### 2020-01-18 code/02-Master.Rmd
Arange preamble

---
title: "AD Target Gene Report"
author: "Jake Gockley"
date: "11/11/2019"
header-includes:
  - \usepackage{multicol}
  - \newcommand{\btwocol}{\begin{multicols}{2}}
  - \newcommand{\etwocol}{\end{multicols}}
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(readr)
library(biomaRt)
library(knitr)
library(kableExtra)
library(tidyverse)
library(grid)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(yaml)
```

```{r yamlIMPT, eval = F, include=FALSE}
#Assign your YAML formated config file to a varable
YAML_Config <- "YAML_FILE"

setwd("~/AD_TargetRank/")
source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
createAndKnitToFolderEntityClient(file = "CMC_ACC_Exp_DataProcessing_V2.Rmd",
                                          parentId ="syn21478397",
                                          folderName = 'ACC')

#Test Genes
Genes <- c('ENSG00000081189','ENSG00000086288', 'ENSG00000137845', 'ENSG00000138600', 'ENSG00000173157', 
           'ENSG00000134463', 'ENSG00000156170', 'ENSG00000120899', 'ENSG00000147065','ENSG00000130203', 
           'ENSG00000165025','ENSG00000174628', 'ENSG00000265148', 'ENSG00000159640', 'ENSG00000154734', 
           'ENSG00000136717', 'ENSG00000095970', 'ENSG00000120885', 'ENSG00000064687')

#Ana's genes
Genes <- c("ENSG00000048392", "ENSG00000051620", "ENSG00000070495", "ENSG00000074800", "ENSG00000077152", "ENSG00000091140", "ENSG00000096060", "ENSG00000100030", "ENSG00000100418", "ENSG00000100605", "ENSG00000105220", "ENSG00000105398", "ENSG00000108424", "ENSG00000109576", "ENSG00000111640", "ENSG00000113657", "ENSG00000117115", "ENSG00000117228", "ENSG00000119125", "ENSG00000121879", "ENSG00000128683", "ENSG00000130540", "ENSG00000136634", "ENSG00000137992", "ENSG00000139112", "ENSG00000140374", "ENSG00000143546", "ENSG00000146701", "ENSG00000146733", "ENSG00000155368", "ENSG00000156709", "ENSG00000162433", "ENSG00000164111", "ENSG00000165025", "ENSG00000166224", "ENSG00000167325", "ENSG00000167720", "ENSG00000168291", "ENSG00000176490", "ENSG00000177511", "ENSG00000186298", "ENSG00000196154", "ENSG00000197448", "ENSG00000197956", "ENSG00000213719")

#Genes <- c('ENSG00000147065','ENSG00000130203')
#Genes <- c('ENSG00000130203')
Tissues <- c('CBE',  'DLPFC' , 'FP', 'IFG', 'PHG', 'STG', 'TCX')
#DE  = 'syn14237651'
#eQTL_Cer = 'syn16984411'
#eQTL_TCX = 'syn16984410'
#eQTL_DLPFC = 'syn16984409'
#eQTL_EasyYesNo = 'syn12514912
#AgoraDrugability = 'syn13363443'
#Proteomics = 'syn18689335'
#AD Associated Mets Networks: syn11932957
#AD Meta Network Modual Associations syn11954640
```

```{r Settings, echo =F, warning=F}
plotdir='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/Plots/'
tabledir='Tables/'
Tissues <- c('CBE', 'DLPFC', 'IFG', 'PHG', 'TCX', 'STG', 'FP')

```

### 2020-01-18 code/Legacy_Profiler.Rmd
Arange preamble

---
title: "AD Target Gene Report"
author: "Jake Gockley"
date: "11/11/2019"
header-includes:
  - \usepackage{multicol}
  - \newcommand{\btwocol}{\begin{multicols}{2}}
  - \newcommand{\etwocol}{\end{multicols}}
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(readr)
library(biomaRt)
library(knitr)
library(kableExtra)
library(tidyverse)
library(grid)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
#library(kableExtra)
#library(dbplyr)
#synapseclient <- reticulate::import("synapseclient")
#syn_temp <- synapseclient$Synapse()
#syn_temp$login()

#MEF2C - ENSG00000081189
#NME8 - ENSG00000086288
#NDUFAF6 - ENSG00000156170 
#ECHDC3 - ENSG00000134463
#ADAMTS20 - ENSG00000173157
#SPPL2A - ENSG00000138600
#ADAM10 - ENSG00000137845
#IQCK - ENSG00000174628
#MIR142/TSPOAP1-AS1 - ENSG00000265148
#ACE - ENSG00000159640
#ADAMTS1 - ENSG00000154734
#BIN1 - ENSG00000136717
#TREM2 - ENSG00000095970
#PTK2B - ENSG00000120899
#CLU - ENSG00000120885
#ABCA7 -

#Test Genes
Genes <- c('ENSG00000081189','ENSG00000086288', 'ENSG00000137845', 'ENSG00000138600', 'ENSG00000173157', 
           'ENSG00000134463', 'ENSG00000156170', 'ENSG00000120899', 'ENSG00000147065','ENSG00000130203', 
           'ENSG00000165025','ENSG00000174628', 'ENSG00000265148', 'ENSG00000159640', 'ENSG00000154734', 
           'ENSG00000136717', 'ENSG00000095970', 'ENSG00000120885', 'ENSG00000064687')

#Ana's genes
Genes <- c("ENSG00000048392", "ENSG00000051620", "ENSG00000070495", "ENSG00000074800", "ENSG00000077152", "ENSG00000091140", "ENSG00000096060", "ENSG00000100030", "ENSG00000100418", "ENSG00000100605", "ENSG00000105220", "ENSG00000105398", "ENSG00000108424", "ENSG00000109576", "ENSG00000111640", "ENSG00000113657", "ENSG00000117115", "ENSG00000117228", "ENSG00000119125", "ENSG00000121879", "ENSG00000128683", "ENSG00000130540", "ENSG00000136634", "ENSG00000137992", "ENSG00000139112", "ENSG00000140374", "ENSG00000143546", "ENSG00000146701", "ENSG00000146733", "ENSG00000155368", "ENSG00000156709", "ENSG00000162433", "ENSG00000164111", "ENSG00000165025", "ENSG00000166224", "ENSG00000167325", "ENSG00000167720", "ENSG00000168291", "ENSG00000176490", "ENSG00000177511", "ENSG00000186298", "ENSG00000196154", "ENSG00000197448", "ENSG00000197956", "ENSG00000213719")

#Genes <- c('ENSG00000147065','ENSG00000130203')
#Genes <- c('ENSG00000130203')
Tissues <- c('CBE',  'DLPFC' , 'FP', 'IFG', 'PHG', 'STG', 'TCX')
#DE  = 'syn14237651'
#eQTL_Cer = 'syn16984411'
#eQTL_TCX = 'syn16984410'
#eQTL_DLPFC = 'syn16984409'
#eQTL_EasyYesNo = 'syn12514912
#AgoraDrugability = 'syn13363443'
#Proteomics = 'syn18689335'
#AD Associated Mets Networks: syn11932957
#AD Meta Network Modual Associations syn11954640
```

```{r Settings, echo =F, warning=F}
plotdir='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/Plots/'
tabledir='Tables/'
Tissues <- c('CBE', 'DLPFC', 'IFG', 'PHG', 'TCX', 'STG', 'FP')

```

### IGAP SNPs

```{r Table1, echo =F, warning=F}
IGAP <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1_2_combined.txt", sep = "\t", header = T)
IGAP1 <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1.txt", sep = "\t", header = T)
```

```{r Table1v2, echo =F, warning=F}
GeneData <- function( id ){

  org = 'hsa'
  id.type = 'ensembl_gene_id'
  #host = 'jul2019.archive.ensembl.org'
  host = 'grch37.ensembl.org'
  ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
  ds <- listDatasets(ensembl)[, "dataset"]
  ds <- grep(paste0("^", org), ds, value = TRUE)

  if (length(ds) == 0){
    stop(paste("Mart not found for:", org))
  } else if (length(ds) > 1) {
    message("Found several marts")
    sapply(ds, function(d) message(paste(which(ds == d), d, sep = ": ")))
    n <- readline(paste0("Choose mart (1-", length(ds), ") : "))
    ds <- ds[as.integer(n)]
  }

  ensembl <- useDataset(ds, mart = ensembl)

  Genes <- getBM(filters = id.type, attributes = c(id.type, 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
  return(Genes)
}

Tab <- GeneData( Genes )
Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
Tab$Interval_hg19 <- NA
for( i in 1:dim(Tab)[1]){
  if(  Tab$start_position[i] < Tab$ensembl_gene_id[i] ){
    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])-1e6, "-", as.numeric(Tab$end_position[i])+1e6)
  }else{
    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])+1e6, "-", as.numeric(Tab$end_position[i])-1e6)
  }
}

Tab$LowestPValCombined <- 1
Tab$TotalSNPsCombined <- 0

Tab$LowestPValPhase1 <- 1
Tab$TotalSNPsPhase1 <- 0

#New Dataframe for the IGAP results that you can track the pval distributions
TotalIGAP <- NULL

#Look at the dataframe 
MedianScoreIGAP <- NULL

#Look at the dataframe 
BestScoreIGAP <- NULL

for( i in 1:dim(Tab)[1]){

  if( isTRUE(as.character(Tab$chromosome_name)[i] %in% as.character(c(1:22))) ){
    Temp <- IGAP[ (as.numeric(as.character(IGAP$Position)) > (as.numeric(as.character(Tab$start_position))[i]-1e6)) &
                  (as.numeric(as.character(IGAP$Position)) < (as.numeric(as.character(Tab$start_position))[i]+1e6)) &
                  as.numeric(as.character(IGAP$Chromosome)) == as.numeric(as.character(Tab$chromosome_name)[i])  , ]
    if( dim(Temp)[1] == 0){
      Temp[1,] <- 0
      Temp$GeneName <- Tab$hgnc_symbol[i]
    }else{
      Temp$GeneName <- Tab$hgnc_symbol[i]
    }
    TotalIGAP <- as.data.frame(rbind(TotalIGAP,Temp))

    TempV1 <- IGAP1[ (as.numeric(as.character(IGAP1$Position)) > (as.numeric(as.character(Tab$start_position))[i]-1e6)) &
                  (as.numeric(as.character(IGAP1$Position)) < (as.numeric(as.character(Tab$start_position))[i]+1e6)) &
                  as.numeric(as.character(IGAP1$Chromosome)) == as.numeric(as.character(Tab$chromosome_name)[i])  , ] 
    TempV1$GeneName <- Tab$hgnc_symbol[i]
    #TotalIGAP <- as.data.frame(rbind(TotalIGAP,TempV1))

      if( median(c(1:length(TempV1$Pvalue))) == floor(median(c(1:length(TempV1$Pvalue)))) & median(c(1:length(TempV1$Pvalue))) == ceiling(median(c(1:length(TempV1$Pvalue)))) ) {
        MedianScoreIGAP <- as.data.frame(rbind(MedianScoreIGAP,TempV1[ median(TempV1$Pvalue) == as.numeric(TempV1$Pvalue), ]))
      }else{

        val <- sort(TempV1$Pvalue)[ floor(median(c(1:length(TempV1$Pvalue)))) ]
        MedianScoreIGAP <- as.data.frame(rbind(MedianScoreIGAP,TempV1[ val == as.numeric(TempV1$Pvalue), ]))
      }

    #BestScoreIGAP <- as.data.frame(rbind(BestScoreIGAP,Temp[ min(Temp$Pvalue) == as.numeric(Temp$Pvalue), ]))

    if( dim(Temp)[1] == 0 ){

    }else{
        Tab$LowestPValCombined[i] <- min( Temp$Pvalue )
        Tab$TotalSNPsCombined[i] <- paste0("N=", dim(Temp)[1] )
    }
    if( dim(TempV1)[1] == 0 ){
    }else{
        Tab$LowestPValPhase1[i] <- min( TempV1$Pvalue )
        Tab$TotalSNPsPhase1[i] <- paste0("N=", dim(TempV1)[1] )
    }
  }else{
    #Temp <- as.data.frame(matrix(0, 1, 9))
    #colnames(Temp) <- c( 'Chromosome', 'Position', 'MarkerName', 'Effect_allele', 'Non_Effect_allele', 'Beta', 'SE', 'Pvalue', 'GeneName' ) 
    #Temp$GeneName <- Tab[ i, ]$GeneName
    #Temp$Pvalue <- 1
    #BestScoreIGAP <- as.data.frame(rbind(BestScoreIGAP,Temp[ 1, ]))

  }
  if( dim(Temp)[1] == 0){
    Temp <- as.data.frame(matrix(0, 1, 10))
    colnames(Temp) <- c( 'Chromosome', 'Position', 'MarkerName', 'Effect_allele', 'Non_Effect_allele', 'Beta', 'SE', 'Pvalue', 'Rank', 'GeneName' ) 
    Temp$GeneName <- Tab[ i, ]$hgnc_symbol
    Temp$Pvalue <- 1
    BestScoreIGAP <- as.data.frame(rbind(BestScoreIGAP,Temp[ 1, ]))
  }else{
    BestScoreIGAP <- as.data.frame(rbind(BestScoreIGAP,Temp[ min(Temp$Pvalue) == as.numeric(Temp$Pvalue), ]))
  }
}
colnames(Tab) <- c("ENSG", "GeneName", "Start", "End", "Chr", "Coordhg19", "Intervalhg19", "LowestPValCombined", "TotalSNPsCombined", "LowestPValPhaseI", "TotalSNPsPhaseI" )
row.names(Tab) <- Tab$ENSG

TotalIGAP$LogP <- -log(TotalIGAP$Pvalue)

p <- ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) + 
  geom_violin() + geom_boxplot(width=0.1) +
  geom_jitter(shape=16, size=.1, position=position_jitter(0.2)) +
  theme(axis.text.x = element_text(angle = 45) ) + ylab('-Log(P-Value)')  
#p

setEPS()
postscript(file = paste0(plotdir,'/TotalIGAP.eps') )
  ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) + 
  geom_violin() + geom_boxplot(width=0.1) +
  geom_jitter(shape=16, size=.1, position=position_jitter(0.2)) +
  theme(axis.text.x = element_text(angle = 45) ) + ylab('-Log(P-Value)') 
dev.off()

IGAP$Rank<-rank(-IGAP$Pvalue)
FOO<-IGAP
FOO$Pvalue <- -log(FOO$Pvalue)
FOO$RankModel <- (FOO$Rank/max(FOO$Rank))

Pr <- ggplot( FOO, aes(x=Pvalue, y=RankModel)) +
  geom_point() + xlab("-log(P-Value") + geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
#Pr
setEPS()
postscript(file = paste0(plotdir,'/IGAP_WeightModel.eps') )
   ggplot( FOO, aes(x=Pvalue, y=RankModel)) +
  geom_point() + xlab("-log(P-Value") + geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
dev.off()

#Fit Logistic Model:
mylogit <- glm(RankModel ~ Pvalue, data = FOO, family = "binomial")

LogisticScorer <- function( X ){
  y <- 1/(1+exp( -( -3.305501+0.420491*X) ))
  return(y)
}

FOO$Weights <- LogisticScorer(FOO$Pvalue)
#Pr <- ggplot( FOO, aes(x=Pvalue, y=Weights)) +
#  geom_point() + xlab("-log(P-Value)") + geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)

Temp <- FOO[ ,c(1:10)]
Temp$Type <- "Actual Rank"
colnames(Temp)[10] <- "y"
temp <- FOO[ ,c(1:9,11)] 
temp$Type <- "Predicted Weight"
colnames(temp)[10] <- "y"

Total <- rbind(Temp,temp)
Pt <- ggplot( Total, aes(x=Pvalue, y=y, col=Type)) + scale_color_manual(values=c( "#E69F00", "#56B4E9")) +
  geom_point() + xlab("-log(P-Value)") + geom_smooth(method = "glm", method.args = list(family = "binomial"), colour="black", size=0.5, se = FALSE)  
#Pt

setEPS()
postscript(file = paste0(plotdir,'/Fitted_IGAP_WeightModel.eps') )
  ggplot( Total, aes(x=Pvalue, y=y, col=Type)) + scale_color_manual(values=c( "#E69F00", "#56B4E9")) +
  geom_point() + xlab("-log(P-Value)") + geom_smooth(method = "glm", method.args = list(family = "binomial"), colour="black", size=0.5, se = FALSE)  
dev.off()

#Remove the pval ties
BestScoreIGAP <- BestScoreIGAP[ !duplicated(BestScoreIGAP$GeneName), ]
BestScoreIGAP$Wts <- LogisticScorer(-log(BestScoreIGAP$Pvalue))
BestScoreIGAP$LogP <- -log(BestScoreIGAP$Pvalue)

Py <- ggplot( BestScoreIGAP, aes(x=LogP, y=Wts, col=GeneName)) +
  geom_point() + xlab("-log(P-Value)") + ylab("IGAP Weight") + geom_smooth( data=Total, aes(x=Pvalue, y=y, col=Type), method = "glm", method.args = list(family = "binomial"), colour="black", size=0.5, se = FALSE) + geom_label_repel(aes(label = GeneName),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50') 
#Py
setEPS()
postscript(file = paste0(plotdir,'/Best_IGAP_Weight_Score.eps') )
  ggplot( BestScoreIGAP, aes(x=LogP, y=Wts, col=GeneName)) +
  geom_point() + xlab("-log(P-Value)") + ylab("IGAP Weight") + geom_smooth( data=Total, aes(x=Pvalue, y=y, col=Type), method = "glm", method.args = list(family = "binomial"), colour="black", size=0.5, se = FALSE) + geom_label_repel(aes(label = GeneName),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')  
dev.off()

lay <- rbind(c(1,2),
             c(3,4))

g1 <- ggplotGrob( p)
g2 <- ggplotGrob( Pr)
g3 <- ggplotGrob( Pt)
g4 <- ggplotGrob( Py )

gs = list( g1, g2, g3, g4 )

```

```{r Table1v3, echo =F, warning=F, eval= F}

plot( grid.arrange(grobs = gs, layout_matrix = lay, top = "IGAP Weight Models Applied to Lowest SNP Within 1MB of Combined Phase I + II)") )

```

```{r Table1v4, echo =F, warning=F}

##Plot Data Table
Tab %>% 
  mutate(
    LowestPValCombined = cell_spec(LowestPValCombined, bold = ifelse(LowestPValCombined < 0.05, TRUE, FALSE ), color = ifelse(LowestPValCombined < 0.05, "red", ifelse(LowestPValCombined == 1, "white", "black") )),
    LowestPValPhaseI = cell_spec(LowestPValPhaseI, bold = ifelse(LowestPValPhaseI < 0.05, TRUE, FALSE ), color = ifelse(LowestPValPhaseI < 0.05, "red", ifelse(LowestPValPhaseI == 1, "white", "black") ))
  ) %>%
  select( ENSG, GeneName, Start, End, Chr, Coordhg19, Intervalhg19, LowestPValCombined, TotalSNPsCombined, LowestPValPhaseI, TotalSNPsPhaseI)  %>%
  kable( escape = F,  format = "html") %>%
  kable_styling("striped", full_width = F)

save_kable(
  Tab %>% 
  mutate(
    LowestPValCombined = cell_spec(LowestPValCombined, bold = ifelse(LowestPValCombined < 0.05, TRUE, FALSE ), color = ifelse(LowestPValCombined < 0.05, "red", ifelse(LowestPValCombined == 1, "white", "black") )),
    LowestPValPhaseI = cell_spec(LowestPValPhaseI, bold = ifelse(LowestPValPhaseI < 0.05, TRUE, FALSE ), color = ifelse(LowestPValPhaseI < 0.05, "red", ifelse(LowestPValPhaseI == 1, "white", "black") ))
  ) %>%
  select( ENSG, GeneName, Start, End, Chr, Coordhg19, Intervalhg19, LowestPValCombined, TotalSNPsCombined, LowestPValPhaseI, TotalSNPsPhaseI)  %>%
  kable( escape = F,  format = "html") %>%
  kable_styling("striped", full_width = F),

  file=paste0(tabledir,'/IGAPStats.pdf')
)

```

### Differential Expression

```{r  DEProcessing, echo =F, warning=F, include=F}
DE <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/differentialExpressionSummary.tsv",
                  header=T, sep ="\t")

ModelConstruct <- function( de, Comp, Mod, Sex ){
  #C#reates a model and plots the target genes
  #'@de the differential  expression Construct (DE)
  #'@Comp The comparison ie NA or "AD-CONTROL" 
  #'@Mod The model is NA or 'Diagnosis'
  #'@Sex The Sex Comparison ie NA or 'MALE' or 'FEMALE'

  #Find the gene expression for the value of the genes where it exists
  if( isTRUE(is.na(Comp)) ){
  }else{
    de <- de[ de$Comparison == Comp ,  ] 
  }
  if( isTRUE(is.na(Mod)) ){
  }else{
    de <- de[ de$Model == Mod ,   ] 
  }
  if( isTRUE(is.na(Sex)) ){
  }else{
    de <- de[ de$Sex == Sex ,  ] 
  }
  de <- de[ , c('hgnc_symbol', 'ensembl_gene_id','Tissue','logFC','adj.P.Val') ]

  Models <- list()
  Plots <- list()

  #Loop through induvidual tissues
  for( Tissue in c('CBE', 'DLPFC', 'FP', "IFG", "PHG", "STG", "TCX" )){
    Work <- de[ de$Tissue == Tissue, ]
    Work$y <- rank(abs( Work$logFC)) / max(rank(abs( Work$logFC)))
    Work$Type <- "Actual Rank"
    #plot( log(abs( Work$logFC)), Work$y )

    #Fit Logistic Model:
    mylogit <- glm(Work$y  ~ abs( Work$logFC), data = FOO, family = "binomial")

    Work$log_abs_PVal <- abs( Work$logFC)

    Work2 <- Work  
    Work2$Type <- "Predicted Weight"
    #Work2$y <- y <- 1/(1+exp( -( mylogit$coefficients[1]+mylogit$coefficients[2]* abs( Work$logFC) ) ))
    Work2$y <- 1/(1+exp( -( mylogit$coefficients[1]+mylogit$coefficients[2]* abs( Work$logFC) ) ))

    tWork <- as.data.frame(rbind( Work ,Work2 ))
    tWork$Sig <- ifelse(tWork$adj.P.Val < 0.05, "YES", "NO" )
    eval(parse( text=paste0( 'Models$', Tissue, ' <- as.data.frame( tWork )' ) ))

    #Colors for the sig and non-sig names
    COLS <- c( "blueviolet","grey29")
    names(COLS) <- c("YES","NO")

    #Plot the model
    if( Tissue == 'CBE' ){
      eval(parse( text=paste0( 
        'Pr <- ggplot( data=Models$', Tissue, '[ Models$', Tissue, '$Type == \'Predicted Weight\' & (Models$', Tissue, '$ensembl_gene_id %in% Genes) == T , ], aes(x=log_abs_PVal, y=y, col=Sig)) +
              geom_point() + 
              geom_smooth( data=Models$', Tissue, '[ Models$', Tissue, '$Type == \'Predicted Weight\', ], aes(x=log_abs_PVal, y=y), 
                    method = "glm", 
                    method.args = list(family = "binomial"), 
                    colour="black", size=0.5, se = FALSE)  + 
              xlab("log(abs(LogFC))") + ylab("Model Weight") + ggtitle("',Tissue,'") +
              theme(plot.title = element_text(vjust = -60, hjust = .9, size = 18, face="bold") ) +
              geom_label_repel(aes(label = hgnc_symbol, col = Sig),
                    box.padding   = 0.35, 
                    point.padding = 0.5,
                    segment.color = \'grey50\') + 
              scale_colour_manual( values = COLS ) + 
              theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
                  axis.text.x=element_blank(), axis.ticks.x=element_blank() )'
      )))
      eval(parse( text=paste0( 'Plots[["', Tissue, '"]] <- Pr' ) ))
    }else{
      eval(parse( text=paste0( 
        'Pr <- ggplot( data=Models$', Tissue, '[ Models$', Tissue, '$Type == \'Predicted Weight\' & (Models$', Tissue, '$ensembl_gene_id %in% Genes) == T , ], aes(x=log_abs_PVal, y=y, col=Sig)) +
              geom_point() + 
              geom_smooth( data=Models$', Tissue, '[ Models$', Tissue, '$Type == \'Predicted Weight\', ], aes(x=log_abs_PVal, y=y), 
                    method = "glm", 
                    method.args = list(family = "binomial"), 
                    colour="black", size=0.5, se = FALSE)  + 
              xlab("log(abs(LogFC))") + ylab("Model Weight") + ggtitle("',Tissue,'") +
              theme(plot.title = element_text(vjust = -60, hjust = .9, size = 18, face="bold") ) +
              geom_label_repel(aes(label = hgnc_symbol, col = Sig),
                    box.padding   = 0.35, 
                    point.padding = 0.5,
                    segment.color = \'grey50\') + 
              scale_colour_manual( values = COLS ) + 
              theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
                  axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank())'
      )))
      eval(parse( text=paste0( 'Plots[["', Tissue, '"]] <- Pr' ) ))

      ggsave( filename = paste0('DiffExpress_', Sex, '_',Tissue,'.eps'),
              plot = Pr,
              device = "eps",
              path = plotdir
      )

    }
  }
  return( list( Plots=Plots, Models=Models ) )
}

#Plot/Run/Build the DE Models

ALL_Mod <- ModelConstruct( DE, 'AD-CONTROL', 'Diagnosis', 'ALL' )
Male_Mod <- ModelConstruct( DE, NA, NA, 'MALE' )
Femal_Mod <- ModelConstruct( DE, NA, NA, 'FEMALE' )

REL <- DE[ (DE$ensembl_gene_id %in% Genes) == T , ]
colnames(REL)[ colnames(REL) == 'ensembl_gene_id'] <- 'ensemblgeneid'
colnames(REL)[ colnames(REL) == 'hgnc_symbol'] <- 'hgncsymbol'

ALL <- REL[ REL$Sex == 'ALL' & REL$Comparison == "AD-CONTROL" & REL$Model == 'Diagnosis',  c('hgncsymbol','ensemblgeneid','Tissue','logFC','adj.P.Val')] 
Males <-  REL[ REL$Sex == 'MALE' ,  c('hgncsymbol','ensemblgeneid','Tissue','logFC','adj.P.Val')]
Females <-  REL[ REL$Sex == 'FEMALE',  c('hgncsymbol','ensemblgeneid','Tissue','logFC','adj.P.Val')]

colnames(Tab)[ colnames(Tab) == 'Gene_Name' ] <- 'GeneName'
DE_Plots <- function( INPUT ){
  #Plots the table of the DE of the gene list
  #'@INPUT The input differential expression objects e.g ALL
  Gen <- as.character(INPUT$ensemblgeneid)[ !duplicated(as.character(INPUT$ensemblgeneid))]
  DF <- as.data.frame(matrix(0,length(Gen),9))
  row.names( DF ) <- Gen
  colnames( DF ) <- c( 'ENSG', 'GeneName', Tissues)
  DF$ENSG <- Gen
  DF$GeneName <- Tab[ DF$ENSG, ]$GeneName

  for( row in row.names(DF)){
    for( col in colnames(DF)[ colnames(DF) %in% c("CBE", "DLPFC", "FP", "IFG", "PHG", "STG", "TCX") ]){
      #Test if the the gene and tissue exist:
      if( row %in% as.character( INPUT[  INPUT$Tissue == col, ]$ensemblgeneid ) ){
        DF[ row, col ] <- signif( INPUT[ INPUT$ensemblgeneid == row & INPUT$Tissue == col, ]$logFC, 3)
      }else{
        DF[ row, col ] <- NA
      }
    }
  }

  #P Vals
  colnames(Tab)[ (colnames(Tab) %in% 'Gene_Name') == T ] <- 'GeneName'
  pDF <- as.data.frame(matrix(0,length(Gen),9))
  row.names( pDF ) <- Gen
  colnames( pDF ) <- c( 'ENSG', 'GeneName', Tissues)
  pDF$ENSG <- Gen
  pDF$GeneName <- Tab[ pDF$ENSG , ]$GeneName

  for( row in row.names(pDF)){
    for( col in colnames(pDF)[ colnames(pDF) %in% c("CBE", "DLPFC", "FP", "IFG", "PHG", "STG", "TCX") ]){
      if( row %in% as.character( INPUT[  INPUT$Tissue == col, ]$ensemblgeneid ) ){
        pDF[ row, col ] <- INPUT[ INPUT$ensemblgeneid == row & INPUT$Tissue == col, ]$adj.P.Val
      }else{
        pDF[ row, col ] <- 40
      }
    }
  }

  return( DF %>% 
    mutate(
      CBE = cell_spec( CBE, bold = ifelse(pDF$CBE < 0.05, TRUE, FALSE ), color = ifelse(pDF$CBE < 0.05, ifelse(DF$CB < 0, "red", "green"), rgb(211/255,211/255,211/255)) ) ,
      DLPFC = cell_spec(DLPFC, bold = ifelse(pDF$DLPFC < 0.05, TRUE, FALSE ),color = ifelse(pDF$DLPFC < 0.05, ifelse(DF$DLPFC < 0, "red", "green"), rgb(211/255,211/255,211/255) )),
      FP = cell_spec(FP, bold = ifelse(pDF$FP < 0.05, TRUE, FALSE ),color = ifelse(pDF$FP < 0.05, ifelse(DF$FP < 0, "red", "green"), rgb(211/255,211/255,211/255) )),
      IFG = cell_spec(IFG, bold = ifelse(pDF$IFG < 0.05, TRUE, FALSE ),color = ifelse(pDF$IFG < 0.05, ifelse(DF$IFG < 0, "red", "green"), rgb(211/255,211/255,211/255) )),
      PHG = cell_spec(PHG, bold = ifelse(pDF$PHG < 0.05, TRUE, FALSE ), color = ifelse(pDF$PHG < 0.05, ifelse(DF$PHG < 0, "red", "green"), rgb(211/255,211/255,211/255) )),
      STG = cell_spec(STG, bold = ifelse(pDF$STG < 0.05, TRUE, FALSE ),color = ifelse(pDF$STG < 0.05, ifelse(DF$STG < 0, "red", "green"), rgb(211/255,211/255,211/255) )),
      TCX = cell_spec(TCX, bold = ifelse(pDF$TCX < 0.05, TRUE, FALSE ),color = ifelse(pDF$TCX < 0.05, ifelse(DF$TCX < 0, "red", "green"), rgb(211/255,211/255,211/255) ))
    ) %>%
    kable( escape = F, row.names = FALSE) %>%
    kable_styling("striped", full_width = FALSE, position="left") )

}
```

#### All Cases Versus Controls

```{r  ALL_DE, echo=FALSE, warning=F, fig.width=12, fig.height=6, warning=F }

DE_Plots(ALL)
save_kable(
  DE_Plots(ALL),
  file=paste0(tabledir,'/DiffExp_ALL.pdf')
)
```

```{r  ALL_DE2, echo=FALSE, fig.width=12, fig.height=6, warning=F }

lay <- rbind(c(1,2,3,4,7),
             c(1,2,5,6,7))

g1 <- ggplotGrob(ALL_Mod$Plots$CBE)
g2 <- ggplotGrob(ALL_Mod$Plots$DLPFC)
g3 <- ggplotGrob( ALL_Mod$Plots$FP)
g4 <- ggplotGrob(ALL_Mod$Plots$IFG)
g5 <- ggplotGrob(ALL_Mod$Plots$PHG)
g6 <- ggplotGrob( ALL_Mod$Plots$STG)
g7 <- ggplotGrob(ALL_Mod$Plots$TCX)

gs = list( g1, g2, g3, g4, g5, g6, g7 )

plot( grid.arrange(grobs = gs, layout_matrix = lay, left = "Calculated Weight", bottom = "Absolute Value of Log(Fold Change)") )
```

#### Female Cases Versus Controls
```{r  Female_DE, echo =F, fig.width=12, fig.height=6, warning=F}

DE_Plots(Females)
save_kable(
  DE_Plots(Females),
  file=paste0(tabledir,'/DiffExp_Females.pdf')
)

lay <- rbind(c(1,2,3,4,7),
             c(1,2,5,6,7))

g1 <- ggplotGrob(Femal_Mod$Plots$CBE)
g2 <- ggplotGrob(Femal_Mod$Plots$DLPFC)
g3 <- ggplotGrob( Femal_Mod$Plots$FP)
g4 <- ggplotGrob(Femal_Mod$Plots$IFG)
g5 <- ggplotGrob(Femal_Mod$Plots$PHG)
g6 <- ggplotGrob( Femal_Mod$Plots$STG)
g7 <- ggplotGrob(Femal_Mod$Plots$TCX)

gs = list( g1, g2, g3, g4, g5, g6, g7 )

grid.arrange(grobs = gs, layout_matrix = lay)

```

#### Male Cases Versus Controls

```{r  Male_DE, echo =F, fig.width=12, fig.height=6, warning=F}

DE_Plots(Males)
save_kable(
  DE_Plots(Males),
  file=paste0(tabledir,'/DiffExp_Males.pdf')
)
```

```{r  Male_DEPlot, echo =F, fig.width=12, fig.height=6, warning=F}
lay <- rbind(c(1,2,3,4,7),
             c(1,2,5,6,7))

g1 <- ggplotGrob(Male_Mod$Plots$CBE)
g2 <- ggplotGrob(Male_Mod$Plots$DLPFC)
g3 <- ggplotGrob( Male_Mod$Plots$FP)
g4 <- ggplotGrob(Male_Mod$Plots$IFG)
g5 <- ggplotGrob(Male_Mod$Plots$PHG)
g6 <- ggplotGrob( Male_Mod$Plots$STG)
g7 <- ggplotGrob(Male_Mod$Plots$TCX)

gs = list( g1, g2, g3, g4, g5, g6, g7 )

grid.arrange(grobs = gs, layout_matrix = lay)
```

### Differential Proteomics

```{r Proteomics, echo =F, warning=F }
Prot <- read.csv( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/Final_Agora_DE.csv", header = T)
Prot <- Prot[ complete.cases(Prot[,c("Log2_FC", "CI_Upr", "CI_Lwr", "PVal", "Cor_PVal")]), ]
Prot <- Prot[complete.cases(Prot),]

TargetP <- Prot[ Prot$ENSG %in% Genes,]

Names <- paste0(TargetP$UniqID, '|', TargetP$ENSG)
Names <- Names[!duplicated(Names)]
PTab <- do.call( rbind, strsplit( Names, '\\|' ))

if( table(Genes %in% PTab[,3])['TRUE'] == length(Genes) ){
  Res <- PTab
}else{
  Adds <- Genes[ (Genes %in% PTab[,3]) == F ]
  Res <- rbind( PTab, cbind( Tab[Adds,]$GeneName, NA, Adds ))
}

row.names(Res) <- paste0( Res[,1], '|', Res[,2])
Res <- as.data.frame(Res)
colnames( Res ) <- c( "GeneName", "UniProt", "ENSG")

Tiss <- as.character(TargetP$Tissue)[ !duplicated(as.character(TargetP$Tissue)) ]
Adds <- matrix(0, dim(Res)[1], 4)
row.names(Adds) <- row.names(Res)
colnames(Adds) <- c("AntPFC", "DLPFC", "MFG", "TCX" )

TotP <- as.data.frame(cbind(Res, Adds))
TotPval <- as.data.frame(cbind(Res, Adds))

for( row in row.names(TotP)){
  for( col in colnames(TotP)){
    if( isTRUE(row %in%TargetP$UniqID) ){
      TP <- TargetP[ TargetP$UniqID == row,]
      if( isTRUE(col %in% TP$Tissue) ){
        TotP[row,col] <- signif( TP[ as.character(TP$Tissue)==col ,]$Log2_FC, 3 )
      }else{}
    }else{

    }
  }
}

for( row in row.names(TotPval)){
  for( col in colnames(TotPval)){
    if( isTRUE(row %in%TargetP$UniqID) ){
      TP <- TargetP[ TargetP$UniqID == row,]
      if( isTRUE(col %in% TP$Tissue) ){
        TotPval[row,col] <- TP[ as.character(TP$Tissue)==col ,]$Cor_PVal
      }else{}
    }else{

    }
  }
}

TotP %>% 
    mutate(
      AntPFC = cell_spec(AntPFC, bold = ifelse(TotPval$AntPFC < 0.05, TRUE, FALSE ),color = ifelse(TotPval$AntPFC == 0, "white", ifelse(TotPval$AntPFC < 0.05, ifelse(TotP$AntPFC < 0, "red", "green"), "grey" ))),
      DLPFC = cell_spec(DLPFC, bold = ifelse(TotPval$DLPFC < 0.05, TRUE, FALSE ), color = ifelse(TotPval$DLPFC == 0, "white", ifelse(TotPval$DLPFC < 0.05, ifelse(TotP$DLPFC < 0, "red", "green"), "grey" ))),
      MFG = cell_spec(MFG, bold = ifelse(TotPval$MFG < 0.05, TRUE, FALSE ),color = ifelse(TotPval$MFG == 0, "white", ifelse(TotPval$MFG < 0.05, ifelse(TotP$MFG < 0, "red", "green"), "grey" ))),
      TCX = cell_spec(TCX, bold = ifelse(TotPval$TCX < 0.05, TRUE, FALSE ),color = ifelse(TotPval$TCX == 0, "white", ifelse(TotPval$TCX < 0.05, ifelse(TotP$TCX < 0, "red", "green"), "grey" )))
    ) %>%
    select(GeneName, UniProt, ENSG, AntPFC, DLPFC, MFG, TCX)  %>%
    kable( escape = F) %>%
    kable_styling("striped", full_width = F)

save_kable(
  TotP %>% 
    mutate(
      AntPFC = cell_spec(AntPFC, bold = ifelse(TotPval$AntPFC < 0.05, TRUE, FALSE ),color = ifelse(TotPval$AntPFC == 0, "white", ifelse(TotPval$AntPFC < 0.05, ifelse(TotP$AntPFC < 0, "red", "green"), "grey" ))),
      DLPFC = cell_spec(DLPFC, bold = ifelse(TotPval$DLPFC < 0.05, TRUE, FALSE ), color = ifelse(TotPval$DLPFC == 0, "white", ifelse(TotPval$DLPFC < 0.05, ifelse(TotP$DLPFC < 0, "red", "green"), "grey" ))),
      MFG = cell_spec(MFG, bold = ifelse(TotPval$MFG < 0.05, TRUE, FALSE ),color = ifelse(TotPval$MFG == 0, "white", ifelse(TotPval$MFG < 0.05, ifelse(TotP$MFG < 0, "red", "green"), "grey" ))),
      TCX = cell_spec(TCX, bold = ifelse(TotPval$TCX < 0.05, TRUE, FALSE ),color = ifelse(TotPval$TCX == 0, "white", ifelse(TotPval$TCX < 0.05, ifelse(TotP$TCX < 0, "red", "green"), "grey" )))
    ) %>%
    select(GeneName, UniProt, ENSG, AntPFC, DLPFC, MFG, TCX)  %>%
    kable( escape = F) %>%
    kable_styling("striped", full_width = F),

  file=paste0(tabledir,'/Diff_Protiens.pdf')
)

```

### eQTL

```{r eQTL, echo = F, warning=F}
####Need eQTL AD analysis from Solly!!!

#syn12514912
eqtl <- read.csv( file ="/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/eqtl_meta_analysis.csv", header=T)
uEQs <- eqtl[ (eqtl$ensembl_gene_id %in% Genes) == T,]

uEQs$hasEqtl <- as.character(uEQs$hasEqtl)
Rep <- c("NO", "YES")
names(Rep) <- c('FALSE', 'TRUE')
uEQs$hasEqtl <- as.character(Rep[ as.character(uEQs$hasEqtl) ])

```

### AD Associated Module Clustering from Meta-Network

```{r Network, echo = F, warning=F}
Nets <- read.csv(file ="/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/Network.csv", header=T)
ZoomNet <- Nets[ (Nets$GeneID %in% Genes) == T, ]

NetPlot <- as.data.frame(matrix(0,length(Genes),5))
row.names(NetPlot) <- Genes
NetPlot$V1 <- row.names(NetPlot)
colnames(NetPlot) <- c("ENSG", "GeneName", "InACluster", "NumberOfClusters", "BrainRegions")

for( g in Genes ){
  if( dim(ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ])[1] > 0 ){
    NetPlot[g,]$GeneName <- names( table( as.character( ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ]$external_gene_name ) ) )
    NetPlot[g,]$InACluster <- 'YES'
    tmp <- ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ]
    NetPlot[g,]$NumberOfClusters <- length(tmp$Module)
    Regions <- c(as.character(tmp$brainRegion))
    NetPlot[g,]$BrainRegions <- paste( Regions[ !duplicated(Regions) ], collapse = ', ')
  }else{
    NetPlot[g,]$GeneName <- as.character( TotP[ TotP$ENSG == g, ]$GeneName )
    NetPlot[g,]$InACluster <- 'NO'
    NetPlot[g, 3:5] <- NA
  }
}

NetPlot %>% 
    select(ENSG, GeneName, InACluster, NumberOfClusters, BrainRegions)  %>%
    kable( escape = F, row.names = FALSE) %>%
    kable_styling("striped", full_width = F)

save_kable(
  NetPlot %>% 
    select(ENSG, GeneName, InACluster, NumberOfClusters, BrainRegions)  %>%
    kable( escape = F, row.names = FALSE) %>%
    kable_styling("striped", full_width = F),

  file=paste0(tabledir,'/InNetCluster.pdf')
)

Regions <- as.data.frame( table(ZoomNet$brainRegion) )
colnames(Regions) <- c("Region", "Targets")

Targets <-as.data.frame(Regions)
Targets %>% 
    select(Region, Targets)  %>%
    kable( escape = F, row.names = FALSE) %>%
    kable_styling("striped", full_width = F)

save_kable(
  Targets %>% 
    select(Region, Targets)  %>%
    kable( escape = F, row.names = FALSE) %>%
    kable_styling("striped", full_width = F),

  file=paste0(tabledir,'/Targets_by_Region.pdf')
)

```

### AD Associated Module Clustering
```{r Heatmap, echo = F, warning=F}
library(gplots)
library(superheat)

Mods <- as.character(ZoomNet$Module)
Mods <- Mods[ !duplicated(Mods) ]

dim( table( ZoomNet[ ZoomNet$brainRegion == 'DLPFC', ]$ModuleName,  ZoomNet[ ZoomNet$brainRegion == 'DLPFC', ]$external_gene_name ) )

HM_Mat <- matrix( 0, length(Genes), length(Mods) )
row.names(HM_Mat) <- as.character(NetPlot$GeneName)
colnames(HM_Mat) <- Mods

trans <- NetPlot$GeneName
names(trans) <- NetPlot$ENSG

for( i in 1:dim(ZoomNet)[1] ){
  HM_Mat[ trans[as.character( ZoomNet[ i, ]$GeneID )], 
          as.character( ZoomNet[ i, ]$Module ) ] <- 1
}

PAL <- colorRampPalette(c( "white", "red"))(n = 2)
heatmap.2(as.matrix(HM_Mat), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)

setEPS()
postscript(file = paste0(plotdir,'/ModuleHeatMap.eps') )
  heatmap.2(as.matrix(HM_Mat), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)
dev.off()

```

```{r ModuleFunction, echo = F, warning=F}
#Figure 2A: https://www.biorxiv.org/content/biorxiv/early/2019/01/03/510420.full.pdf
Consensus_Cluster_A <- c("TCXblue", "IFGyellow", "PHGyellow")
Consensus_Cluster_B <- c("DLPFCblue", "CBEturquoise", "STGblue", "PHGturquoise", "IFGturquoise", "TCXturquoise", "FPturquoise")
Consensus_Cluster_C <- c("IFGbrown", "STGbrown", "DLPFCyellow", "TCXgreen", "FPyellow", "CBEyellow", "PHGbrown" )
Consensus_Cluster_D <- c("DLPFCbrown", "STGyellow", "PHGgreen", "CBEbrown", "TCXyellow", "IFGblue", "FPblue")
Consensus_Cluster_E <- c("FPbrown", "CBEblue", "DLPFCturquoise", "TCXbrown", "STGturquoise", "PHGblue")

ConClust <- as.data.frame( cbind( c(Consensus_Cluster_A, Consensus_Cluster_B, Consensus_Cluster_C, Consensus_Cluster_D, Consensus_Cluster_E),
                      c( rep("A", length(Consensus_Cluster_A)) ,rep("B", length(Consensus_Cluster_B)), rep("C", length(Consensus_Cluster_C)),
                      rep("D", length(Consensus_Cluster_D)), rep("E", length(Consensus_Cluster_E))) ))

AllGenes <- Nets$GeneID[ !duplicated(Nets$GeneID) ]

colnames(ConClust) <- c("Module", "ConesusCluster")
row.names(ConClust) <- ConClust$Module

#-#TempClust <- as.data.frame( matrix(0, length(Genes), 5))
TempClust <- as.data.frame( matrix(0, length(AllGenes), 5))
#-#row.names(TempClust) <- Genes
row.names(TempClust) <- AllGenes
colnames(TempClust) <- c("A", "B", "C", "D", "E")

for( i in AllGenes){
  #tmp_cs <- table( as.character( ConClust[ as.character(ZoomNet[ ZoomNet$GeneID == i, ]$Module), ]$ConesusCluster ) )
  tmp_cs <- table( as.character( ConClust[ as.character(Nets[ Nets$GeneID == i, ]$Module), ]$ConesusCluster ) )
  for(y in names(tmp_cs)){
    TempClust[ i, y] <- as.numeric(tmp_cs[y])
  }
}

#-#row.names(TempClust) <- as.character( trans[ row.names(TempClust)] )
tNets <- Nets[ !duplicated(Nets$GeneID), ]
row.names(tNets) <- tNets$GeneID

#row.names(TempClust) <-tNets[ row.names(tNets),]$external_gene_name

#Fix the repeats...
PAL <- colorRampPalette(c( "white", "yellow", "red"))(n = 10)

HM <- heatmap.2(as.matrix(TempClust), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)
hc <- hclust(dist(TempClust), "complete")
plot(hc, xlab="", ylab="", main="", sub="", axes=FALSE)
par(cex=1)

#TC.pca <- prcomp(t(TempClust), center = TRUE)
#summary( TC.pca )

#plot(TC.pca$)

Tempclust <- TempClust[rownames(TempClust) %in% Genes,] 
row.names(Tempclust) <- trans[row.names(Tempclust)]

save_kable(
  Tempclust %>% 
    select( A, B, C, D, E)  %>%
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F),
  file=paste0(tabledir,'/ConsensousCluster.pdf')
)

Tempclust %>% 
    select( A, B, C, D, E)  %>%
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F)

#####
#TotalCls <- table( Nets$GeneID, Nets$Module )
#TC.pca <- prcomp(TotalCls, center = TRUE)
#summary( TC.pca )
#dim(TC.pca$x)
#plot(TC.pca$x[,1], TC.pca$x[,2])

#screeplot( TC.pca )

#library(irlba)
#library(cluster)
#library(tidyverse)  # data manipulation
#library(factoextra)
#library(sparcl)

#foo <- KMeansSparseCluster(FOO, K=10, wbounds = NULL, nstart = 20, silent =
#FALSE, maxiter=6)
#plot(foo)

#Clus <- hclust(dist(TotalCls), "complete")
#Dis <- as.dist(1 - cor(t(TotalCls)))
#plot(hclust(Dis), main="Dissimilarity = 1 - Correlation", labels=FALSE)

#plot(hclust(dist(TotalCls)), labels=FALSE)

#s <- irlba(TotalCls, nu = 0, nv=10)
#TotalCls_reduced <- as.matrix(TotalCls %*% s$v)
#clust_kmeans <- kmeans(TotalCls_reduced, 10)
#summary(silhouette(clust_kmeans$cluster, dist(TotalCls_reduced)))

#FOO <- matrix( 0, dim(TotalCls)[1], dim(TotalCls)[2] )
#row.names(FOO) <- row.names(TotalCls)
#colnames(FOO) <- colnames(TotalCls)

#for( i in 1:30 ){
#  FOO[,i] <- TotalCls[,i]
#}

#Dis <- dist(FOO)
#Cor <- cor(t(FOO))
#
#Att <- as.vector(Cor)

#N <- 0

#Att <- as.list(Cor)
#foo<- as.data.frame(as.matrix(Att))

#Meh <- NULL
#for( i in 1:length(row.names(Cor))){ Meh <- c(Meh, rep( row.names(Cor)[i], length(row.names(Cor)) )) }
#foo$V3 <- rep( row.names(Cor), length(row.names(Cor)) ) 
#foo$V2 <- Meh

#for( i in 1:(dim(Cor)[1]-1) ){
#  N <- N+i
#}
#Comps <- as.data.frame( matrix( 0, 0, 3 ) )

#Rs <- row.names(Cor)
#Cs <- colnames(Cor)
#i <- 0

#for( Name in Rs ){
#  temp <- Cs[ (Cs %in% Name) == F ]
#  for( Other in temp){
    #i <- i+1
    #Comps <- rbind( Comps[ 1:dim(Comps)[1], ], c(Name, Other, Cor[ Name,Other ]) )
#  }
#}

#fviz_cluster(clust_kmeans, data = FOO, geom = c("point"))

```

### Module Celltype Enrichments

```{r CelltpyeEnrich, echo = F, warning=F }

Lake <- read.csv( file="/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/lakeCellDf.csv", header = T )
Lake <- Lake[ Lake$adj.pval < 0.05, ]
Lake$Module <- as.character(Lake$Module)
Zhang <- read.csv( file="/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/zhangCellDf.csv", header = T )
Zhang <- Zhang[ Zhang$adj.pval < 0.05, ]
Zhang$Module <- as.character(Zhang$Module)

Zrich <- as.data.frame( matrix(0, length(Genes), length(table(Zhang$category)) ))
row.names(Zrich) <- trans[Genes]
colnames(Zrich) <- names(table(Zhang$category))

Lrich <- as.data.frame( matrix(0, length(Genes), length(table(Lake$category)) ))
row.names(Lrich) <- trans[Genes]
colnames(Lrich) <- names(table(Lake$category))

for( gene in Genes ){
  lyst <- as.character( ZoomNet[ ZoomNet$GeneID == gene, ]$Module )
  foo <- Lake[ as.character(Lake$Module) %in% lyst ==T , ]
  for( Cell in as.character(foo$category) ){
    Lrich[ trans[gene], Cell  ] <- Lrich[ trans[gene], Cell  ] + 1
  }
  foo <- Zhang[ as.character(Zhang$Module) %in% lyst ==T , ]
  for( Cell in as.character(foo$category) ){
    Zrich[ trans[gene], Cell  ] <- Zrich[ trans[gene], Cell  ] + 1
  }
}
save_kable(
   Zrich %>% 
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F),
   file=paste0(tabledir,'/ZhangSingleCell.pdf')
)
save_kable(
   Lrich %>% 
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F),
   file=paste0(tabledir,'/LakeSingleCell.pdf')
)
Zrich %>% 
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F)

Lrich %>% 
    kable( escape = F, row.names = TRUE) %>%
    kable_styling("striped", full_width = F)

PAL <- colorRampPalette(c( "white", "yellow", "red"))(n = 50)

setEPS()
postscript(file = paste0(plotdir,'/ZhangCellTypeHeatMap.eps') )
  heatmap.2(as.matrix(Zrich), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)
dev.off()

setEPS()
postscript(file = paste0(plotdir,'/LakeCellTypeHeatMap.eps') )
  heatmap.2(as.matrix(Lrich), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)
dev.off()

heatmap.2(as.matrix(Zrich), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)

heatmap.2(as.matrix(Lrich), col=PAL, key = FALSE,
          tracecol=NA, cexRow = .9, cexCol=.75)

```

#### Load the Modules and create tissue specific Consensus Groups
```{r GeneConsensus, echo = F, warning=F  }

Consensus_Cluster_A <- c("TCXblue", "IFGyellow", "PHGyellow")
Consensus_Cluster_B <- c("DLPFCblue", "CBEturquoise", "STGblue", "PHGturquoise", "IFGturquoise", "TCXturquoise", "FPturquoise")
Consensus_Cluster_C <- c("IFGbrown", "STGbrown", "DLPFCyellow", "TCXgreen", "FPyellow", "CBEyellow", "PHGbrown" )
Consensus_Cluster_D <- c("DLPFCbrown", "STGyellow", "PHGgreen", "CBEbrown", "TCXyellow", "IFGblue", "FPblue")
Consensus_Cluster_E <- c("FPbrown", "CBEblue", "DLPFCturquoise", "TCXbrown", "STGturquoise", "PHGblue")

Trans <- c(rep( "A", 3), rep( "B", 7), rep( "C", 7), rep( "D", 7), rep( "E", 6))
names(Trans) <- c("TCXblue", "IFGyellow", "PHGyellow", "DLPFCblue", "CBEturquoise", "STGblue", "PHGturquoise", "IFGturquoise", "TCXturquoise", "FPturquoise", "IFGbrown", "STGbrown", "DLPFCyellow", "TCXgreen", "FPyellow", "CBEyellow", "PHGbrown", "DLPFCbrown", "STGyellow", "PHGgreen", "CBEbrown", "TCXyellow", "IFGblue", "FPblue", "FPbrown", "CBEblue", "DLPFCturquoise", "TCXbrown", "STGturquoise", "PHGblue")

Nets$ConsensusCluster <- Trans[Nets$Module]

GeneLvl_ConsensusClust <- list()
for( tissue in names(table( Nets$brainRegion )) ){
  for( clust in names(table( Nets$ConsensusCluster ))){
    eval(parse( text=paste0( 'temp <- as.character(Nets[ as.character(Nets$brainRegion) == "', tissue, '" & as.character(Nets$ConsensusCluster) == "', clust,'", ]$GeneID) ')))
    if( length(temp) > 0 ){
          eval(parse( text=paste0( 'GeneLvl_ConsensusClust$', tissue, '$', clust, ' <- temp[ !duplicated(temp) ]' ) ))

    }
  }
}
names(table( Nets$ConsensusCluster ))

#Load the Tissue Expression 
library(data.table)
MSBB <- fread( file ='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/MSSM_FP_STG_PHG_IFG_netResidualExpression.tsv', sep = '\t', header = T )
row.names(MSBB) <- MSBB$ensembl_gene_id
MSBB <-as.data.frame(MSBB)
MSBB_Cov <- fread( file ='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/MSSM_FP_STG_PHG_IFG_Covariates.tsv', sep = '\t', header = T )

Mayo <- fread( file ='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/MAYO_CBE_TCX_netResidualExpression.tsv', sep = '\t', header = T )
row.names(Mayo) <- Mayo$ensembl_gene_id
Mayo <-as.data.frame(Mayo)

Rosmap <- fread( file ='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/ROSMAP_DLPFC_netResidualExpression.tsv', sep = '\t', header = T )
row.names(Rosmap) <- Rosmap$ensembl_gene_id
Rosmap <-as.data.frame(Rosmap)

MSBB_CoVar <- fread( file ='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/MSBB_SVA_Covariates.tsv', header=T, sep='\t' )
MSBB_CoVar$Tissue <- do.call( rbind, strsplit( MSBB_CoVar$Tissue.Diagnosis, "[.]") )[,1]

row.names( Mayo ) <- Mayo$ensembl_gene_id
row.names( Rosmap ) <- Rosmap$ensembl_gene_id
row.names( MSBB ) <- MSBB$ensembl_gene_id

#Get Tissue Specific Expression Data Frames
DLPFC_exp  <- Rosmap[ , colnames(Rosmap)[colnames(Rosmap) != "ensembl_gene_id"]]
TCX_exp <- Mayo[ , colnames(Mayo)[grepl( 'TCX', colnames(Mayo)) == T]   ]
CBE_exp <- Mayo[ , colnames(Mayo)[grepl( 'CER', colnames(Mayo)) == T]   ]
FP_exp <- MSBB[ ,MSBB_CoVar[ MSBB_CoVar$Tissue == 'FP', ]$SampleID ]
STG_exp <- MSBB[ ,MSBB_CoVar[ MSBB_CoVar$Tissue == 'STG', ]$SampleID ]
IFG_exp <- MSBB[ ,MSBB_CoVar[ MSBB_CoVar$Tissue == 'IFG', ]$SampleID ]
PHG_exp <- MSBB[ ,MSBB_CoVar[ MSBB_CoVar$Tissue == 'PHG', ]$SampleID ]

#Calculate Eigen values
#library(rsvd)
EigenComp <- function( Exp, ENSG ){
  #'@Exp an expression matrix eg. CBE_exp
  #'@ENSG the gene ENSGs in the merged consensus cluster modules for moduleXtissue eg. GeneLvl_ConsensusClust$CBE$A
  #Exp <- CBE_exp
  #ENSG <- GeneLvl_ConsensusClust$CBE$A

  tmp <- row.names(Exp)
  Exp <- as.matrix( as.data.frame(lapply(Exp, as.numeric)) ) 
  row.names(Exp) <- tmp
  colnames(Exp) <- gsub( "X", "", colnames(Exp))

  foo <- Exp[ ENSG, ]
  foo<- as.matrix(foo)
  temp <- svd( dist(t(foo)) )
  eigengenes <- as.vector( svd(dist(t(foo)), nv = 1 )$v )
  names(eigengenes) <- colnames( Exp[ ENSG, ] )
  temp <- cor.test( as.numeric(as.character(Exp[ ENSG, ][1,])), eigengenes)

  Look <- as.data.frame(matrix( NA, 1,3))#CBE_exp

  for( i in row.names( Exp )){
    temp <- cor.test( Exp[ i, ], eigengenes)
    Look <- as.data.frame(rbind( Look[1:dim(Look)[1],], c( i, temp$estimate, temp$p.value) ))
    Look <- na.omit(Look)
  }
  Look <- na.omit(Look)
  Look$fdrCorP <- p.adjust(Look$V3, method = "fdr", n = dim(Look)[1])
  Look$InConsensusMod <- Look$V1 %in% GeneLvl_ConsensusClust$CBE$A
  return(Look)
}

GeneStandModuleByTissue <- list()
```

```{r GeneConsensus_V3, echo = F, warning=F, eval=F  }
##Need to fix
GeneStandModuleByTissue$CBE$A <- EigenComp(CBE_exp,GeneLvl_ConsensusClust$CBE$A)
GeneStandModuleByTissue$CBE$B <- EigenComp(CBE_exp,GeneLvl_ConsensusClust$CBE$B)

GeneStandModuleByTissue$CBE$B <- EigenComp(CBE_exp,GeneLvl_ConsensusClust$CBE$B)
GeneStandModuleByTissue$CBE$B <- EigenComp(CBE_exp,GeneLvl_ConsensusClust$CBE$B)

wilcox.test( -log(Look[ Look$InConsensusMod == T,]$fdrCorP), -log(Look[ Look$InConsensusMod == F,]$fdrCorP))
wilcox.test( -log(Look[ Look$InConsensusMod == T,]$fdrCorP), -log(Look$fdrCorP))
```

### 2020-01-18 code/01-Initializer.r

@@ -1,0 +1,21 @@
+#This script takes a user specified Yaml file and creates a markdown file from it
+#To run: Rscript Run.r <Configuration YAML FILE> <
+
+library(yaml)
+args <- commandArgs(trailingOnly=TRUE)
+setwd( "~/AD_TargetRank/")
+
+#READ IN Config
+#config<-'configs/TestConfig.yaml'
+configuration <- args[1]
+config <- read_yaml(configuration)
+
+#CREATE the RMD to RUN
+system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ~/AD_TargetRank/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))
+
+setwd("~/AD_TargetRank/")
+source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
+source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
+createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/03-Run.Rmd",
+                                  parentId =config$parentID,
+                                  folderName = config$runID)

### 2020-01-18 code/02-Master.Rmd
Preamble Work

@@
 
-```{r yamlIMPT, eval = F, include=FALSE}
+```{r yamlIMPT, eval = T, include=T}
 #Assign your YAML formated config file to a varable
-YAML_Config <- "YAML_FILE"
+configuration <- "YAML_FILE"
+config <- read_yaml(configuration)
 
-setwd("~/AD_TargetRank/")
-source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
-source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
-createAndKnitToFolderEntityClient(file = "CMC_ACC_Exp_DataProcessing_V2.Rmd",
-                                          parentId ="syn21478397",
-                                          folderName = 'ACC')
+setwd(config$filedir)
+plotdir=paste0( 'runs/', config$runname, '/figures' )
+tabledir=paste0( 'runs/', config$runname, '/tables' )
 
-#Test Genes
-Genes <- c('ENSG00000081189','ENSG00000086288', 'ENSG00000137845', 'ENSG00000138600', 'ENSG00000173157', 
-           'ENSG00000134463', 'ENSG00000156170', 'ENSG00000120899', 'ENSG00000147065','ENSG00000130203', 
-           'ENSG00000165025','ENSG00000174628', 'ENSG00000265148', 'ENSG00000159640', 'ENSG00000154734', 
-           'ENSG00000136717', 'ENSG00000095970', 'ENSG00000120885', 'ENSG00000064687')
+Tissues <- config$tissue
 
-#Ana's genes
-Genes <- c("ENSG00000048392", "ENSG00000051620", "ENSG00000070495", "ENSG00000074800", "ENSG00000077152", "ENSG00000091140", "ENSG00000096060", "ENSG00000100030", "ENSG00000100418", "ENSG00000100605", "ENSG00000105220", "ENSG00000105398", "ENSG00000108424", "ENSG00000109576", "ENSG00000111640", "ENSG00000113657", "ENSG00000117115", "ENSG00000117228", "ENSG00000119125", "ENSG00000121879", "ENSG00000128683", "ENSG00000130540", "ENSG00000136634", "ENSG00000137992", "ENSG00000139112", "ENSG00000140374", "ENSG00000143546", "ENSG00000146701", "ENSG00000146733", "ENSG00000155368", "ENSG00000156709", "ENSG00000162433", "ENSG00000164111", "ENSG00000165025", "ENSG00000166224", "ENSG00000167325", "ENSG00000167720", "ENSG00000168291", "ENSG00000176490", "ENSG00000177511", "ENSG00000186298", "ENSG00000196154", "ENSG00000197448", "ENSG00000197956", "ENSG00000213719")
-
-#Genes <- c('ENSG00000147065','ENSG00000130203')
-#Genes <- c('ENSG00000130203')
-Tissues <- c('CBE',  'DLPFC' , 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 #DE  = 'syn14237651'
@@
 ```
-
-```{r Settings, echo =F, warning=F}
-plotdir='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/Plots/'
-tabledir='Tables/'
-Tissues <- c('CBE', 'DLPFC', 'IFG', 'PHG', 'TCX', 'STG', 'FP')
+```{r Module1, eval = T, include=T}
 

### 2020-01-18 code/01-Initializer.r

@@ -1,14 +1,17 @@
 #This script takes a user specified Yaml file and creates a markdown file from it
-#To run: Rscript Run.r <Configuration YAML FILE> <
+#To run: Rscript Run.r <Configuration YAML FILE>
 
 library(yaml)
 args <- commandArgs(trailingOnly=TRUE)
-setwd( "~/AD_TargetRank/")
 
 #READ IN Config
-#config<-'configs/TestConfig.yaml'
+#configuration<-'configs/TestConfig.yaml'
 configuration <- args[1]
 config <- read_yaml(configuration)
+
+setwd( config$filedir )
+system( paste0('mkdir -p runs/', config$runname, '/figures' ) )
+system( paste0('mkdir -p runs/', config$runname, '/tables' ) )
 
 #CREATE the RMD to RUN
 system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ~/AD_TargetRank/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))

### 2020-01-20 code/02-Master.Rmd
Gene Name Munger

@@
 configuration <- "YAML_FILE"
-config <- read_yaml(configuration)
+config <- read_yaml(file=paste0('./',configuration))
 
@@
 ```
-```{r Module1, eval = T, include=T}
+
+Importing genelist `r thisFile` as `r config$genelisttype`
+```{r GeneNames, eval=T, include=T}
+
+GeneData <- function( id, type ){
+  #Retreives gene name info
+  #'@id the gene ids to source a list from
+  #'@type either ENSG or GNAME for type of gen ids
+
+  org = 'hsa'
+  host = 'grch37.ensembl.org'
+  ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
+  ds <- listDatasets(ensembl)[, "dataset"]
+  ds <- grep(paste0("^", org), ds, value = TRUE)
+
+  if (length(ds) == 0){
+    stop(paste("Mart not found for:", org))
+  } else if (length(ds) > 1) {
+    message("Found several marts")
+    sapply(ds, function(d) message(paste(which(ds == d), d, sep = ": ")))
+    n <- readline(paste0("Choose mart (1-", length(ds), ") : "))
+    ds <- ds[as.integer(n)]
+  }
+
+  if(type == 'ENSG'){
+    id.type = 'ensembl_gene_id'
+    ensembl <- useDataset(ds, mart = ensembl)
+    Genes <- getBM(filters = id.type, attributes = c(id.type, 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
+  }else{
+    if(type == 'GNAME'){
+      id.type = 'hgnc_symbol'
+      ensembl <- useDataset(ds, mart = ensembl)
+      Genes <- getBM(filters = id.type, attributes = c('ensembl_gene_id', id.type, 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
+      ##Cut Decmals
+    }else{
+      #Error Gene names not specififed correctly
+      stop("ERROR: SOURCE=Config.yaml Issue=genelisttype must be ENSG or GNAME")
+    }
+  }
+  return(Genes)
+}
+
+#Load Gene Set:
+Genes <- read.table( paste0(config$filedir, '/genelists/', config$genelistfile ), header=F, sep='\t')
+
+Tab <- GeneData( Genes, config$genelisttype )
+Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
+Tab$Interval_hg19 <- NA
+for( i in 1:dim(Tab)[1]){
+  if(  Tab$start_position[i] < Tab$ensembl_gene_id[i] ){
+    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])-1e6, "-", as.numeric(Tab$end_position[i])+1e6)
+  }else{
+    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])+1e6, "-", as.numeric(Tab$end_position[i])-1e6)
+  }
+}
 
 ```
+
+```{r Module1, eval = T, include=T}
+if( i %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence')
+
+  source()
+
+}else{
+  writeLines('Skipping IGAP based Model and Gene-Ranking: User specified module 1 not included in <Config>.yaml ')
+}
+
+```

### 2020-01-20 code/03-Run.Rmd
Arange preamble
Gene Name Munger

--- 2020-01-18 code/02-Master.Rmd
@@
 
-```{r yamlIMPT, eval = F, include=FALSE}
+```{r yamlIMPT, eval = T, include=T}
 #Assign your YAML formated config file to a varable
-YAML_Config <- "YAML_FILE"
+configuration <- "configs/TestConfig.yaml"
+config <- read_yaml(configuration)
 
-setwd("~/AD_TargetRank/")
-source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
-source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
-createAndKnitToFolderEntityClient(file = "CMC_ACC_Exp_DataProcessing_V2.Rmd",
-                                          parentId ="syn21478397",
-                                          folderName = 'ACC')
+setwd(config$filedir)
+plotdir=paste0( 'runs/', config$runname, '/figures' )
+tabledir=paste0( 'runs/', config$runname, '/tables' )
 
-#Test Genes
-Genes <- c('ENSG00000081189','ENSG00000086288', 'ENSG00000137845', 'ENSG00000138600', 'ENSG00000173157', 
-           'ENSG00000134463', 'ENSG00000156170', 'ENSG00000120899', 'ENSG00000147065','ENSG00000130203', 
-           'ENSG00000165025','ENSG00000174628', 'ENSG00000265148', 'ENSG00000159640', 'ENSG00000154734', 
-           'ENSG00000136717', 'ENSG00000095970', 'ENSG00000120885', 'ENSG00000064687')
+Tissues <- config$tissue
 
-#Ana's genes
-Genes <- c("ENSG00000048392", "ENSG00000051620", "ENSG00000070495", "ENSG00000074800", "ENSG00000077152", "ENSG00000091140", "ENSG00000096060", "ENSG00000100030", "ENSG00000100418", "ENSG00000100605", "ENSG00000105220", "ENSG00000105398", "ENSG00000108424", "ENSG00000109576", "ENSG00000111640", "ENSG00000113657", "ENSG00000117115", "ENSG00000117228", "ENSG00000119125", "ENSG00000121879", "ENSG00000128683", "ENSG00000130540", "ENSG00000136634", "ENSG00000137992", "ENSG00000139112", "ENSG00000140374", "ENSG00000143546", "ENSG00000146701", "ENSG00000146733", "ENSG00000155368", "ENSG00000156709", "ENSG00000162433", "ENSG00000164111", "ENSG00000165025", "ENSG00000166224", "ENSG00000167325", "ENSG00000167720", "ENSG00000168291", "ENSG00000176490", "ENSG00000177511", "ENSG00000186298", "ENSG00000196154", "ENSG00000197448", "ENSG00000197956", "ENSG00000213719")
-
-#Genes <- c('ENSG00000147065','ENSG00000130203')
-#Genes <- c('ENSG00000130203')
-Tissues <- c('CBE',  'DLPFC' , 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 #DE  = 'syn14237651'
@@
 
-```{r Settings, echo =F, warning=F}
-plotdir='/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/Plots/'
-tabledir='Tables/'
-Tissues <- c('CBE', 'DLPFC', 'IFG', 'PHG', 'TCX', 'STG', 'FP')
+```{r Process, eval=T, include=T}
+
+  writeLines('Importing genelist ', config$genelistfile, ' as ', config$Genelisttype)
 
 ```
+
+```{r GeneNames, eval=T, include=T}
+GeneData <- function( id ){
+
+  org = 'hsa'
+  id.type = 'ensembl_gene_id'
+  #host = 'jul2019.archive.ensembl.org'
+  host = 'grch37.ensembl.org'
+  ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
+  ds <- listDatasets(ensembl)[, "dataset"]
+  ds <- grep(paste0("^", org), ds, value = TRUE)
+
+  if (length(ds) == 0){
+    stop(paste("Mart not found for:", org))
+  } else if (length(ds) > 1) {
+    message("Found several marts")
+    sapply(ds, function(d) message(paste(which(ds == d), d, sep = ": ")))
+    n <- readline(paste0("Choose mart (1-", length(ds), ") : "))
+    ds <- ds[as.integer(n)]
+  }
+
+  ensembl <- useDataset(ds, mart = ensembl)
+
+  Genes <- getBM(filters = id.type, attributes = c(id.type, 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
+  return(Genes)
+}
+
+Tab <- GeneData( Genes )
+Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
+Tab$Interval_hg19 <- NA
+for( i in 1:dim(Tab)[1]){
+  if(  Tab$start_position[i] < Tab$ensembl_gene_id[i] ){
+    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])-1e6, "-", as.numeric(Tab$end_position[i])+1e6)
+  }else{
+    Tab$Interval_hg19[i] <- paste0("chr", Tab$chromosome_name[i], ":", as.numeric(Tab$start_position[i])+1e6, "-", as.numeric(Tab$end_position[i])-1e6)
+  }
+}
+
+```
+```{r Module1, eval = T, include=T}
+if( i %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence')
+
+  source()
+
+}else{
+  writeLines('Skipping IGAP based Model and Gene-Ranking: User specified module 1 not included in <Config>.yaml ')
+}
+
+```

### 2020-01-20 code/modules/01-IGAP_module.R

@@ -1,0 +1,3 @@
+#Load IGAP files
+IGAP <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1_2_combined.txt", sep = "\t", header = T)
+IGAP1 <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1.txt", sep = "\t", header = T)

### 2020-01-21 code/02-Master.Rmd
update

@@
 knitr::opts_chunk$set(echo = F)
+opts_knit$set(root.dir = '../')
 library(readr)
@@
 library(yaml)
+#webshot::install_phantomjs()
 ```
@@
 Tab$Interval_hg19 <- NA
+
 for( i in 1:dim(Tab)[1]){
@@
 }
-
 ```
@@
 ```{r Module1, eval = T, include=T}
-if( i %in% config$modules ){
+if( 1 %in% config$modules ){
   writeLines('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence')
 
-  source()
+  source('code/modules/01-IGAP_module.R')
 
@@
 ```
+
+```{r Module2, eval = T, include=T}
+if( 2 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence')
+
+  source('modules/02-DiffExp_module.R')
+
+}else{
+  writeLines('Skipping Differential Expression based Model and Gene-Ranking: User specified module 2 not included in <Config>.yaml ')
+}
+
+```

### 2020-01-21 code/03-Run.Rmd
update

@@
 configuration <- "configs/TestConfig.yaml"
-config <- read_yaml(configuration)
+config <- read_yaml(file=paste0(configuration))
 
@@
 
-```{r Process, eval=T, include=T}
+Importing genelist `r thisFile` as `r config$genelisttype`
+```{r GeneNames, eval=T, include=T}
 
-  writeLines('Importing genelist ', config$genelistfile, ' as ', config$Genelisttype)
-
-```
-
-```{r GeneNames, eval=T, include=T}
-GeneData <- function( id ){
+GeneData <- function( id, type ){
+  #Retreives gene name info
+  #'@id the gene ids to source a list from
+  #'@type either ENSG or GNAME for type of gen ids
 
   org = 'hsa'
-  id.type = 'ensembl_gene_id'
-  #host = 'jul2019.archive.ensembl.org'
   host = 'grch37.ensembl.org'
@@
 
+  if(type == 'ENSG'){
+    id.type = 'ensembl_gene_id'
     ensembl <- useDataset(ds, mart = ensembl)
-
     Genes <- getBM(filters = id.type, attributes = c(id.type, 'hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
+  }else{
+    if(type == 'GNAME'){
+      id.type = 'hgnc_symbol'
+      ensembl <- useDataset(ds, mart = ensembl)
+      Genes <- getBM(filters = id.type, attributes = c('ensembl_gene_id', id.type, 'chromosome_name', 'start_position', 'end_position'), values = id, mart = ensembl)
+      ##Cut Decmals
+    }else{
+      #Error Gene names not specififed correctly
+      stop("ERROR: SOURCE=Config.yaml Issue=genelisttype must be ENSG or GNAME")
+    }
+  }
   return(Genes)
@@
 
-Tab <- GeneData( Genes )
+#Load Gene Set:
+Genes <- read.table( paste0(config$filedir, '/genelists/', config$genelistfile ), header=F, sep='\t')
+
+Tab <- GeneData( Genes, config$genelisttype )
 Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
@@
 }
+```
 
-```
 ```{r Module1, eval = T, include=T}
-if( i %in% config$modules ){
+if( 1 %in% config$modules ){
   writeLines('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence')
 
-  source()
+  source('code/modules/01-IGAP_module.R')
 

### 2020-01-21 code/01-Initializer.r

@@ -16,6 +16,11 @@ system( paste0('mkdir -p runs/', config$runname, '/tables' ) )
 #CREATE the RMD to RUN
 system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ~/AD_TargetRank/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))
 
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
 setwd("~/AD_TargetRank/")
 source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
 source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")

### 2020-01-21 code/modules/01-IGAP_module.R

@@ -1,3 +1,203 @@
 #Load IGAP files
-IGAP <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1_2_combined.txt", sep = "\t", header = T)
-IGAP1 <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1.txt", sep = "\t", header = T)
+#IGAP <- read.table( synfile = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1_2_combined.txt", sep = "\t", header = T)
+#IGAP1 <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1.txt", sep = "\t", header = T)
+IGAP <- read.table( syn_temp$get('syn21534587')$path, sep = "\t", header = T)
+IGAP1 <- read.table( syn_temp$get('syn21534586')$path, sep = "\t", header = T)
+
+Tab$LowestPValCombined <- 1
+Tab$TotalSNPsCombined <- 0
+
+Tab$LowestPValPhase1 <- 1
+Tab$TotalSNPsPhase1 <- 0
+
+#New Dataframe for the IGAP results that you can track the pval distributions
+TotalIGAP <- NULL
+
+#Look at the dataframe 
+MedianScoreIGAP <- NULL
+
+#Look at the dataframe 
+BestScoreIGAP <- NULL
+
+for( i in 1:dim(Tab)[1]){
+  
+  if( isTRUE(as.character(Tab$chromosome_name)[i] %in% as.character(c(1:22))) ){
+    Temp <- IGAP[ (as.numeric(as.character(IGAP$Position)) > (as.numeric(as.character(Tab$start_position))[i]-1e6)) &
+                    (as.numeric(as.character(IGAP$Position)) < (as.numeric(as.character(Tab$start_position))[i]+1e6)) &
+                    as.numeric(as.character(IGAP$Chromosome)) == as.numeric(as.character(Tab$chromosome_name)[i])  , ]
+    if( dim(Temp)[1] == 0){
+      Temp[1,] <- 0
[…]

### 2020-01-21 code/test.r

@@ -1,0 +1,1 @@
+print(head(Tab))

### 2020-01-23 code/02-Master.Rmd
pre Mod7 final

@@
   pdf_document: default
+editor_options: 
+  chunk_output_type: console
 ---
@@
 knitr::opts_chunk$set(echo = F)
-opts_knit$set(root.dir = '../')
 library(readr)
@@
 library(yaml)
+knitr::opts_knit$set(root.dir = config$filedir )
+print(getwd())
 #webshot::install_phantomjs()
@@
 ```{r yamlIMPT, eval = T, include=T}
+print(getwd())
 #Assign your YAML formated config file to a varable
@@
 Tissues <- config$tissue
+
+#Create Synapse Run Folder and retrieve parentID
+CODE <- syn_temp$store(synapseclient$Folder(name = config$runID, parentId = config$parentID))
+
+foo <- syn_temp$getChildren('syn21532475')
+Temp <- reticulate::iterate(foo, simplify = F)
+for( i in 1:length(Temp) ){
+  if( Temp[[i]]$name == config$runID ){
+    RunParent <- Temp[[i]]$id
+  }else{
+  }
+}
 
@@
 
-Importing genelist `r thisFile` as `r config$genelisttype`
+Importing genelist `r config$genelistfile` as `r config$genelisttype`
 ```{r GeneNames, eval=T, include=T}
@@
   org = 'hsa'
-  host = 'grch37.ensembl.org'
+  host = 'grch38.ensembl.org'
   ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
@@
 Tab <- GeneData( Genes, config$genelisttype )
+#Trans used as raw for module 7
+Trans <- Tab
 Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
@@
 }else{
-  writeLines('Skipping IGAP based Model and Gene-Ranking: User specified module 1 not included in <Config>.yaml ')
+  writeLines( paste0('Skipping IGAP based Model and Gene-Ranking: User specified module 1 not included in ', config$name) )
 }
@@
 
-  source('modules/02-DiffExp_module.R')
+  source('code/modules/02-DiffExp_module.R')
 
 }else{
-  writeLines('Skipping Differential Expression based Model and Gene-Ranking: User specified module 2 not included in <Config>.yaml ')
+  writeLines( paste0('Skipping Differential Expression based Model and Gene-Ranking: User specified module 2 not included in ', config$name) )
 }
@@
 ```
+
+```{r Module3, eval = T, include=T}
+if( 3 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Proteomic Expression evidence')
+
+  source('code/modules/03-DiffProt_module.R')
+
+}else{
+  writeLines( paste0('Skipping Differential Proteomic Expression based Model and Gene-Ranking: User specified module 3 not included in ', config$name) )
+}
+
+```
+
+```{r Module4, eval = T, include=T}
+if( 4 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by eQTL evidence')
+
+  source('code/modules/04-eQTL_module.R')
+
+}else{
+  writeLines( paste0('Skipping eQTL based Model and Gene-Ranking: User specified module 4 not included in ', config$name) )
+}
+
+```
+
+```{r Module5, eval = T, include=T}
+if( 5 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Coexpression Module evidence')
+
+  source('code/modules/05-CoExp_module.R')
+
+}else{
+  writeLines( paste0('Skipping Coexpression Module based Model and Gene-Clustering: User specified module 5 not included in ', config$name) )
+}
+
+```
+
+```{r Module6, eval = T, include=T}
+if( 6 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Module Cell-Type Enrichments evidence')
+
+  source('code/modules/06-CellType_module.R')
+
+}else{
+  writeLines( paste0('Skipping Coexpression Module based Model and Gene-Clustering: User specified module 6 not included in ', config$name) )
+}
+
+```
+
+```{r Module7, eval = T, include=T}
+setwd( config$filedir )
+if( 7 %in% config$modules ){
+  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence')
+
+  source('code/modules/07-Coexpression.R')
+
+}else{
+  writeLines( paste0('Skipping Partial Coexpression Analysis: User specified module 7 not included in ', config$name) )
+}
+
+```

### 2020-01-23 code/03-Run.Rmd
pre Mod7 final

@@
 ```{r yamlIMPT, eval = T, include=T}
+print(getwd())
 #Assign your YAML formated config file to a varable
 configuration <- "configs/TestConfig.yaml"
-config <- read_yaml(file=paste0(configuration))
+config <- read_yaml(file=paste0('./',configuration))
 
@@
 Tab <- GeneData( Genes, config$genelisttype )
+#Trans used as raw for module 7
+Trans <- Tab
 Tab$Coord_hg19 <- paste0("chr", Tab$chromosome_name, ":", Tab$start_position, "-",Tab$end_position)
 Tab$Interval_hg19 <- NA
+
 for( i in 1:dim(Tab)[1]){

### 2020-01-23 code/modules/01-IGAP_module.R

@@ -38,7 +38,8 @@ for( i in 1:dim(Tab)[1]){
                        as.numeric(as.character(IGAP1$Chromosome)) == as.numeric(as.character(Tab$chromosome_name)[i])  , ] 
     if( dim(TempV1)[1]>0 ){
       TempV1$GeneName <- Tab$hgnc_symbol[i]
-      #TotalIGAP <- as.data.frame(rbind(TotalIGAP,TempV1))
+      #####
+      TotalIGAP <- as.data.frame(rbind(TotalIGAP,TempV1))
     }else{}
     
     if( median(c(1:length(TempV1$Pvalue))) == floor(median(c(1:length(TempV1$Pvalue)))) & median(c(1:length(TempV1$Pvalue))) == ceiling(median(c(1:length(TempV1$Pvalue)))) ) {
@@ -78,7 +79,7 @@ for( i in 1:dim(Tab)[1]){
     BestScoreIGAP <- as.data.frame(rbind(BestScoreIGAP,Temp[ min(Temp$Pvalue) == as.numeric(Temp$Pvalue), ]))
   }
 }
-colnames(Tab) <- c("ENSG", "GeneName", "Start", "End", "Chr", "Coordhg19", "Intervalhg19", "LowestPValCombined", "TotalSNPsCombined", "LowestPValPhaseI", "TotalSNPsPhaseI" )
+colnames(Tab) <- c("ENSG", "GeneName", "Chr", "Start", "End", "Coordhg19", "Intervalhg19", "LowestPValCombined", "TotalSNPsCombined", "LowestPValPhaseI", "TotalSNPsPhaseI" )
 row.names(Tab) <- Tab$ENSG
 
 TotalIGAP$LogP <- -log(TotalIGAP$Pvalue)
@@ -99,6 +100,9 @@ ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) +
   theme(axis.text.x = element_text(angle = 45) ) + ylab('-Log(P-Value)') 
 dev.off()
 
+#Replace Infinite values with max + one
+FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)
+
 IGAP$Rank<-rank(-IGAP$Pvalue)
 FOO<-IGAP
 FOO$Pvalue <- -log(FOO$Pvalue)
[…]

### 2020-01-23 code/modules/07-Coexpression.R

@@ -1,0 +1,148 @@
+##Coexpression Analysis module:
+#install.packages("~/Desktop/Programs/spike/spike/", repos = NULL, type = "source")
+
+#config$tissue <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
+
+# Github link
+thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
+
+setwd( config$filedir )
+
+GL_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$genelistfile)
+CF_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$name)
+I_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/01-Initializer.r')
+M_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/02-Master.Rmd')
+
+CODE <- syn_temp$store(synapseclient$Folder(name = 'figures', parentId = RunParent))
+
+#SynIDs of expression data to pull from
+ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
+names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
+
+Study <- c( 'RosMap', 'Mayo', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
+names(Study) <- c('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
+
+row.names(Trans) <- Trans$ensembl_gene_id
+Trans <- Trans[ Trans$chromosome_name %in% c(1:22,"X","Y"), ]
+
+writeLines(paste0('For each tissue specified in ', 
+                  paste( config$name, collapse = ', '), 
+                  " the coexpression of genes specified in ", 
+                  config$genelistfile, 
+                  " is given. This is measured as a percentage of 1000 pemutations in which the gene was a significant predictor of expression of its comparison gene by a linear model."
+                  )
+           )
+#List of Genes Not Expressed In Tissue
+#dev.off()
+Missing <- list()
+Plots <- list()
+
+for( Tissue in config$tissue ){
+  
[…]

### 2020-01-24 code/02-Master.Rmd
Mod 7 Plots

@@
   org = 'hsa'
-  host = 'grch38.ensembl.org'
-  ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
+  #host = 'grch38.ensembl.org'
+  #"ensembl",dataset="hsapiens_gene_ensembl")
+  #ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = host)
+  ensembl <- useMart("ensembl",dataset="hsapiens_gene_ensembl")
   ds <- listDatasets(ensembl)[, "dataset"]
@@
 #Load Gene Set:
-Genes <- read.table( paste0(config$filedir, '/genelists/', config$genelistfile ), header=F, sep='\t')
+Genes <- read.table( config$genelistfile , header=F, sep='\t')
 
@@
 if( 1 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence')
+  writeLines(paste0('Modeling and ranking genelist ', config$genelistfile ,' by IGAP SNP evidence'))
 
@@
 if( 2 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence')
+  writeLines(paste0( 'Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence'))
 
@@
 if( 3 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Proteomic Expression evidence')
+  writeLines(paste0('Modeling and ranking genelist ', config$genelistfile ,' by Differential Proteomic Expression evidence'))
 
@@
 if( 4 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by eQTL evidence')
+  writeLines(paste0( 'Modeling and ranking genelist ', config$genelistfile ,' by eQTL evidence') )
 
@@
 if( 5 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Coexpression Module evidence')
+  writeLines(paste0( 'Modeling and ranking genelist ', config$genelistfile ,' by Coexpression Module evidence') )
 
@@
 if( 6 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Module Cell-Type Enrichments evidence')
+  writeLines(paste0( 'Modeling and ranking genelist ', config$genelistfile ,' by Module Cell-Type Enrichments evidence'))
 
@@
 if( 7 %in% config$modules ){
-  writeLines('Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence')
+  writeLines(paste0('Modeling and ranking genelist ', config$genelistfile ,' by Differential Expression evidence'))
 

### 2020-01-24 code/01-Initializer.r

@@ -1,5 +1,11 @@
 #This script takes a user specified Yaml file and creates a markdown file from it
-#To run: Rscript Run.r <Configuration YAML FILE>
+#To run: Rscript code/01-Initializer.r <Configuration YAML FILE>
+#eg. Rscript code/01-Initializer.r configs/TestConfig.yaml
+#eg. Rscript code/01-Initializer.r configs/M62.yaml
+#eg. Rscript code/01-Initializer.r configs/pQTL.yaml
+#eg. Rscript code/01-Initializer.r configs/Retromer.yaml
+#eg. Rscript code/01-Initializer.r configs/Moesin.yaml
+#eg. Rscript code/01-Initializer.r configs/EmoryTargets.yaml
 
 library(yaml)
 args <- commandArgs(trailingOnly=TRUE)

### 2020-01-24 code/modules/07-Coexpression.R

@@ -1,13 +1,12 @@
 ##Coexpression Analysis module:
 #install.packages("~/Desktop/Programs/spike/spike/", repos = NULL, type = "source")
 
-#config$tissue <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
-
-# Github link
-thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
-
+# Load Libraries and se working directory 
+library(pheatmap)
 setwd( config$filedir )
 
+# Set Github Provenance links
+thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
 GL_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$genelistfile)
 CF_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$name)
 I_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/01-Initializer.r')
@@ -19,9 +18,11 @@ CODE <- syn_temp$store(synapseclient$Folder(name = 'figures', parentId = RunPare
 ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
 names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 
+#Study ID Translator
 Study <- c( 'RosMap', 'Mayo', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
 names(Study) <- c('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
 
+#Useable Chrs for genelist ENSG  translations and wiki analysis preamble
 row.names(Trans) <- Trans$ensembl_gene_id
 Trans <- Trans[ Trans$chromosome_name %in% c(1:22,"X","Y"), ]
 
@@ -32,14 +33,13 @@ writeLines(paste0('For each tissue specified in ',
                   " is given. This is measured as a percentage of 1000 pemutations in which the gene was a significant predictor of expression of its comparison gene by a linear model."
                   )
            )
-#List of Genes Not Expressed In Tissue
-#dev.off()
+# Make objects for list of Genes Not Expressed In Tissue and dataframes to plot
 Missing <- list()
 Plots <- list()
 
 for( Tissue in config$tissue ){
   
-  # Set annotations
[…]

### 2020-01-24 code/modules/07-Coexpression.R
Add Parallel vbrbootstrap

@@ -3,6 +3,9 @@
 
 # Load Libraries and se working directory 
 library(pheatmap)
+library(parallel)
+library(doParallel)
+
 setwd( config$filedir )
 
 # Set Github Provenance links
@@ -90,8 +93,8 @@ for( Tissue in config$tissue ){
   row.names(Exp) <- row.names(t(exp)) 
   colnames(Exp) <- colnames(t(exp)) 
   
-  x <- Exp[ , colnames(Exp) %in% row.names(Trans)]
-  
+  #x <- Exp[ , colnames(Exp) %in% row.names(Trans)]
+  x <- Exp
   #Record Genes missing in tissue of interest
   if( as.numeric(table(row.names(Trans) %in% colnames(x) )["TRUE"] ) == dim(Trans)[1] ){
     Missing[[ Tissue ]] <- paste0("All Listed Genes Expressed in ", Tissue)
@@ -101,24 +104,49 @@ for( Tissue in config$tissue ){
                                  )
   }
   #Build output matrix for partial correlation
-  if(length(Missing))
-  Final <- matrix( NA, length(colnames(x)), length(colnames(x)) )
-  row.names(Final) <- colnames(x)
-  colnames(Final) <- colnames(x)
   
+  #Pairwise spearman correlation of genes
+  XCor <- cor(x, method = "spearman") 
+  
+  #Prep for partial correlation detection
+  Final <- data.frame()
+  core <- parallel::detectCores()-2 
+  source("~/AD_TargetRank/utilityFunctions/Parallel_vbsrBootstrap.R")
   #Run Partial correlations for each gene
-  for( i in colnames(x) ){
+  mark <- Sys.time()
+  for( i in row.names(Trans) ){
     OBS <- i
     y <- as.matrix(x[,OBS]) 
     colnames(y) <- i
     X <- x[,(colnames(x) %in% OBS) == F ]
-    att <- spike::vbsrBootstrap( y=y,x=X,nsamp=1000,cores=4)
     
-    eval(parse(text = paste0("Final[ i,] <- c( att[ (names(att) %in% 'intercept')==F ], ", i, " = 1)[colnames(Final)]") ))
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=100, cores=core )
+    
+    #Build Iteration based DF
+    temp <- as.data.frame( cbind( Target_Gene = rep(i,length(att)), 
+                Seed_Set = rep(config$runID,length(att)),
[…]

### 2020-01-24 utilityFunctions/Parallel_vbsrBootstrap.R

@@ -1,0 +1,44 @@
+pvbsrBootstrap = function(y,x,nsamp=100,cores=8){
+  library(dplyr)
+  n <- length(y)
+  replicateMatrix = sample(1:(n*nsamp),replace=TRUE) %>%
+    matrix(n,nsamp)
+  replicateMatrix = replicateMatrix%%n +1
+  #print(replicateMatrix[1:5,])
+  fxn1 <- function(shuf,y,x){
+    #library(utilityFunctions)
+    #return(utilityFunctions::fastlmbeta(y[shuf],x[shuf,]))
+    #library(vbsr)
+    #y=y[shuf]
+    #x=x[shuf,]
+    res <- vbsr::vbsr(y[shuf],x[shuf,])
+    
+    ###identify significant features
+    ###return unpenalized betas
+    baz = rep(0,ncol(x)+1)
+    names(baz) = c('intercept',colnames(x))
+    whichSig = which(res$pval < 0.05/ncol(x))
+    if(length(whichSig)>0){
+      #write a fastlmbeta function that also returns correlation
+      baz2 = c()
+      try(baz2 <- spike::fastlmbeta2(y[shuf],x[shuf,whichSig],colnames(x)[whichSig]),silent=T)
+      if(length(baz2)>0){
+        baz[names(baz2)] <- baz2
+      }
+    }
+    
+    #return(c(res$alpha,res$beta))
+    return(baz)
+  }
+  
+  cl <- parallel::makeCluster(cores)
+  registerDoParallel(cl)
+  #betaMatrix <- foreach(i=1:nsamp,.combine='rbind') %dopar% fxn1(replicateMatrix[,i],y,data.matrix(x))
+  betaMatrix <- t(parallel::parApply(cl,replicateMatrix,2,fxn1,y,data.matrix(x)))
+  #betaMatrix <- t(apply(replicateMatrix,2,fxn1,y,data.matrix(x)))
+  parallel::stopCluster(cl)
+  #colnames(betaMatrix) <- c(colnames(x))
+  return(apply(betaMatrix!=0,2,mean))
+  #return(betaMatrix)
+  
+}

### 2020-01-24 code/01-Initializer.r
WIP

@@ -11,7 +11,7 @@ library(yaml)
 args <- commandArgs(trailingOnly=TRUE)
 
 #READ IN Config
-#configuration<-'configs/TestConfig.yaml'
+#configuration<-'configs/M62.yaml'
 configuration <- args[1]
 config <- read_yaml(configuration)
 

### 2020-01-24 code/modules/07-Coexpression.R

@@ -1,6 +1,6 @@
 ##Coexpression Analysis module:
 #install.packages("~/Desktop/Programs/spike/spike/", repos = NULL, type = "source")
-
+source("~/AD_TargetRank/utilityFunctions/Parallel_vbsrBootstrap.R")
 # Load Libraries and se working directory 
 library(pheatmap)
 library(parallel)
@@ -25,6 +25,8 @@ names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 Study <- c( 'RosMap', 'Mayo', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
 names(Study) <- c('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
 
+Syn <- list('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
+
 #Useable Chrs for genelist ENSG  translations and wiki analysis preamble
 row.names(Trans) <- Trans$ensembl_gene_id
 Trans <- Trans[ Trans$chromosome_name %in% c(1:22,"X","Y"), ]
@@ -40,28 +42,12 @@ writeLines(paste0('For each tissue specified in ',
 Missing <- list()
 Plots <- list()
 
+FinalTis <- data.frame()
+
 for( Tissue in config$tissue ){
-  
-  # Set annotations for synapse objects ( figures and tables )
-  all.annotations = list(
-    dataType = 'Coexpression',
-    dataSubType = 'pheatmap PDF',
-    summaryLevel = 'gene',
-    assay	 = 'RNAseq',
-    tissueTypeAbrv	= Tissue, 
-    study = Study[Tissue], 
-    organism = 'HomoSapiens',
-    consortium	= 'AMPAD',
-    normalizationStatus	= TRUE,
-    normalizationType	= 'CQN',
-    rnaquantification = 'RSEM',
-    genomeAssemblyID = 'GRCh38'
-  )
   
   Syns_Used <- NULL
   #Tissue <- 'STG'
-  activityName = paste0( Tissue, 'Coexpression heatmap')
-  activityDescription = paste0( Tissue, 'Coexpression heatmap for genelist: ', config$genelistfile);
   
   #Load expression for tissue
[…]

### 2020-01-25 code/prepscripts/07_1-SpearmanCor.R
SpearCor

@@ -1,0 +1,47 @@
+library(parallel)
+library(doParallel)
+
+ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
+names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
+
+for( Tissue in c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX') ){
+  
+  thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
+  thisFile  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/prepscripts/07_1-SpearmanCor.R')
+  
+  LIST <- colnames(x)
+  LIST <- combn(LIST,2)
+  
+  OUT_Cor <- as.data.frame( matrix(0, length(colnames(x)), length(colnames(x))) )
+  colnames(OUT_Cor) <- colnames(x)
+  row.names(OUT_Cor) <- colnames(x)
+  
+  OUT_pVal <- as.data.frame( matrix(0, length(colnames(x)), length(colnames(x))) )
+  colnames(OUT_pVal) <- colnames(x)
+  row.names(OUT_pVal) <- colnames(x)
+  
+  library(doParallel)
+  cl <- makeCluster(detectCores()-1)
+  registerDoParallel(cl)
+  
+  mark <- Sys.time()
+  foreach(i=c(1:dim(LIST)[2])) %do% {
+    
+    mark <- Sys.time()
+    foo<-LIST[1,i]
+    foo2<-LIST[2,i]
+    test <- cor.test( x[,foo], x[,foo2], method = "spearman" )
+    
+    OUT_Cor[ foo,foo2 ] <- test$estimate
+    OUT_Cor[ foo2,foo ] <- test$estimate
+    
+    #pVal Test
+    OUT_pVal[ foo,foo2 ] <- test$p.value
+    OUT_pVal[ foo2,foo ] <- test$p.value
+    Sys.time()-mark
+  }
+  Sys.time()-mark
+  
+  stopCluster(cl)
+  
+}

### 2020-01-25 code/03-Run.Rmd
Initial Mod7

@@
 #Assign your YAML formated config file to a varable
-configuration <- "configs/TestConfig.yaml"
+configuration <- "configs/EmoryTargets.yaml"
 config <- read_yaml(file=paste0('./',configuration))

### 2020-01-25 code/modules/07-Coexpression.R

@@ -74,6 +74,8 @@ for( Tissue in config$tissue ){
   }
   
   #Impute svalues for given gene-patient NA values
+  exp <- exp[rowSums(is.na(exp))<ncol(exp), ]
+  
   foo <- bcv::impute.svd( t(exp) )
   Exp <- foo$x
   row.names(Exp) <- row.names(t(exp)) 
@@ -94,26 +96,9 @@ for( Tissue in config$tissue ){
   #Pairwise spearman correlation of genes
   XCor <- cor(x, method = "spearman")
   
-  mark<-Sys.time()
-  FOO <- psych::corr.test(x, method = "spearman")
-  Sys.time()-mark
-  
-  cor.test.p <- function(x){
-    FUN <- function(x, y) cor.test(x, y, method = "spearman")[["p.value"]]
-    z <- outer(
-      colnames(x), 
-      colnames(x), 
-      Vectorize(function(i,j) FUN(x[,i], x[,j]))
-    )
-    dimnames(z) <- list(colnames(x), colnames(x))
-    z
-  }
-  
-  mark<-Sys.time()
-  FOO <- cor.test.p(x)
-  Sys.time()-mark
-  
-  
+  #mark<-Sys.time()
+  #FOO <- psych::corr.test(x, method = "spearman")
+  #Sys.time()-mark
   
   #Prep for partial correlation detection
   Final <- data.frame()
@@ -200,15 +185,15 @@ for( Tissue in config$tissue ){
   }
   
   #Exclude Hit Targets with zero probability:
-  Ps <- as.numeric(Ps)
+  #Ps <- as.numeric(Ps)
   at <- as.data.frame(sapply(Ps, as.numeric))
   row.names(at) <- row.names(Ps)
   
   Ps = at[,colSums(at) > 0.4  ]
   
   
-  Plots[[ Tissue ]] <- list(ParCor = t(Ps), main=Tissue, show_rownames=F)
-  pheatmap(t(Ps), main=Tissue, show_rownames=F)
+  Plots[[ Tissue ]] <- list(ParCor = t(Ps), main=Tissue)
+  #pheatmap(t(Ps), main=Tissue, show_rownames=F)
   
   eval( parse( text= paste0( 'Syn$', Tissue, '<- Syns_Used' ) ))
 }
@@ -218,22 +203,22 @@ for( Tissue in config$tissue ){
                               Tissue,
                               '$ParCor, main=Plots$',
                               Tissue,
-                              '$main)' 
[…]

### 2020-01-25 run.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+Rscript code/01-Initializer.r configs/M62.yaml
+Rscript code/01-Initializer.r configs/pQTL.yaml
+Rscript code/01-Initializer.r configs/Retromer.yaml
+Rscript code/01-Initializer.r configs/Moesin.yaml
+Rscript code/01-Initializer.r configs/EmoryTargets.yaml

### 2020-01-25 code/prepscripts/07_1-SpearmanCor.R
Working on Spearman

@@ -1,47 +1,118 @@
 library(parallel)
 library(doParallel)
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
+setwd("~/AD_TargetRank/")
+source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
+source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
 
 ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
 names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 
 for( Tissue in c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX') ){
   
-  thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
-  thisFile  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/prepscripts/07_1-SpearmanCor.R')
+  Syns_Used <- NULL
   
-  LIST <- colnames(x)
-  LIST <- combn(LIST,2)
+  #Load expression for tissue
+  exp <- read.table(syn_temp$get(as.character(ExpressionDS[Tissue]))$path, header =T, sep ='\t', row.names=1)
+  colnames( exp ) <- gsub( "X", "", colnames( exp ) )
+  Syns_Used <- c(Syns_Used, as.character(ExpressionDS[Tissue] ) )
+  #Seperate exp by tissue
+  if( Tissue == 'DLPFC'){
+  }else{
+    if( Tissue == 'TCX' | Tissue == 'CBE' ){
+      if( Tissue == 'CBE' ){
+        slec <- 'CER'
+      }else{ slec <- 'TCX' }
+      colnames( exp ) <- gsub( "TC", "TCX", colnames( exp ) )
+      exp <- exp[ , grepl( slec, colnames(exp)) ]
+    }else{
+      if( Tissue %in% c('FP', 'IFG', 'PHG', 'STG') ){
+        Syns_Used <- c( Syns_Used, 'syn21285520' )
+        Meta <- read.table( syn_temp$get('syn21285520')$path, header =T, sep ='\t', row.names=1 )
[…]

### 2020-02-11 code/03-Run.Rmd
update Mod1and7

@@
 #Assign your YAML formated config file to a varable
-configuration <- "configs/EmoryTargets.yaml"
+configuration <- "configs/M62.yaml"
 config <- read_yaml(file=paste0('./',configuration))

### 2020-02-11 code/modules/01-IGAP_module.R

@@ -1,6 +1,7 @@
 #Load IGAP files
 #IGAP <- read.table( synfile = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1_2_combined.txt", sep = "\t", header = T)
 #IGAP1 <- read.table( file = "/Users/jgockley/Desktop/Projects/AMP-AD_Targets/Candidate_Target_Profiler/RefrenceData/IGAP_stage_1.txt", sep = "\t", header = T)
+syn_temp$login()
 IGAP <- read.table( syn_temp$get('syn21534587')$path, sep = "\t", header = T)
 IGAP1 <- read.table( syn_temp$get('syn21534586')$path, sep = "\t", header = T)
 
@@ -18,6 +19,8 @@ MedianScoreIGAP <- NULL
 
 #Look at the dataframe 
 BestScoreIGAP <- NULL
+
+#save.image("TempA.RData")
 
 for( i in 1:dim(Tab)[1]){
   
@@ -84,13 +87,14 @@ row.names(Tab) <- Tab$ENSG
 
 TotalIGAP$LogP <- -log(TotalIGAP$Pvalue)
 
+#Replace Infinite values with max + one
+TotalIGAP[ TotalIGAP$LogP == Inf, ]$LogP <- max(TotalIGAP$LogP[TotalIGAP$LogP!=Inf]+1)
+
 p <- ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) + 
-  geom_violin() + geom_boxplot(width=0.1) +
+  geom_violin( ) + geom_boxplot(width=0.1, outlier.shape=NA) +
   geom_jitter(shape=16, size=.1, position=position_jitter(0.2)) +
   theme(axis.text.x = element_text(angle = 45) ) + ylab('-Log(P-Value)')  
 #p 
-#Replace Infinite values with max + one
-TotalIGAP[ TotalIGAP$LogP == Inf, ]$LogP <- max(TotalIGAP$LogP[TotalIGAP$LogP!=Inf]+1)
 
 setEPS()
 postscript(file = paste0(plotdir,'/TotalIGAP.eps') )
@@ -106,19 +110,20 @@ FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)
[…]

### 2020-02-11 code/modules/07-Coexpression.R

@@ -105,24 +105,45 @@ for( Tissue in config$tissue ){
   core <- parallel::detectCores()-2 
   source("~/AD_TargetRank/utilityFunctions/Parallel_vbsrBootstrap.R")
   #Run Partial correlations for each gene
+  RUNNe <- function( i=i, x=x){
+    OBS <- i
+    y <- as.matrix(x[,OBS]) 
+    colnames(y) <- i
+    X <- x[,(colnames(x) %in% OBS) == F ]
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=10, cores=1 )
+    names( att["intercept"] ) <- eval( parse(text = OBS))
+    return(att)
+  }
   mark <- Sys.time()
+  
+  Sys.time()-mark
+  
   for( i in row.names(Trans) ){
     OBS <- i
+    #OBS <- 'ENSG00000084234'
     y <- as.matrix(x[,OBS]) 
     colnames(y) <- i
     X <- x[,(colnames(x) %in% OBS) == F ]
     
     
     mark <- Sys.time()
-    att <- pvbsrBootstrap( y=y, x=X, nsamp=200, cores=15 )
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=10, cores=1 )
     Sys.time()-mark
     
     mark <- Sys.time()
-    att <- pvbsrBootstrap( y=y, x=X, nsamp=500, cores=15 )
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=100, cores=1 )
     Sys.time()-mark
     
     mark <- Sys.time()
-    att <- pvbsrBootstrap( y=y, x=X, nsamp=1000, cores=15 )
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=200, cores=1 )
+    Sys.time()-mark
+    
+    mark <- Sys.time()
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=500, cores=1 )
+    Sys.time()-mark
+    
+    mark <- Sys.time()
+    att <- pvbsrBootstrap( y=y, x=X, nsamp=1000, cores=1 )
     Sys.time()-mark
     
     

### 2020-02-12 code/prepscripts/PartialCorelations.R
Partial Cor Prep Script

@@ -1,0 +1,175 @@
+##Coexpression Analysis module:
+#devtools::install_github("blogsdon/spike/spike/")#, repos = NULL, type = "source")
+source("~/AD_TargetRank/utilityFunctions/Parallel_vbsrBootstrap.R")
+# Load Libraries and se working directory 
+library(pheatmap)
+library(parallel)
+library(doParallel)
+library(spike)
+library(reshape2)
+setwd( config$filedir )
+
+# Set Github Provenance links
+thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
+GL_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$genelistfile)
+CF_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$name)
+I_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/01-Initializer.r')
+M_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/02-Master.Rmd')
+
+CODE <- syn_temp$store(synapseclient$Folder(name = 'figures', parentId = RunParent))
+
+#SynIDs of expression data to pull from
+ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
+names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
+
+#Study ID Translator
+Study <- c( 'RosMap', 'Mayo', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
+names(Study) <- c('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
+
+Syn <- list('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
+
+#Useable Chrs for genelist ENSG  translations and wiki analysis preamble
+row.names(Trans) <- Trans$ensembl_gene_id
+Trans <- Trans[ Trans$chromosome_name %in% c(1:22,"X","Y"), ]
+
+writeLines(paste0('For each tissue specified in ', 
+                  paste( config$name, collapse = ', '), 
+                  " the coexpression of genes specified in ", 
+                  config$genelistfile, 
[…]

### 2020-02-12 code/prepscripts/PartialCorelations.R
Partial Cor Prep Script

@@ -59,8 +59,8 @@ for( Tissue in config$tissue ){
   activityDescription = paste0( Tissue,'Pairwise Partial Correlation Calculation');
   thisFileName <- 'PartialCorelations.R'
   # Github link
-  thisRepo <- githubr::getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
-  thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
+  thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
+  thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts/',thisFileName))
   
   activityName = 'Covariate and Diagnosis Regression';
   activityDescription = 'Covariate analysis and Regression of aligned effective counts with GRCh38 with CQN normalisation (IFG, STG, FP, PHG)';

### 2020-02-12 code/prepscripts/PartialCorelations.R
Partial Cor Prep Script

@@ -170,6 +170,14 @@ for( Tissue in config$tissue ){
   
   #Push to Synapse
   
+  # Store Matrix
+  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0(Tissue,"_PartialCorMatrix.tsv"), name = paste0(Tissue, ' Pairwise Partial correlation Matrix'), parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+  all.annotations$dataSubType = 'PartialCorrelationsForADTargetRank'
+  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
+  # Store Tables
+  ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path=paste0(Tissue,"_PartialCorTable.tsv"), name = paste0(Tissue, ' Pairwise Partial correlation Table'), parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+  all.annotations$dataSubType = 'PartialCorrelationsForADTargetRank'
+  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
 } 
 
 stopCluster(cl)

### 2020-02-18 code/modules/01-IGAP_module.R
Legend Issue Mod 1

@@ -88,8 +88,9 @@ row.names(Tab) <- Tab$ENSG
 TotalIGAP$LogP <- -log(TotalIGAP$Pvalue)
 
 #Replace Infinite values with max + one
-TotalIGAP[ TotalIGAP$LogP == Inf, ]$LogP <- max(TotalIGAP$LogP[TotalIGAP$LogP!=Inf]+1)
-
+if( Inf %in% TotalIGAP$LogP ){
+  TotalIGAP[ TotalIGAP$LogP == Inf, ]$LogP <- max(TotalIGAP$LogP[TotalIGAP$LogP!=Inf]+1)
+}else{}
 p <- ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) + 
   geom_violin( ) + geom_boxplot(width=0.1, outlier.shape=NA) +
   geom_jitter(shape=16, size=.1, position=position_jitter(0.2)) +
@@ -105,8 +106,11 @@ ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) +
 dev.off()
 
 #Replace Infinite values with max + one
-FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)
-
+if( Inf %in% FOO$Pvalue ){
+  FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)
+}else{
+  
+}
 IGAP$Rank<-rank(-IGAP$Pvalue)
 FOO<-IGAP
 FOO$Pvalue <- -log(FOO$Pvalue)
@@ -130,6 +134,7 @@ dev.off()
 FOO$Rank<-rank(FOO$Pvalue)
 FOO$RankModel <- (FOO$Rank/dim(FOO)[1])
 mylogit <- glm(RankModel ~ Pvalue, data = FOO, family = "binomial")
+family = quasibinomial(link = 'logit')
 #FOO$RankModel
 
 LogisticScorer <- function( X ){
@@ -170,7 +175,7 @@ BestScoreIGAP$LogP <- -log(BestScoreIGAP$Pvalue)
 Py <- ggplot( BestScoreIGAP, aes(x=LogP, y=Wts, col=GeneName)) +
   geom_point() + xlab("-log(P-Value)") + ylab("IGAP Weight") + 
[…]

### 2020-02-18 code/modules/03-DiffProt_module.R
Prot

@@ -1,0 +1,197 @@
+Prot <- read.csv( file = syn_temp$get('syn21534585')$path, header = T)
+Prot <- Prot[ complete.cases(Prot[,c("Log2_FC", "CI_Upr", "CI_Lwr", "PVal", "Cor_PVal")]), ]
+Prot <- Prot[complete.cases(Prot),]
+
+TargetP <- Prot[ Prot$ENSG %in% Genes,]
+
+Names <- paste0(TargetP$UniqID, '|', TargetP$ENSG)
+Names <- Names[!duplicated(Names)]
+PTab <- do.call( rbind, strsplit( Names, '\\|' ))
+
+if( table(Genes %in% PTab[,3])['TRUE'] == length(Genes) ){
+  Res <- PTab
+}else{
+  Adds <- Genes[ (Genes %in% PTab[,3]) == F ]
+  Res <- rbind( PTab, cbind( Tab[Adds,]$GeneName, NA, Adds ))
+}
+
+row.names(Res) <- paste0( Res[,1], '|', Res[,2])
+Res <- as.data.frame(Res)
+colnames( Res ) <- c( "GeneName", "UniProt", "ENSG")
+
+Tiss <- as.character(TargetP$Tissue)[ !duplicated(as.character(TargetP$Tissue)) ]
+Adds <- matrix(0, dim(Res)[1], 4)
+row.names(Adds) <- row.names(Res)
+colnames(Adds) <- c("AntPFC", "DLPFC", "MFG", "TCX" )
+
+TotP <- as.data.frame(cbind(Res, Adds))
+TotPval <- as.data.frame(cbind(Res, Adds))
+
+row.names(TotP) <- gsub('\\.', '|', row.names(TotP))
+row.names(TotPval) <- gsub('\\.', '|', row.names(TotPval))
+
+for( row in row.names(TotP)){
+  for( col in colnames(TotP)){
+    if( isTRUE(row %in%TargetP$UniqID) ){
+      TP <- TargetP[ TargetP$UniqID == row,]
+      if( isTRUE(col %in% TP$Tissue) ){
+        TotP[row,col] <- signif( TP[ as.character(TP$Tissue)==col ,]$Log2_FC, 3 )
+      }else{}
+    }else{
+      
+    }
+  }
+}
+
+for( row in row.names(TotPval)){
+  for( col in colnames(TotPval)){
+    if( isTRUE(row %in%TargetP$UniqID) ){
[…]

### 2020-02-19 code/modules/04-eQTL_module.R
Mod4

@@ -1,0 +1,24 @@
+eqtl <- read.csv( file =syn_temp$get('syn21534584')$path, header=T)
+
+Genes <- Tab$ENSG
+
+uEQs <- eqtl[ (eqtl$ensembl_gene_id %in% Genes) == T,]
+
+uEQs$hasEqtl <- as.character(uEQs$hasEqtl)
+Rep <- c("NO", "YES")
+names(Rep) <- c('FALSE', 'TRUE')
+uEQs$hasEqtl <- as.character(Rep[ as.character(uEQs$hasEqtl) ])
+
+uEQs %>%
+  select(ensembl_gene_id, hgnc_symbol, hasEqtl)  %>%
+  kable( escape = F, row.names = FALSE ) %>%
+  kable_styling("striped", full_width = F)
+
+save_kable(
+  uEQs %>%
+    select(ensembl_gene_id, hgnc_symbol, hasEqtl)  %>%
+    kable( escape = F, row.names = FALSE ) %>%
+    kable_styling("striped", full_width = F),
+  
+  file=paste0(tabledir,'/eQTL_inBrain.pdf')
+)

### 2020-02-19 code/modules/05-CoExp_module.R
Mod 5

@@ -1,0 +1,142 @@
+library(gplots)
+library(superheat)
+
+Genes <- Tab$ENSG
+Nets <- read.csv(file = syn_temp$get('syn21534603')$path, header=T)
+ZoomNet <- Nets[ (Nets$GeneID %in% Genes) == T, ]
+
+NetPlot <- as.data.frame(matrix(0,length(Genes),5))
+row.names(NetPlot) <- Genes
+NetPlot$V1 <- row.names(NetPlot)
+colnames(NetPlot) <- c("ENSG", "GeneName", "InACluster", "NumberOfClusters", "BrainRegions")
+
+for( g in Genes ){
+  if( dim(ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ])[1] > 0 ){
+    NetPlot[g,]$GeneName <- names( table( as.character( ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ]$external_gene_name ) ) )
+    NetPlot[g,]$InACluster <- 'YES'
+    tmp <- ZoomNet[ as.character(ZoomNet$GeneID) == NetPlot[g,]$ENSG, ]
+    NetPlot[g,]$NumberOfClusters <- length(tmp$Module)
+    Regions <- c(as.character(tmp$brainRegion))
+    NetPlot[g,]$BrainRegions <- paste( Regions[ !duplicated(Regions) ], collapse = ', ')
+  }else{
+    NetPlot[g,]$GeneName <- as.character( TotP[ TotP$ENSG == g, ]$GeneName )
+    NetPlot[g,]$InACluster <- 'NO'
+    NetPlot[g, 3:5] <- NA
+  }
+}
+
+NetPlot %>% 
+  select(ENSG, GeneName, InACluster, NumberOfClusters, BrainRegions)  %>%
+  kable( escape = F, row.names = FALSE) %>%
+  kable_styling("striped", full_width = F)
+
+save_kable(
+  NetPlot %>% 
+    select(ENSG, GeneName, InACluster, NumberOfClusters, BrainRegions)  %>%
+    kable( escape = F, row.names = FALSE) %>%
+    kable_styling("striped", full_width = F),
+  
+  file=paste0(tabledir,'/InNetCluster.pdf')
+)
+
+Regions <- as.data.frame( table(ZoomNet$brainRegion) )
+colnames(Regions) <- c("Region", "Targets")
+
[…]

### 2020-02-19 code/modules/06-CellType_Module.R
mod6

@@ -1,0 +1,67 @@
+
+Lake <- read.csv( file=syn_temp$get('syn21534588')$path, header = T )
+Lake <- Lake[ Lake$adj.pval < 0.05, ]
+Lake$Module <- as.character(Lake$Module)
+Zhang <- read.csv( file=syn_temp$get('syn21534606')$path, header = T )
+Zhang <- Zhang[ Zhang$adj.pval < 0.05, ]
+Zhang$Module <- as.character(Zhang$Module)
+
+Zrich <- as.data.frame( matrix(0, length(Genes), length(table(Zhang$category)) ))
+row.names(Zrich) <- trans[Genes]
+colnames(Zrich) <- names(table(Zhang$category))
+
+Lrich <- as.data.frame( matrix(0, length(Genes), length(table(Lake$category)) ))
+row.names(Lrich) <- trans[Genes]
+colnames(Lrich) <- names(table(Lake$category))
+
+for( gene in Genes ){
+  lyst <- as.character( ZoomNet[ ZoomNet$GeneID == gene, ]$Module )
+  foo <- Lake[ as.character(Lake$Module) %in% lyst ==T , ]
+  for( Cell in as.character(foo$category) ){
+    Lrich[ trans[gene], Cell  ] <- Lrich[ trans[gene], Cell  ] + 1
+  }
+  foo <- Zhang[ as.character(Zhang$Module) %in% lyst ==T , ]
+  for( Cell in as.character(foo$category) ){
+    Zrich[ trans[gene], Cell  ] <- Zrich[ trans[gene], Cell  ] + 1
+  }
+}
+save_kable(
+  Zrich %>% 
+    kable( escape = F, row.names = TRUE) %>%
+    kable_styling("striped", full_width = F),
+  file=paste0(tabledir,'/ZhangSingleCell.pdf')
+)
+save_kable(
+  Lrich %>% 
+    kable( escape = F, row.names = TRUE) %>%
+    kable_styling("striped", full_width = F),
+  file=paste0(tabledir,'/LakeSingleCell.pdf')
+)
+Zrich %>% 
+  kable( escape = F, row.names = TRUE) %>%
+  kable_styling("striped", full_width = F)
+
+Lrich %>% 
+  kable( escape = F, row.names = TRUE) %>%
+  kable_styling("striped", full_width = F)
+
[…]

### 2020-02-19 code/01-Initializer.r
intit alt

@@ -20,7 +20,7 @@ system( paste0('mkdir -p runs/', config$runname, '/figures' ) )
 system( paste0('mkdir -p runs/', config$runname, '/tables' ) )
 
 #CREATE the RMD to RUN
-system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ~/AD_TargetRank/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))
+system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ", config$filedir, "/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))
 
 reticulate::use_python("/usr/bin/python", required = TRUE)
 synapseclient <- reticulate::import("synapseclient")

### 2020-02-19 code/01-Initializer.r
intit alt

@@ -20,7 +20,7 @@ system( paste0('mkdir -p runs/', config$runname, '/figures' ) )
 system( paste0('mkdir -p runs/', config$runname, '/tables' ) )
 
 #CREATE the RMD to RUN
-system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ", config$filedir, "/code/02-Master.Rmd > ~/AD_TargetRank/code/03-Run.Rmd"))
+system( paste0( "sed 's/YAML_FILE/", sub("/", "\\\\/",config$name), "/g' ", config$filedir, "/code/02-Master.Rmd > ", config$filedir, "/code/03-Run.Rmd"))
 
 reticulate::use_python("/usr/bin/python", required = TRUE)
 synapseclient <- reticulate::import("synapseclient")

### 2020-02-19 code/01-Initializer.r
intit alt

@@ -27,7 +27,7 @@ synapseclient <- reticulate::import("synapseclient")
 syn_temp <- synapseclient$Synapse()
 syn_temp$login()
 
-setwd("~/AD_TargetRank/")
+setwd(config$filedir)
 source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
 source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
 createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/03-Run.Rmd",

### 2020-02-19 code/01-Initializer.r
intit alt

@@ -28,8 +28,8 @@ syn_temp <- synapseclient$Synapse()
 syn_temp$login()
 
 setwd(config$filedir)
-source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
-source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
-createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/03-Run.Rmd",
+source( paste0( config$filedir, "/utilityFunctions/knitfile2synapseClient.R"))
+source( paste0( config$filedir, "/utilityFunctions/hook_synapseMdSyntax_plot.R"))
+createAndKnitToFolderEntityClient(file = paste0( config$filedir, "/code/03-Run.Rmd" ),
                                   parentId =config$parentID,
                                   folderName = config$runID)

### 2020-02-19 code/modules/01-IGAP_module.R
Mod1

@@ -105,14 +105,17 @@ ggplot(TotalIGAP, aes(x=GeneName, y=LogP)) +
   theme(axis.text.x = element_text(angle = 45) ) + ylab('-Log(P-Value)') 
 dev.off()
 
+
+IGAP$Rank<-rank(-IGAP$Pvalue)
+FOO<-IGAP
+
 #Replace Infinite values with max + one
 if( Inf %in% FOO$Pvalue ){
   FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)
 }else{
   
 }
-IGAP$Rank<-rank(-IGAP$Pvalue)
-FOO<-IGAP
+
 FOO$Pvalue <- -log(FOO$Pvalue)
 #Replace Infinite values with max + one and zero vals w/ min/100
 FOO[ FOO$Pvalue == Inf,]$Pvalue <- max(FOO$Pvalue[FOO$Pvalue!=Inf]+1)

### 2020-02-19 code/03-Run.Rmd
Module2
Total Catchup

@@
 #Assign your YAML formated config file to a varable
-configuration <- "configs/M62.yaml"
+configuration <- "configs/TestConfig.yaml"
 config <- read_yaml(file=paste0('./',configuration))
@@
 
-  source('code/modules/02-DiffExp_module.R')
+  source('code/modules/02-DiffExpression_module.R')
 

### 2020-02-19 code/modules/03-DiffProt_module.R
Fix DEProt

@@ -148,7 +148,7 @@ ModelConstructProt <- function( de ){
       eval(parse( text=paste0( 'Plotz[["', Tissue, '"]] <- Pr' ) ))
     }else{
       eval(parse( text=paste0( 
-        'Pr <- ggplot( data=Modelz$', Tissue, '[ Modelz$', Tissue, '$Type == \'Predicted Weight\' & (Models$', Tissue, '$ensembl_gene_id %in% Genes) == T , ], aes(x=log_abs_PVal, y=y, col=Sig)) +
+        'Pr <- ggplot( data=Modelz$', Tissue, '[ Modelz$', Tissue, '$Type == \'Predicted Weight\' & (Modelz$', Tissue, '$ensembl_gene_id %in% Genes) == T , ], aes(x=log_abs_PVal, y=y, col=Sig)) +
               geom_point() + 
               geom_smooth( data=Modelz$', Tissue, '[ Modelz$', Tissue, '$Type == \'Predicted Weight\', ], aes(x=log_abs_PVal, y=y), 
                     method = "glm", 

### 2020-02-19 code/modules/03-DiffProt_module.R
Fix DEProt

@@ -166,7 +166,7 @@ ModelConstructProt <- function( de ){
       )))
       eval(parse( text=paste0( 'Plotz[["', Tissue, '"]] <- Pr' ) ))
       
-      ggsave( filename = paste0('DiffExpress_', Sex, '_',Tissue,'.eps'),
+      ggsave( filename = paste0('DiffExpress_',Tissue,'.eps'),
               plot = Pr,
               device = "eps",
               path = plotdir

### 2020-02-19 code/modules/03-DiffProt_module.R
Fix DEProt

@@ -1,3 +1,5 @@
+Genes <- Tab$ENSG
+
 Prot <- read.csv( file = syn_temp$get('syn21534585')$path, header = T)
 Prot <- Prot[ complete.cases(Prot[,c("Log2_FC", "CI_Upr", "CI_Lwr", "PVal", "Cor_PVal")]), ]
 Prot <- Prot[complete.cases(Prot),]

### 2020-02-20 code/modules/05-CoExp_module.R
rownames mod 5

@@ -5,8 +5,8 @@ Genes <- Tab$ENSG
 Nets <- read.csv(file = syn_temp$get('syn21534603')$path, header=T)
 ZoomNet <- Nets[ (Nets$GeneID %in% Genes) == T, ]
 
-NetPlot <- as.data.frame(matrix(0,length(Genes),5))
-row.names(NetPlot) <- Genes
+NetPlot <- as.data.frame(matrix(0,length(Genes[!duplicated(Genes)]),5))
+row.names(NetPlot) <- Genes[!duplicated(Genes)]
 NetPlot$V1 <- row.names(NetPlot)
 colnames(NetPlot) <- c("ENSG", "GeneName", "InACluster", "NumberOfClusters", "BrainRegions")
 

### 2020-04-27 code/03-Run.Rmd
Partials

--- 2020-01-25 code/03-Run.Rmd
@@
 }
+Tab <- Tab[ Tab$chromosome_name %in% c( c(1:22), "X","Y"), ]
 ```

### 2020-04-27 code/01-Initializer.r

@@ -11,7 +11,7 @@ library(yaml)
 args <- commandArgs(trailingOnly=TRUE)
 
 #READ IN Config
-#configuration<-'configs/M62.yaml'
+#configuration<-'configs/EmoryTargets.yaml'
 configuration <- args[1]
 config <- read_yaml(configuration)
 

### 2020-04-27 code/prepscripts/07_1-SpearmanCor.R

@@ -2,6 +2,7 @@ library(parallel)
 library(doParallel)
 library(data.table)
 reticulate::use_python("/usr/bin/python", required = TRUE)
+#reticulate::use_python("../miniconda2/bin/python", required = TRUE)
 synapseclient <- reticulate::import("synapseclient")
 syn_temp <- synapseclient$Synapse()
 syn_temp$login()
@@ -18,9 +19,8 @@ names( ExpressionDS ) <- c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX')
 Study <- c( 'RosMap', 'Mayo', 'Mayo', 'MSBB', 'MSBB', 'MSBB', 'MSBB')
 names(Study) <- c('DLPFC', 'TCX', 'CBE', 'FP', 'IFG', 'PHG', 'STG')
 
-
 for( Tissue in c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX') ){
-  
+  #Tissue<-'DLPFC'
   Syns_Used <- NULL
   
   #Load expression for tissue
@@ -65,15 +65,15 @@ for( Tissue in c('CBE', 'DLPFC', 'FP', 'IFG', 'PHG', 'STG', 'TCX') ){
           )
   }
   
-  mark <- Sys.time()
+  #mark <- Sys.time()
   LIST <- combn(1:length(colnames(x)),2)
-  Sys.time()-mark
+  #Sys.time()-mark
   
   rm(foo)
   rm(Exp)
   rm(exp)
   
-  cl <- makeCluster(detectCores()-2)
+  cl <- makeCluster(detectCores()-12)
   registerDoParallel(cl)
   #mark <- Sys.time()
   foo <- t(parApply(cl,LIST,2,RUNNEr,x))

### 2020-04-27 code/prepscripts/PartialCorelations.R

@@ -7,16 +7,10 @@ library(parallel)
 library(doParallel)
 library(spike)
 library(reshape2)
+library(parallel)
+library(doParallel)
+
 setwd( config$filedir )
-
-# Set Github Provenance links
-thisRepo <- githubr::getRepo(repository = "jgockley62/AD_TargetRank", ref="branch", refName='master')
-GL_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$genelistfile)
-CF_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath=config$name)
-I_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/01-Initializer.r')
-M_File  <- githubr::getPermlink(repository = thisRepo, repositoryPath='code/02-Master.Rmd')
-
-CODE <- syn_temp$store(synapseclient$Folder(name = 'figures', parentId = RunParent))
 
 #SynIDs of expression data to pull from
 ExpressionDS <- c('syn21291908','syn21292041','syn21285564','syn21285564','syn21285564','syn21285564','syn21291908')
@@ -45,49 +39,45 @@ Plots <- list()
 
 FinalTis <- data.frame()
 
-library(parallel)
-library(doParallel)
-cores <- detectCores()-2 
-cl <- makePSOCKcluster(cores)
-registerDoParallel(cl)
+cores <- parallel::detectCores()-2 
+cl <- parallel::makePSOCKcluster(cores)
+doParallel::registerDoParallel(cl)
 
+# Set Github Provenance links
+# Github link
+parentId = 'syn21534582';
+activityName = 'Partial Correlations';
+activityDescription = 'Partial correltation Data';
+thisFileName <- 'PartialCorelations.R'
+thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
+thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts/',thisFileName))
+
+### Store files in synapse
+activityName = 'Partial Correlations ';
+activityDescription = 'Pairwise partial correlations of genes for Target Prioritization';
+CODE <- syn_temp$store(synapseclient$Folder(name = "Reference Data", parentId = 'syn21534582'))
+
[…]

### 2020-04-27 utilityFunctions/Parallel_vbsrBootstrap.R

@@ -1,5 +1,7 @@
 pvbsrBootstrap = function(y,x,nsamp=100,cores=8){
   library(dplyr)
+  library(parallel)
+  library(doParallel)
   n <- length(y)
   replicateMatrix = sample(1:(n*nsamp),replace=TRUE) %>%
     matrix(n,nsamp)

### 2020-04-27 code/03-Run.Rmd

@@
 #Assign your YAML formated config file to a varable
+
 configuration <- "configs/EmoryTargets.yaml"
+configuration <- "configs/TestConfig.yaml"
 config <- read_yaml(file=paste0('./',configuration))

### 2020-05-22 code/prepscripts/MSBB_TranscriptNorm.Rmd
Trans DE MSBB

---
title: Covariate and analysis of MSBB reprocessed counts Regressing out Diagnosis as well for TWAS
author: 'JKG - Adapted from: Thanneer Perumal'
output: html_notebook
editor_options:
  chunk_output_type: console
---
  Date of analysis update: "`r date()`"

| *Synapse ID* | *File Name* |
|  -----------------------------------------  |   ---------                      |
| Old Count Matrix ------syn8691099.1 | |
| syn20963389 | |
| syn20963390 | |
| syn10156693.3 | | 
| Old Metrics------syn8698270.1 | |
| syn20963391 | | 
| Old Length GC Data-----syn8449369.2 | | 
| syn21156161 | |
| syn6100548.7 | | 
| syn6101474.4 | |

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##It is assumed your working directory is "~/ampad-DiffExp/gene_level_analysis"
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(psych)
library(limma)
library(edgeR)
#BiocManager::install("biomaRt")
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(glmnet)
library(knitr)
library(doParallel)
library(foreach)
library(githubr)
#cl = makeCluster(detectCores()-2)
cl = makeCluster(14)
registerDoParallel(cl)
options(xtable.type="html")
setwd('~/AD_TargetRank/code/')
# Source modified utility functions from limma and sva package
source('../utilityFunctions/parallelDuplicateCorrelation.R')
source('../utilityFunctions/irwsva.build.R')
source('../utilityFunctions/f.pvalue.R')
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Data download
Obtain count matrix and metadata from synapse
```{r download.data}
# Get reprocessed sample counts for temporal cortex
COUNT_ID <- 'syn10507730';
ALL_USED_IDs <- COUNT_ID
COUNT <- syn_temp$get(COUNT_ID)$path %>%
  read.table(header=T, sep='\t', check.names = F, row.names = 1)

Feature_Name_Sink <- row.names(COUNT)
row.names(COUNT) <- do.call( rbind, strsplit(row.names(COUNT),split='\\|' ) )[,1]

COUNT <- COUNT[ rowSums(COUNT[,] > 0) / (ncol(COUNT)) >= 0.05, ]
# Convert rownames of counts from tracking id to ensemble gene id
tmp = data.frame(Gene.ID = rownames(COUNT)) %>%
  dplyr::mutate(ID = Gene.ID) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(tmp) = tmp$Gene.ID
rownames(COUNT) = tmp[rownames(COUNT), 'ensembl_gene_id']
# Get sample ids
SampleID = data.frame(SampleID = colnames(COUNT),
                      ID = colnames(COUNT)) %>% 
  tidyr::separate(ID, c('A','B','ID'), sep = '_') %>%
  dplyr::select(SampleID, ID) %>%
  dplyr::mutate(ID = as.numeric(ID))
# Get technical metadata
METADATA_TECH_ID <- 'syn6100548'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TECH_ID
METADATA_TECH <- syn_temp$get(METADATA_TECH_ID, version = 7)$path %>%
  read.table(sep=',',header=T, row.names=1) %>%
  group_by(sampleIdentifier) %>%
  top_n(1, -rRNA.rate) %>%
  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN) %>%
  unique %>%
  ddply(.(barcode), .fun = function(x){
    x = dplyr::mutate(x, batch = paste(batch, collapse = ''), RIN = sum(RIN, na.rm = T))  
  }) 
# Get 90+ age
METADATA_AGE_ID = 'syn10156693'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_AGE_ID
METADATA_AGE <- syn_temp$get(METADATA_AGE_ID, version = 3)$path %>%
  read.table(sep='\t', header=T)
# Get clinical metadata
METADATA_CLINICAL_ID <- 'syn6101474'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CLINICAL_ID
METADATA_CLINICAL <- syn_temp$get(METADATA_CLINICAL_ID)$path %>%
  read.csv(header = T)
# Get picard metrics from synapse
METADATA.PICARD_ID <- 'syn8698270';#'syn20963391';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_ID
METADATA.PICARD <- syn_temp$get(METADATA.PICARD_ID)$path %>%
  read.table(sep='\t', header=T, stringsAsFactors=F) %>% 
  dplyr::rename(sampleIdentifier = sample)

#REplace Syn IDs
#row.names(METADATA.PICARD_Old) <- as.character( Translator[ METADATA.PICARD_Old$sampleIdentifier , ][,2] )
row.names(METADATA.PICARD) <- METADATA.PICARD$sampleIdentifier 
row.names(METADATA_CLINICAL) <- METADATA_CLINICAL$individualIdentifier
#METADATA_CLINICAL <- METADATA_CLINICAL[ , colnames(METADATA_CLINICAL) != 'individualIdentifier']

###Left Here:
colnames(METADATA_CLINICAL)[ colnames(METADATA_CLINICAL) == 'individualID'] <- 'individualIdentifier'
colnames(METADATA_CLINICAL)[ colnames(METADATA_CLINICAL) == 'ageDeath'] <- 'AOD'
row.names(METADATA_TECH) <- METADATA_TECH$sampleIdentifier
METADATA <- dplyr::full_join(METADATA_TECH, METADATA_CLINICAL) %>%
  dplyr::inner_join(METADATA.PICARD ) %>%
  dplyr::select(-AOD) %>%
  dplyr::left_join(METADATA_AGE)
colnames(METADATA) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA))
colnames(METADATA) = gsub('RnaSeqMetrics__','',colnames(METADATA))

#Get Transcript parameters...
#gene_id <- 'ensembl_transcript_id'
#filters <- 'ensembl_transcript_id'
#host <- 'ensembl.org'
#organism <- 'hsa'

# Parse gene IDs to use in query
#gene_ids <- row.names(COUNT)
#gene_ids <- do.call( rbind, strsplit( gene_ids, "[.]"))[,1]
#attrs <- c(filters, "ensembl_exon_id", "chromosome_name", "exon_chrom_start", "exon_chrom_end")

#ensembl <- biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host = host)

#ds <- biomaRt::listDatasets(ensembl)[, "dataset"]
#ds <- grep(paste0("^", organism), ds, value = TRUE)
#  if (length(ds) == 0) {
#    stop(paste("Mart not found for:", organism))
#  } else if (length(ds) > 1) {
#    message("Found several marts")
#    sapply(ds, function(d) message(paste(which(ds == d), d, sep = ": ")))
#    n <- readline(paste0("Choose mart (1-", length(ds), ") : "))
#    ds <- ds[as.integer(n)]
#  }
#  ensembl <- biomaRt::useDataset(ds, mart = ensembl)

#coords <- biomaRt::getBM(filters = filters,
#                             attributes = attrs,
#                             values = gene_ids,
#                             mart = ensembl,
#                             useCache = FALSE)

#length <- plyr::ldply( names(coords), function(x) c( x , sum(IRanges::width(coords[[x]]))), .id = c("ensembl_gene_id","length") )  %>% 
#      dplyr::rename( gene=.data$V1 ) %>% 
#      dplyr::rename( length=.data$V2 )
#lsink<-length
#gc_content <- biomaRt::getBM(filters = filters,
#                                 attributes = c(filters, "hgnc_symbol", "percentage_gene_gc_content","gene_biotype", "chromosome_name"),
#                                 values = gene_ids,
#                                 mart = ensembl,
#                                 useCache = FALSE)
#colnames(length)[1] <- 'ensembl_transcript_id'
#biomart_results <- dplyr::full_join(gc_content, length)
#sink<-biomart_results

biomart_results <- read.table(syn_temp$get('syn22087693')$path, header=T, sep='\t', stringsAsFactors = F)

# Duplicate Ensembl Ids are collapsed into a single entry
collapse_duplicate_trans <- function(biomart_results){
  biomart_results %>%
    dplyr::group_by(.data$ensembl_transcript_id) %>%
    dplyr::mutate(hgnc_symbol = paste(.data$hgnc_symbol, collapse = ", ")) %>%
    unique()
}
biomart_results <- as.data.frame( collapse_duplicate_trans(biomart_results), stringsAsFactors = F )
row.names(biomart_results) <- biomart_results$ensembl_transcript_id
biomart_results <- biomart_results[ row.names(COUNT), ]
# Biomart IDs as rownames
#biomart_results <- tibble::column_to_rownames(biomart_results, var = 'ensembl_transcript_id')
#biomart_results <-  as.data.frame(biomart_results)
#biomart_results <- read.table('MSBB_biomart_results.tsv', sep='\t', header =T)
#write.table(biomart_results, file='MSBB_biomart_results.tsv', col.names =T, row.names=F, sep='\t', quote=F)

GENE.PARAM = biomart_results
GENE.GC = dplyr::select(GENE.PARAM, ensembl_transcript_id, percentage_gene_gc_content) %>% unique()
colnames(GENE.GC) = c('ensembl_gene_id', 'percentage_gc_content')

GENE.LEN = dplyr::select(GENE.PARAM, ensembl_transcript_id, length) %>% unique()
colnames(GENE.LEN) = c('ensembl_gene_id', 'gene.length')

```

### Data preprocessing
```{r preprocess.data,cache=FALSE, echo=FALSE}
# Choose samples with rRNA rate less than 5% and more PF_READS_ALIGNED
colnames(METADATA)[ colnames(METADATA) %in% 'pmi' ] <- 'PMI'
colnames(METADATA)[ colnames(METADATA) %in% 'Braak' ] <- 'bbscore'
colnames(METADATA)[ colnames(METADATA) %in% 'race' ] <- 'RACE'
colnames(METADATA)[ colnames(METADATA) %in% 'CERAD' ] <- 'NP.1'

#colnames(METADATA)[ colnames(METADATA) %in% '' ] <- ''

METADATA = as_tibble(METADATA, .name_repair = "unique") %>%
  plyr::ddply(.(barcode), .fun = function(x){
    x %>%
      dplyr::filter(PCT_RIBOSOMAL_BASES < 0.05) %>%
      dplyr::top_n(1, PF_READS_ALIGNED)
  }) %>%
  dplyr::filter(!is.na(BrodmannArea), 
                !is.na(CDR), 
                !is.na(PMI), 
                !is.na(RIN),
                !is.na(bbscore),
                !is.na(NP.1),
                !is.na(sampleIdentifier),
                RIN >= 5,
                BrodmannArea %in% c('BM10', 'BM22', 'BM36', 'BM44'))
# Fix metadata
ST <- c('M', 'F')
names(ST) <- c('male', 'female')

METADATA$sex <- as.character( ST[as.character(METADATA$sex)] )
METADATA = METADATA %>%
  dplyr::rename(Sex = sex) %>%
  dplyr::mutate(AOD = gsub('\\+','',AOD),
                Tissue = factor(BrodmannArea,
                                     levels = c('BM10', 'BM22', 'BM36', 'BM44'),
                                     labels = c('BM10' = 'FP', 'BM22' = 'STG', 
                                                'BM36' = 'PHG', 'BM44' = 'IFG')), 
                RIN2 = RIN^2, 
                Sex = factor(Sex, levels = c('M','F'), labels = c('MALE', 'FEMALE')),
                Tissue1 = Tissue)
rownames(METADATA) = METADATA$sampleIdentifier
# Get harmonised case control definition
METADATA$Diagnosis = 'OTHER'
METADATA$Diagnosis[METADATA$CDR <= 0.5 & METADATA$bbscore <= 3 & METADATA$NP.1 <= 1] = 'CONTROL'
METADATA$Diagnosis[METADATA$CDR >= 1 & METADATA$bbscore >= 4 & METADATA$NP.1 >= 2] = 'AD'
METADATA$Tissue.Diagnosis = paste(METADATA$Tissue, METADATA$Diagnosis, sep = '.')
# Fix missing batch
levels(METADATA$batch)[levels(METADATA$batch) == ''] = 'NoBatch'
# Introduce ApoE4
METADATA$APOE4 = 0

METADATA$Apo1 <- do.call( rbind, strsplit( as.character(METADATA$apoeGenotype), "" ))[,1] 
METADATA$Apo2 <- do.call( rbind, strsplit( as.character(METADATA$apoeGenotype), "" ))[,2]

METADATA$APOE4[METADATA$Apo1 == 4 | METADATA$Apo2 == 4] = 1
METADATA$APOE4[METADATA$Apo1 == 4 & METADATA$Apo2 == 4] = 2
METADATA$Tissue.APOE4 = paste(METADATA$Tissue, METADATA$APOE4, sep = '.')
# Match covariates to expression data
indToRetain = intersect(METADATA$sampleIdentifier, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$sampleIdentifier)
COUNT <- COUNT[, indToRetain]
METADATA = as.data.frame(METADATA)
rownames(METADATA) = METADATA$sampleIdentifier
METADATA = METADATA[indToRetain,]
```
`r dim(COUNT)[2]` samples from `r length(unique(METADATA$individualIdentifier))` subjects were obtained from the MSSM cohorts in AMP-AD reprocessed rnaseq project.

Following `r length(indRemoved)` samples were removed: `r indRemoved`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering}
primaryVariable <- c("CDR", "Tissue.Diagnosis", "Tissue.APOE4")
#_# ADD DIAGNOSIS HERE TO REGRESS FOR TWAS
FactorCovariates <- c("individualIdentifier", "batch", "RACE", "Sex", "Tissue.Diagnosis", "Tissue.APOE4")
ContCovariates <- c("CDR", "RIN", "RIN2", "AOD","PMI",  "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES")
METADATA$Diagnosis <- as.factor(METADATA$Diagnosis)
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$sampleIdentifier
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, cache=FALSE, fig.width=10, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```
### Explore metadata
```{r data.explore, cache=FALSE, fig.width = 12, fig.height = 8}
my.theme = theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme
# Age Of Death
p[[2]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(AOD))) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AOD') + my.theme
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PMI))) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + my.theme
# Ribosomal bases
p[[4]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_RIBOSOMAL_BASES))) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Fraction of Ribosomal Bases') + my.theme
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTRONIC_BASES))) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + my.theme
# Intergenic bases
p[[6]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTERGENIC_BASES))) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Intergenic Bases') + my.theme
multiplot(plotlist = p, cols = 3)
```

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Tissue x Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}

####Need to conserve computation space: filter for protien coding
COUNT <- COUNT[ row.names(COUNT) %in% biomart_results[ biomart_results$gene_biotype=='protein_coding',]$ensembl_transcript_id, ]

row.names(COUNT) <- do.call(rbind, strsplit( row.names(COUNT), "[.]") )[,1]
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(Tissue.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object
#mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "jul2019.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
## Query biomart
#Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
                       #filters = "ensembl_gene_id", 
                       #values = PROCESSED_COUNTS$filteredExprMatrix$genes,
                       #mart = mart)
Ensemble2HGNC <- biomart_results[ biomart_results$ensembl_transcript_id %in% row.names(PROCESSED_COUNTS$filteredExprMatrix$genes), ]

summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```

### Library Normalisation
Library normalisation is performed using cqn (conditional quantile normalisation)
```{r cqn}
GENE.GC <- as.data.frame(GENE.GC[complete.cases(GENE.GC),] )
row.names(GENE.GC) <- GENE.GC$ensembl_gene_id
GENE.LEN <-as.data.frame(GENE.LEN[complete.cases(GENE.LEN),])
row.names(GENE.LEN) <- GENE.LEN$ensembl_gene_id
GENE.LEN$gene.length <- as.numeric(GENE.LEN$gene.length)
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Outlier Analysis
#### Sample outliers
Outlier analysis is performed before library normalisation with raw cpm counts
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c('hB_RNA_10622')
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Diagnosis, size=RIN, shape = Sex))
p <- p + theme_bw() + theme(legend.position="right")
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Diagnosis+.~Tissue)
p
indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]
tmp = COVARIATES %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.') %>%
  group_by(Tissue, Diagnosis) %>%
  summarise(count = n()) %>%
  spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are: `r kable(tmp)`

#### Gene outliers
Assign NA values to genes that are above and below 3 std deviation of its distribution
```{r winsorise.data}
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
CQN.GENE_EXPRESSION$E.no.na = CQN.GENE_EXPRESSION$E
CQN.GENE_EXPRESSION$E.no.na[is.na(CQN.GENE_EXPRESSION$E.no.na)] = 0
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```
### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data1.1, cache=FALSE, fig.height=5, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Region','Diagnosis'), sep = '\\.') %>% 
  dplyr::mutate(Region = factor(Region), batch = factor(batch))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Diagnosis, size=RIN, shape = Sex))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(.~Region, scales = 'free_y')
p
```
Tree based classification of samples
```{r decompse.normalise.data1.2, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = (COVARIATES[,c(FactorCovariates), drop = F])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```
Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(CQN.GENE_EXPRESSION$E.no.na))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, cache=FALSE, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix.

NOTE:
1. Using a mixed effect model where random effect is chosen as individualIdentifier
2. Adding batch and Sex a priori to variable selection
3. Primary variable of interest Tissue.Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis', warning = FALSE}
# Primary variable of interest
postAdjustCovars = c('batch', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'individualIdentifier'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
loopCount = 0 
notEstimable <- NULL
Start <- Sys.time()
foo <- 0
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){ #_# <= 8
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects and individualIdentifier as random effect'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  cnts = NEW.COUNTS
  cnts[is.na(cnts)] = 0
  VOOM.GENE_EXPRESSION = voom(cnts, design=DM1$design, 
                              plot=F)#, na.rm = T)

  # Calculate correlation values of random effects
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$individualIdentifier,
                                             method = 'lmer')
  # correlation = list(); correlation$cor = 0.3819258

  # Re-estimate voom weights
  if (!is.nan(correlation$cor)){
    # Re-estimate voom weights
    VOOM.GENE_EXPRESSION = voom(cnts, 
                                design=DM1$design, plot=F, 
                                block = COVARIATES$individualIdentifier, 
                                correlation = correlation$cor)

    # Fit linear model using new weights and new design
    VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                              design = DM1$design,
                              weights = VOOM.GENE_EXPRESSION$weights,
                              block = COVARIATES$individualIdentifier, 
                              correlation = correlation$cor)

    # Residuals after normalisation
    RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                  CQN.GENE_EXPRESSION$E)
  }else {
    notEstimable = c(notEstimable, postAdjustCovars[length(postAdjustCovars)])
    postAdjustCovars = postAdjustCovars[1:(length(postAdjustCovars)-1)]
  }

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'individualIdentifier'))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  expr = RESIDUAL.GENE_EXPRESSION; expr[is.na(expr)] = 0; expr = expr[rowSums(expr)!=0,]
  residualSigCovars = runPCAandPlotCorrelations(expr, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
  # print(loopCount)
}
modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')
tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

### Sanity check
```{r residual.adj, cache=FALSE, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(expr))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Region','Diagnosis'), sep = '\\.') %>% 
  dplyr::mutate(Region = factor(Region))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Diagnosis, size=RIN, shape = Sex))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(.~Region, scales = 'free_y')
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, cache=FALSE, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,FactorCovariates])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
#save.image('MSBB_Image.RData')
```

### Adjust data with covariates for Network Analysis
Identified covariates are regressed out from the expression matrix for network analysis
```{r network.adjust}
# Get design matrix
DESIGN.NET = getDesignMatrix(COVARIATES[, postAdjustCovars, drop = F], Intercept = F)
DESIGN.NET = DESIGN.NET$design[,linColumnFinder(DESIGN.NET$design)$indepCols]
# Re-estimate voom weights
VOOM.NET.WEIGHTS = voom(cnts, 
                        design = DESIGN.NET, plot=F,
                        block = COVARIATES$individualIdentifier, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS)
# Residuals after normalisation
RESIDUAL.NET.GENE_EXPRESSION = residuals.MArrayLM(VOOM.NET.FIT,
                                                  CQN.GENE_EXPRESSION$E)
```

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Tissue.Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust, cache=FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts, 
                    design = MODEL1, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION;
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.2)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
##JKG Add in in case lack of valuse adds singularity issue
#if( length(table(MODEL0[,'PCT_RIBOSOMAL_BASES'])) < .05*dim(MODEL0)[1] ){
#  MODEL0 <- MODEL0[,colnames(MODEL0)[colnames(MODEL0)!='PCT_RIBOSOMAL_BASES'] ]
#}else{}
#if( length(table(MODEL1[,'PCT_RIBOSOMAL_BASES'])) < .05*dim(MODEL1)[1] ){
#  MODEL1 <- MODEL1[,colnames(MODEL1)[colnames(MODEL1)!='PCT_RIBOSOMAL_BASES'] ]
#}else{}
SURR.VAR = sva::sva(CQN.GENE_EXPRESSION$E.no.na, MODEL1, MODEL0, n.sv = NUM.SV, B = 30)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$individualIdentifier, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Tissue.Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
#save.image('MSBB_Image.RData')
```

### Differential expression analysis (with Tissue.CDR as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Tissue.CDR
1. FP.CDR: change in CDR in frontal pole where 0 no cognitive decline and 5 severe cognitive decline
2. IFG.CDR: change in CDR in inferior frontal gyrus where 0 no cognitive decline and 5 severe cognitive decline
3. PHG.CDR: change in CDR in parahippocampal gyrus where 0 no cognitive decline and 5 severe cognitive decline
4. STG.CDR: change in CDR in superior temporal gyrus where 0 no cognitive decline and 5 severe cognitive decline

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=6, fig.width=15}
# Get design matrix
COVARIATES.tmp = COVARIATES %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.') %>%
  dplyr::mutate(Tissue = factor(Tissue))
DESIGN = getDesignMatrix(COVARIATES.tmp[, c('Tissue', postAdjustCovars), drop = F], Intercept = F)
ind = grep('Tissue', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind] * as.numeric(COVARIATES.tmp[,'CDR'])
colnames(DESIGN$design)[ind] = paste0( colnames(DESIGN$design)[ind],'.CDR')
DESIGN$design = DESIGN$design[,c(linColumnFinder(DESIGN$design)$indepCols)]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Fit contrast
contrast = makeContrasts(contrasts = c("TissueFP.CDR", 
                                       "TissueIFG.CDR", 
                                       "TissuePHG.CDR", 
                                       "TissueSTG.CDR"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
#sink<-DE
#_# Reformat Gene Param for combination:
GENE.PARAM <- GENE.PARAM[ row.names(DE[[1]]) , ]
colnames(GENE.PARAM)[ colnames(GENE.PARAM) == 'ensembl_transcript_id' ] <- 'ensembl_gene_id'

DE = DE %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue','',Comparison),
                Study = 'MSSM',
                Sex = 'ALL') %>%
  tidyr::separate(Comparison, c('Tissue', 'Comparison1'), sep = '\\.') %>%
  dplyr::rename(Comparison = Comparison1) %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique())
DE$Direction = 'NONE'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC <= -log2(1.2)] = 'DOWN'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC >= log2(1.2)] = 'UP'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue) %>%
#  group_by(Tissue, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position = 'top')
p = p + facet_grid(Comparison+.~Tissue, scales = 'free')
p
all.diff.exp = list(SourceDiagnosis = DE)
all.fit = list(SourceDiagnosis = FIT)
#save.image('MSBB_Image.RData')
```

### Differential expression analysis (with Tissue.Diagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Tissue.Diagnosis
1. FP.AD: Samples from frontal pole region with CDR >= 1 and bbscore >= 4 and NP.1 >= 2
2. FP.CONTROL: Samples from frontal pole region with CDR <= 0.5 and bbscore <= 3 and NP.1 <= 1
3. FP.OTHER: All the other frontal pole samples
4. PHG.AD: Samples from parahippocampal gyrus region with CDR >= 1 and bbscore >= 4 and NP.1 >= 2
5. PHG.CONTROL: Samples from parahippocampal gyrus region with CDR <= 0.5 and bbscore <= 3 and NP.1 <= 1
6. PHG.OTHER:All the other parahippocampal gyrus samples
7. STG.AD: Samples from frontal superior temporal gyrus with CDR >= 1 and bbscore >= 4 and NP.1 >= 2
8. STG.CONTROL: Samples from superior temporal gyrus region with CDR <= 0.5 and bbscore <= 3 and NP.1 <= 1
9. STG.OTHER: All the other superior temporal gyrus samples
10. IFG.AD: Samples from inferior frontal gyrus region with CDR >= 1 and bbscore >= 4 and NP.1 >= 2
11. IFG.CONTROL: Samples from inferior frontal gyrus region with CDR <= 0.5 and bbscore <= 3 and NP.1 <= 1
12. IFG.OTHER: All the other inferior frontal gyrus samples

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp1, fig.height=15, fig.width=13}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Fit contrast
cntr = c()
for (region in c('FP', 'STG', 'PHG', 'IFG')){
  tmp = combn(paste0('Tissue.Diagnosis', region, '.', c('AD', 'OTHER', 'CONTROL')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.Diagnosis','',Comparison),
                Study = 'MSSM',
                Sex = 'ALL') %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue1','against'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  dplyr::select(-Tissue1) %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique())
DE$Direction = 'NONE'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC <= -log2(1.2)] = 'DOWN'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC >= log2(1.2)] = 'UP'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue) %>%
#  group_by(Tissue, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position='top')
p = p + facet_grid(Tissue+.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis = DE))
all.fit = c(all.fit, list(Diagnosis = FIT))
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=15, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = log10(length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(Tissue~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')
pl[[2]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = percentage_gene_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gene_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(Tissue~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')
pl[[3]] = ggplot(DE %>% dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(Tissue~.+Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')
multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (with Tissue.APOE4 as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp2, fig.height=15, fig.width=13}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[3], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Fit contrast
cntr = c()
for (region in c('FP', 'STG', 'PHG', 'IFG')){
  tmp = combn(paste0('Tissue.APOE4', region, '.', c('2', '1', '0')),2)
  cntr = c(cntr, apply(tmp, 2, paste, collapse = '-'))
}
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.APOE4','',Comparison),
                Study = 'MSSM',
                Sex = 'ALL') %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue1','against'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  dplyr::select(-Tissue1) %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique())
DE$Direction = 'NONE'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC <= -log2(1.2)] = 'DOWN'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC >= log2(1.2)] = 'UP'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue) %>%
#  group_by(Tissue, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red')) + theme(legend.position='top')
p = p + facet_grid(Tissue+.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(APOE4 = DE))
all.fit = c(all.fit, list(APOE4 = FIT))
```

### Differential expression analysis (with Tissue.Diagnosis and Sex)
Differential expression is performed on the Tissue.Diagnosis x Sex variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp3, fig.height=15, fig.width=13}
## Modeling Diagnosis, msex and age_death conjointly
COVARIATES$Tissue.Diagnosis.Sex = paste(COVARIATES$Tissue.Diagnosis,COVARIATES$Sex, sep = '.') %>% factor
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('Tissue.Diagnosis.Sex', setdiff(postAdjustCovars, 'Sex')), drop = F], 
                         Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Fit contrast
cntr = c('Tissue.Diagnosis.SexIFG.AD.MALE-Tissue.Diagnosis.SexIFG.CONTROL.MALE',
         'Tissue.Diagnosis.SexIFG.AD.FEMALE-Tissue.Diagnosis.SexIFG.CONTROL.FEMALE',
         'Tissue.Diagnosis.SexFP.AD.MALE-Tissue.Diagnosis.SexFP.CONTROL.MALE',
         'Tissue.Diagnosis.SexFP.AD.FEMALE-Tissue.Diagnosis.SexFP.CONTROL.FEMALE',
         'Tissue.Diagnosis.SexPHG.AD.MALE-Tissue.Diagnosis.SexPHG.CONTROL.MALE',
         'Tissue.Diagnosis.SexPHG.AD.FEMALE-Tissue.Diagnosis.SexPHG.CONTROL.FEMALE',
         'Tissue.Diagnosis.SexSTG.AD.MALE-Tissue.Diagnosis.SexSTG.CONTROL.MALE',
         'Tissue.Diagnosis.SexSTG.AD.FEMALE-Tissue.Diagnosis.SexSTG.CONTROL.FEMALE')
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.Diagnosis.Sex','',Comparison),
                Study = 'MSSM') %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue','reference','Sex'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue1','against','Sex1'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  dplyr::select(-Tissue1, -Sex1) %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique())
DE$Direction = 'NONE'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC <= -log2(1.2)] = 'DOWN'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC >= log2(1.2)] = 'UP'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Tissue, Direction, Sex) %>%
#  dplyr::group_by(Comparison, Tissue, Direction,Sex) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  tidyr::spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
all.diff.exp = c(all.diff.exp, list(Diagnosis.Sex = DE))
all.fit = c(all.fit, list(Diagnosis.Sex = FIT))
```

### Differential expression analysis (with Tissue.Diagnosis and AOD)
Differential expression is performed on the Tissue.Diagnosis x AOD variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp4, fig.height=15, fig.width=13}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], setdiff(postAdjustCovars, 'AOD')), drop = F], Intercept = F)
ind = grep('Diagnosis', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind] * COVARIATES[rownames(DESIGN$design),'AOD']
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
colnames(DESIGN$design)[ind] = gsub('Tissue.Diagnosis', '', paste0(colnames(DESIGN$design)[ind], '.AOD'))
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$individualIdentifier, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)
# Fit contrast
cntr = c("FP.AD.AOD-FP.CONTROL.AOD",
         "IFG.AD.AOD-IFG.CONTROL.AOD",
         "PHG.AD.AOD-PHG.CONTROL.AOD",
         "STG.AD.AOD-STG.CONTROL.AOD")
contrast = makeContrasts(contrasts=cntr,
                         levels = colnames(FIT$coefficients))
colnames(contrast) = cntr
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  data.table::rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Study = 'MSSM', Sex = 'ALL') %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue','reference','Sex'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue1','against','Sex1'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  dplyr::select(-Tissue1, -Sex1, -Sex) %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique())
DE$Direction = 'NONE'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC <= -log2(1.2)] = 'DOWN'
DE$Direction[DE$adj.P.Val <= 0.05 & DE$logFC >= log2(1.2)] = 'UP'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Tissue, Comparison, Direction) %>%
#  group_by(Comparison, Tissue, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
#save.image('MSBB_Image.RData')
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22092490';
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective protien coding transcript level counts with CQN normalisation (IFG, STG, FP, PHG)';
thisFileName <- 'MSBB_TranscriptNorm.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts',thisFileName))
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective protien coding transcript level counts with CQN normalisation (IFG, STG, FP, PHG)'
CODE <- syn_temp$store(synapseclient$Folder(name = "MSBB", parentId = parentId))
#Set Used SynIDs For Provenance
Syns_Used <- c("syn10507730", "syn6100548", "syn10156693", "syn6101474", "syn8698270", "syn22087693")
# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= c('IFG', 'STG', 'FP', 'PHG'), 
  study = 'MSBB', 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh38'
)
# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'MSBB_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'MSBB_SVA_Diagnosis_Covariates.tsv', row.names = F, sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_SVA_Diagnosis_Covariates.tsv', name = 'Covariates', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'covariates'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_Counts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_Counts.tsv', name = 'Counts (filtered raw)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store Estimated counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_eCounts.tsv', name = 'Counts (estimated)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'estimatedCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_logCPM.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredLCPM'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_offset.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_offset.tsv', name = 'Gene length and GC content offset', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'offset'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_Design.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_Design.tsv', name = 'Design Matrix', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'designMatrix'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_SurVar.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_SurVar.tsv', name = 'Surrogate Variables', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'surrogateVariables'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MSBB_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSBB_Diagnosis_Regressed_netResidualExpression.tsv', name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store DE Results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'MSSM_FP_STG_PHG_IFG_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MSSM_FP_STG_PHG_IFG_DiffExpression.tsv', name = 'Differential Expression Results (Tissue.Diagnosis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'diffExp'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

#_#stopCluster(cl)
```

```{r DelFiles, echo=T, results='hide'}
#system('rm *.tsv')
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp  <- synapseclient$Synapse()
syn_temp$login()
setwd("~/AD_TargetRank/code/")
source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
createAndKnitToFolderEntityClient(file = "MSBB_TranscriptNorm.Rmd",
                                          parentId ="syn22092490",
                                          folderName = 'MSBB')
```

### 2020-05-22 code/prepscripts/MSBB_TranscriptNorm.Rmd
Trans DE MSBB

@@
 thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
-thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts',thisFileName))
+thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts/',thisFileName))
 ```

### 2020-05-23 code/prepscripts/Mayo_TranscriptNorm.Rmd
Trans DE MSBB

---
title: Covariate Analysis of MAYO Reprocessed counts Regressing out Diagnosis as well for TWAS
author: 'JKG - Adapted from: Thanneer Perumal'
output: html_notebook
editor_options:
  chunk_output_type: console
---
Date of analysis update: "`r date()`"

| *Synapse ID* | *File Name* |
|  -----------------------------------------  |   ---------                      |
| syn5223705.5 | | 
| syn8698211.1 | | 
| syn3817650.8 | | 
| syn6126119.3 | | 
| syn6126114.3 | | 
| syn6145639.5 | | 
| syn20830935.1 | |
| syn20825567.1 | |
| syn21156161.2 | |

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##It is assumed your working directory is "~/ampad-DiffExp/gene_level_analysis"
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(psych)
library(limma)
library(edgeR)
#BiocManager::install("biomaRt")
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(glmnet)
library(knitr)
library(doParallel)
library(foreach)
library(githubr)
#cl = makeCluster(detectCores()-2)
cl = makeCluster(14)
registerDoParallel(cl)
options(xtable.type="html")
setwd('~/AD_TargetRank/code/')
# Source modified utility functions from limma and sva package
source('../../utilityFunctions/parallelDuplicateCorrelation.R')
source('../../utilityFunctions/irwsva.build.R')
source('../../utilityFunctions/f.pvalue.R')
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Data download
Obtain count matrix and metadata from Synapse
```{r download.data}
# Get reprocessed sample counts for temporal cortex
COUNT_ID <- 'syn10499227';
ALL_USED_IDs <- COUNT_ID
COUNT_CBE <- syn_temp$get(COUNT_ID)$path %>%
  read.table(header=T, sep='\t', check.names = F, row.names = 1)

COUNT_ID <- 'syn10507722';
ALL_USED_IDs <- COUNT_ID
COUNT_TCX <- syn_temp$get(COUNT_ID)$path %>%
  read.table(header=T, sep='\t', check.names = F, row.names = 1)

COUNT <- as.data.frame( cbind(COUNT_CBE, COUNT_TCX) )

Feature_Name_Sink <- row.names(COUNT)
row.names(COUNT) <- do.call( rbind, strsplit(row.names(COUNT),split='\\|' ) )[,1]
row.names(COUNT) <- do.call( rbind, strsplit(row.names(COUNT),split='[.]' ) )[,1]

#Need to filter low counts here to remove the PAR_Y Regions
COUNT <- COUNT[ rowSums(COUNT[,] > 0) / (ncol(COUNT)) >= 0.05, ]

tmp = data.frame(Gene.ID = rownames(COUNT)) %>%
  dplyr::mutate(ID = Gene.ID) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(tmp) = tmp$Gene.ID
rownames(COUNT) = tmp[rownames(COUNT), 'ensembl_gene_id']
# Get clinical metadata
METADATA_TC_ID <- 'syn3817650'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_TC_ID
METADATA_TC <- syn_temp$get(METADATA_TC_ID)$path %>%
  read.table(sep=',',header=T, stringsAsFactors=F) %>%
  tidyr::separate(ID, c('Donor_ID', 'Tissue'), sep = '_') %>%
  dplyr::mutate(ID = paste(Donor_ID, Tissue, sep = '_'))
# Get clinical metadata
METADATA_CE_ID <- 'syn5223705'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_CE_ID
METADATA_CE <- syn_temp$get(METADATA_CE_ID)$path %>%
  read.table(sep=',',header=T, stringsAsFactors=F) %>%
  tidyr::separate(SampleID, c('Donor_ID', 'Tissue')) %>%
  dplyr::mutate(ID = paste(Donor_ID, Tissue, sep = '_')) %>%
  dplyr::rename(Gender = Sex, FLOWCELL = Flowcell)
# Get picard metrics from synapse
METADATA_PICARD_CE_ID <- 'syn8698214';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_PICARD_CE_ID
METADATA_PICARD_CE <- syn_temp$get(METADATA_PICARD_CE_ID)$path %>%
  read.table(sep='\t', header=T, stringsAsFactors=F) %>% 
  dplyr::rename(ID = sample)
colnames(METADATA_PICARD_CE) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA_PICARD_CE))
colnames(METADATA_PICARD_CE) = gsub('RnaSeqMetrics__','',colnames(METADATA_PICARD_CE))
METADATA_PICARD_TC_ID <- 'syn8698211';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_PICARD_TC_ID
METADATA_PICARD_TC <- syn_temp$get(METADATA_PICARD_TC_ID)$path %>%
  read.table(sep='\t', header=T, stringsAsFactors=F) %>% 
  dplyr::rename(ID = sample)
 colnames(METADATA_PICARD_TC) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA_PICARD_TC))
 colnames(METADATA_PICARD_TC) = gsub('RnaSeqMetrics__','',colnames(METADATA_PICARD_TC))

METADATA_PICARD_TC <- METADATA_PICARD_TC[, !duplicated(colnames(METADATA_PICARD_TC))]
METADATA_PICARD_CE <- METADATA_PICARD_CE[, !duplicated(colnames(METADATA_PICARD_CE))]
METADATA_PICARD <- dplyr::full_join( METADATA_PICARD_TC, METADATA_PICARD_CE )

# Merge all metadata
METADATA = list(METADATA_TC, METADATA_CE) %>%
  data.table::rbindlist(use.names = T, fill = T) %>%
  dplyr::inner_join(list(METADATA_PICARD_TC, METADATA_PICARD_CE) %>% 
                      data.table::rbindlist(use.names = T, fill = T)) %>%
                      as.data.frame

biomart_results <- read.table(syn_temp$get('syn22087693')$path, header=T, sep='\t', stringsAsFactors = F)

# Duplicate Ensembl Ids are collapsed into a single entry
collapse_duplicate_trans <- function(biomart_results){
  biomart_results %>%
    dplyr::group_by(.data$ensembl_transcript_id) %>%
    dplyr::mutate(hgnc_symbol = paste(.data$hgnc_symbol, collapse = ", ")) %>%
    unique()
}
biomart_results <- as.data.frame( collapse_duplicate_trans(biomart_results), stringsAsFactors = F )
row.names(biomart_results) <- biomart_results$ensembl_transcript_id
biomart_results <- biomart_results[ row.names(COUNT), ]

GENE.PARAM = biomart_results
GENE.GC = dplyr::select(GENE.PARAM, ensembl_transcript_id, percentage_gene_gc_content) %>% unique()
colnames(GENE.GC) = c('ensembl_gene_id', 'percentage_gc_content')

GENE.LEN = dplyr::select(GENE.PARAM, ensembl_transcript_id, length) %>% unique()
colnames(GENE.LEN) = c('ensembl_gene_id', 'gene.length')

SAMPLES.EXCLUDE = c(read.table(syn_temp$get('syn6126119', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name, read.table(syn_temp$get('syn6126114', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name)

```
Following `r length(SAMPLES.EXCLUDE)` samples were removed from analysis as original investigators tagged them as outliers for various reasons as described here syn6126119, syn6126114
Samples removed are `r paste(SAMPLES.EXCLUDE, collapse = ', ')`

### Data preprocessing
```{r preprocess.data, echo=FALSE}
# Filter metadata
METADATA = as_tibble(METADATA, .name_repair = "unique") %>%
  dplyr::filter(ID %in% colnames(COUNT)) %>%
  dplyr::filter(!is.na(RIN)) %>%
  dplyr::filter(!is.na(PCT_INTRONIC_BASES)) %>%
  dplyr::filter(!is.na(AgeAtDeath)) %>%
  dplyr::filter(!is.na(Source)) %>%
  dplyr::filter(!(ID %in% SAMPLES.EXCLUDE)) %>%
  as.data.frame()
# Fix Tissue
METADATA$Tissue = gsub('CER', 'CBE', METADATA$Tissue)
# Fix AgeAtDeath
METADATA = METADATA %>%
  dplyr::mutate(AgeAtDeath = gsub("_or_above", "", AgeAtDeath))
# Fix Diagnosis
METADATA = METADATA %>%
  dplyr::mutate(Source.Diagnosis = Diagnosis) %>%
  dplyr::mutate(Source.Diagnosis = factor(Source.Diagnosis,
                                          levels = c('AD', 'Control', 'Pathologic Aging', 'PSP'),
                                          labels = c('AD', 'CONTROL', 'PATH_AGE', 'PSP'))) %>%
  dplyr::mutate(Diagnosis = gsub("Pathologic Aging", "OTHER", Diagnosis),
                Diagnosis = gsub("PSP", "OTHER", Diagnosis),
                Diagnosis = factor(Diagnosis, 
                                   levels = c("AD", "Control", "OTHER"),
                                   labels = c('AD', 'CONTROL', 'OTHER'))) %>%
  dplyr::mutate(Tissue.Diagnosis = paste(Tissue, Diagnosis, sep = '.'),
                Tissue.SourceDiagnosis = paste(Tissue, Source.Diagnosis, sep = '.'))
# Fix Gender
METADATA = METADATA %>%
  dplyr::mutate(Sex = Gender,
                Sex = factor(Sex, levels = c('F','M'), labels = c('FEMALE','MALE')))
# Get RIN square
METADATA = METADATA %>%
  dplyr::mutate(RIN2 = RIN^2)
# Add apoe4 genotype (0, 1, 2)
METADATA$APOE4 = 0
METADATA$APOE4[METADATA$ApoE %in% c(24, 34)] = 1
METADATA$APOE4[METADATA$ApoE %in% c(44)] = 2
METADATA$Tissue.APOE4 = paste(METADATA$Tissue, METADATA$APOE4, sep = '.')
# Match covariates to expression data
indToRetain = intersect(METADATA$ID, colnames(COUNT))
indRemoved = setdiff(colnames(COUNT), METADATA$ID)
COUNT <- COUNT[, indToRetain]
rownames(METADATA) = METADATA$ID
METADATA = METADATA[indToRetain,]
```
`r dim(COUNT)[2]` samples from `r length(unique(METADATA$Donor_ID))` subjects were obtained from the Mayo cohorts in AMP-AD reprocessed rnaseq project.

Following samples were removed `r indRemoved`

### Predict missing PMI
```{r predict.pmi}
# Get PMI from the gene level data used to predict
GeneLvlMeta_WithPMI <- fread(syn_temp$get('syn20801992')$path, data.table = F)
row.names(GeneLvlMeta_WithPMI) <- GeneLvlMeta_WithPMI$SampleID
METADATA$PMI <- NA
METADATA[row.names(GeneLvlMeta_WithPMI),]$PMI <-  GeneLvlMeta_WithPMI$PMI

```
Fitted R2 values are, (i) with all gx features `r round(R2.allgenes, 3)`, (ii) with small set of inferred gx features `r round(R2.smallgenes,3)`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering}
primaryVariable <- c("Tissue.SourceDiagnosis", "Tissue.Diagnosis", "Tissue.APOE4")
FactorCovariates <- c("Donor_ID", "Source", "Sex", "FLOWCELL", "Tissue.SourceDiagnosis", "Tissue.Diagnosis", "Tissue.APOE4")
ContCovariates <- c("RIN", "RIN2", "AgeAtDeath","PMI", "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
METADATA <- METADATA[ !duplicated(METADATA), ]
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
rownames(COVARIATES) <- METADATA$ID
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, cache=FALSE, fig.width=10, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm'))
```
### Explore metadata
```{r data.explore, cache=FALSE, fig.width = 12, fig.height = 8}
my.theme = theme(legend.position = 'top', axis.text.x = element_text(angle = 90, hjust = 1))
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = RIN)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + my.theme
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(AgeAtDeath))) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeAtDeath') + my.theme
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PMI))) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + my.theme
# Ribosomal bases
p[[4]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_RIBOSOMAL_BASES))) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Fraction of Ribosomal Bases') + my.theme
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTRONIC_BASES))) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + my.theme
# Intergenic bases
p[[6]] = ggplot(COVARIATES, aes(x = Tissue.Diagnosis, y = as.numeric(PCT_INTERGENIC_BASES))) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Intergenic Bases') + my.theme
multiplot(plotlist = p, cols = 3)
```

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Tissue x Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(Tissue.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object

Ensemble2HGNC <- biomart_results[ biomart_results$ensembl_transcript_id %in% row.names(PROCESSED_COUNTS$filteredExprMatrix$genes), ]

summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable

```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Library Normalisation
Library normalisation is performed using cqn (conditional quantile normalisation)
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Outlier Analysis
#### Sample outliers
Outlier analysis is performed before library normalisation with raw cpm counts
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c("11311_CER", "1923_CER", "1923_TCX", "11396_TCX", "11294_TCX", "11408_TCX",
                '1950_TCX', '1950_CER', '1925_TCX', '1957_CER')
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") 
p <- p + facet_grid(Tissue~.+Diagnosis, scales = 'free')
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove) %>%
             droplevels(),
           aes(x = logCPM)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Tissue~.+Diagnosis)
p
indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]
tmp = COVARIATES %>%
  dplyr::group_by(Tissue.Diagnosis) %>%
  dplyr::summarise(count = n()) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.') %>%
  tidyr::spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

Distribution of samples are: `r kable(tmp)`

#### Gene outliers
Assign NA values to genes that are above and below 3 std deviation of its distribution
```{r winsorise.data}
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
CQN.GENE_EXPRESSION$E.no.na = CQN.GENE_EXPRESSION$E
CQN.GENE_EXPRESSION$E.no.na[is.na(CQN.GENE_EXPRESSION$E.no.na)] = 0
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, cache=FALSE, fig.height=8, fig.width=8, results='asis', cache=FALSE}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="top") + facet_grid(Tissue+.~Diagnosis, scales = 'free_y')
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
COVARIATES.tmp = (COVARIATES[,c(FactorCovariates), drop = F])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE, fig.height=10, fig.width=20}
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.')
p = ggplot(tmp1, aes(x = logCPM)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(Tissue+.~Diagnosis, scale = 'free')
p
```

Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(CQN.GENE_EXPRESSION$E.no.na))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=TRUE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                   COVARIATES,
                                                   'NULL design(voom-normalized)', 
                                                   isKeyPlot=TRUE, 
                                                   MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, cache=FALSE, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix.

NOTE:
1. Using a mixed effect model where random effect is chosen as Donor_ID
2. Adding Source, FLOWCELL and Gender a priori to variable selection
3. Primary variable of interest Tissue.Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, results='asis'}

# Primary variable of interest
postAdjustCovars = c('Source', 'FLOWCELL', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects and Donor_ID as random effect'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  cnts = NEW.COUNTS; cnts[is.na(cnts)] = 0
  VOOM.GENE_EXPRESSION = voom(cnts, design=DM1$design, 
                              plot=F)#, na.rm = T)

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$Donor_ID,
                                             method = 'lmer')
  # correlation = list(); correlation$cor = 0.3819258

  if (!is.nan(correlation$cor)){
    # Re-estimate voom weights
    VOOM.GENE_EXPRESSION = voom(cnts, 
                                design=DM1$design, plot=F, 
                                block = COVARIATES$Donor_ID, 
                                correlation = correlation$cor)

    # Fit linear model using new weights and new design
    VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                              design = DM1$design,
                              weights = VOOM.GENE_EXPRESSION$weights,
                              block = COVARIATES$Donor_ID, 
                              correlation = correlation$cor)

    # Residuals after normalisation
    #_# Fix NA coeefcient to stop PCA error
    CQN.GENE_EXPRESSION$E[is.na(CQN.GENE_EXPRESSION$E)] = 0
    VOOM.ADJUSTED.FIT$coefficients[is.na(VOOM.ADJUSTED.FIT$coefficients)] = 0
    RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                  CQN.GENE_EXPRESSION$E)
  } else {
    notEstimable = c(notEstimable, postAdjustCovars[length(postAdjustCovars)])
    postAdjustCovars = postAdjustCovars[1:(length(postAdjustCovars)-1)]
  }

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'Donor_ID', notEstimable))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  expr = RESIDUAL.GENE_EXPRESSION; expr[is.na(expr)] = 0
  residualSigCovars = runPCAandPlotCorrelations(expr, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, notEstimable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')
tmp <- paste('Using following covariates in the final model:', modelStr)
#save.image('Mayo_Image.RData')
```
`r tmp`

### Sanity check
```{r residual.adj, cache=FALSE, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(expr))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, cache=FALSE, fig.height=10, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Tissue.Diagnosis, c('Region','Diagnosis'), sep = '\\.') %>% 
  dplyr::mutate(Region = factor(Region),
                Source = factor(Source))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Source, color=Diagnosis, size=RIN))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Region~., scales = 'free_y')
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, cache=FALSE, fig.height=8, fig.width=12, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,FactorCovariates])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Adjust data with covariates for Network Analysis
Identified covariates are regressed out from the expression matrix for network analysis
```{r network.adjust}
# Get design matrix
DESIGN.NET = getDesignMatrix(COVARIATES[, postAdjustCovars, drop = F], Intercept = F)
DESIGN.NET = DESIGN.NET$design[,linColumnFinder(DESIGN.NET$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.NET.WEIGHTS = voom(cnts, design=DESIGN.NET, plot=F)
# Re-estimate voom weights
VOOM.NET.WEIGHTS = voom(cnts, 
                        design = DESIGN.NET, plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS$E,
                     weights = VOOM.NET.WEIGHTS$weights,
                     design = VOOM.NET.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.NET.GENE_EXPRESSION = residuals.MArrayLM(VOOM.NET.FIT,
                                                  CQN.GENE_EXPRESSION$E)
#save.image('Mayo_Image.RData')

```

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Tissue.Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.2)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E.no.na,
                        MODEL1, MODEL0, n.sv = NUM.SV, 
                        B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(cnts, design=cbind(MODEL1, SURR.VAR), plot=F)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS$E,
                     weights = VOOM.SVA.WEIGHTS$weights,
                     design = VOOM.SVA.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Tissue.Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
#save.image('Mayo_Image.RData')
```

### Differential expression analysis (with Tissue.SourceDiagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Tissue.SourceDiagnosis
1. CBE.AD: Alzheimers Disease samples from Cerebellum region
2. CBE.CONTROL: Control samples from Cerebellum region
3. CBE.PATH_AGE: Pathological ageing samples from Cerebellum region
4. CBE.PSP: Palsy samples from Cerebellum region
5. TCX.AD: Alzheimers Disease samples from temporal cortex region
6. TCX.CONTROL: Control samples from temporal cortex region
7. TCX.PATH_AGE: Pathalogical ageing samples from temporal cortex region
8. TCX.PSP: Palsy samples from temporal cortex region

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=10, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[1], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "Tissue.SourceDiagnosisCBE.AD-Tissue.SourceDiagnosisCBE.CONTROL",
  "Tissue.SourceDiagnosisCBE.PATH_AGE-Tissue.SourceDiagnosisCBE.CONTROL",
  "Tissue.SourceDiagnosisCBE.PSP-Tissue.SourceDiagnosisCBE.CONTROL",
  "Tissue.SourceDiagnosisCBE.AD-Tissue.SourceDiagnosisCBE.PATH_AGE",
  "Tissue.SourceDiagnosisTCX.AD-Tissue.SourceDiagnosisTCX.CONTROL",
  "Tissue.SourceDiagnosisTCX.PATH_AGE-Tissue.SourceDiagnosisTCX.CONTROL",
  "Tissue.SourceDiagnosisTCX.PSP-Tissue.SourceDiagnosisTCX.CONTROL",
  "Tissue.SourceDiagnosisTCX.AD-Tissue.SourceDiagnosisTCX.PATH_AGE"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)

#_# Reformat Gene Param for combination:
GENE.PARAM <- GENE.PARAM[ row.names(DE[[1]]) , ]
colnames(GENE.PARAM)[ colnames(GENE.PARAM) == 'ensembl_transcript_id' ] <- 'ensembl_gene_id'

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.SourceDiagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue.ref','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue.ag','against'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue.ref) %>%
#  group_by(Tissue.ref, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Tissue.ref+.~Comparison, scales = 'fixed')
p
all.diff.exp = list(SourceDiagnosis = DE)
all.fit = list(SourceDiagnosis = FIT)
```

### Differential expression analysis (with Tissue.Diagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Tissue.Diagnosis
1. CER.AD: Alzheimers disease samples from cerebellum region
2. CER.CONTROL: Control samples from cerebellum region
3. CER.OTHER: Progressive supranuclear palsy and pathological aging samples from Cerebellum region
4. TCX.AD: Alzheimers disease samples from temporal cortex region
5. TCX.CONTROL: Control samples from temporal cortex region
6. TCX.OTHER: Progressive supranuclear palsy and pathological aging samples from temporal cortex region

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp.1, fig.height=10, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "Tissue.DiagnosisCBE.AD-Tissue.DiagnosisCBE.CONTROL",
  "Tissue.DiagnosisCBE.AD-Tissue.DiagnosisCBE.OTHER",
  "Tissue.DiagnosisCBE.OTHER-Tissue.DiagnosisCBE.CONTROL",
  "Tissue.DiagnosisTCX.AD-Tissue.DiagnosisTCX.CONTROL",
  "Tissue.DiagnosisTCX.AD-Tissue.DiagnosisTCX.OTHER",
  "Tissue.DiagnosisTCX.OTHER-Tissue.DiagnosisTCX.CONTROL"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.Diagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue.ref','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue.ag','against'), sep = '\\.') %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue.ref) %>%
#  group_by(Tissue.ref, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Tissue.ref+.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis = DE))
all.fit = c(all.fit, list(Diagnosis = FIT))
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=10, fig.width=15}
pl = list()
pl[[1]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = log10(length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(Tissue.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')
pl[[2]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = percentage_gene_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gene_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(Tissue.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')
pl[[3]] = ggplot(DE %>%
                   dplyr::filter(Comparison == 'AD-CONTROL'), 
                 aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(Tissue.ref+.~Comparison) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')
multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (with Tissue.APOE as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Tissue.Diagnosis
1. CBE.0: No APOE4 allele in cerebellum
2. CBE.1: 1 APOE4 allele in cerebellum
3. CBE.2: 2 APOE4 allele in cerebellum
4. TCX.0: No APOE4 allele in temporal cortex
5. TCX.1: 1 APOE4 allele in temporal cortex
6. TCX.2: 2 APOE4 allele in temporal cortex

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp.2, fig.height=10, fig.width=18}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[3], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "Tissue.APOE4CBE.2-Tissue.APOE4CBE.0",
  "Tissue.APOE4CBE.2-Tissue.APOE4CBE.1",
  "Tissue.APOE4CBE.1-Tissue.APOE4CBE.0",
  "Tissue.APOE4TCX.2-Tissue.APOE4TCX.0",
  "Tissue.APOE4TCX.2-Tissue.APOE4TCX.1",
  "Tissue.APOE4TCX.1-Tissue.APOE4TCX.0"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.APOE4','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue.ref','reference'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue.ag','against'), sep = '\\.') %>%
  dplyr::mutate(reference = paste0('APOE4',reference),
                against = paste0('APOE4',against)) %>%
  tidyr::unite(Comparison, reference, against, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue.ref) %>%
#  group_by(Tissue.ref, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Tissue.ref+.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(APOE4 = DE))
all.fit = c(all.fit, list(APOE4 = FIT))
```

### Differential expression analysis (with Tissue.Diagnosis x Sex as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp2, fig.height=5, fig.width=18}
# Get design matrix
COVARIATES$Tissue.Diagnosis.Sex = paste(COVARIATES$Tissue.Diagnosis, COVARIATES$Sex, sep = '.') %>%
  factor()
DESIGN = getDesignMatrix(COVARIATES[, c('Tissue.Diagnosis.Sex', setdiff(postAdjustCovars, 'Sex')), drop = F], 
                         Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "Tissue.Diagnosis.SexCBE.AD.FEMALE-Tissue.Diagnosis.SexCBE.CONTROL.FEMALE",
  "Tissue.Diagnosis.SexCBE.AD.MALE-Tissue.Diagnosis.SexCBE.CONTROL.MALE",

  "Tissue.Diagnosis.SexTCX.AD.FEMALE-Tissue.Diagnosis.SexTCX.CONTROL.FEMALE",
  "Tissue.Diagnosis.SexTCX.AD.MALE-Tissue.Diagnosis.SexTCX.CONTROL.MALE"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.Diagnosis.Sex','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue.ref','Dx.ref','Sex.ref'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue.ag','Dx.ag','Sex.ag'), sep = '\\.') %>%
  tidyr::unite(Comparison, Dx.ref, Dx.ag, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue.ref, Sex.ref) %>%
#  group_by(Tissue.ref, Sex.ref, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Comparison+.~Tissue.ref+Sex.ref, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis.Sex = DE))
all.fit = c(all.fit, list(Diagnosis.Sex = FIT))
```

### Differential expression analysis (with Tissue.Diagnosis x AgeAtDeath as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp3, fig.height=5, fig.width=18, cache = FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('Tissue.Diagnosis', setdiff(postAdjustCovars, 'AgeAtDeath')), drop = F], 
                         Intercept = F)
DESIGN$design[,grep('Tissue.Diagnosis',colnames(DESIGN$design))] = DESIGN$design[,grep('Tissue.Diagnosis',colnames(DESIGN$design))] * COVARIATES[,'AgeAtDeath']
colnames(DESIGN$design)[grep('Tissue.Diagnosis',colnames(DESIGN$design))] = 
  paste(colnames(DESIGN$design)[grep('Tissue.Diagnosis',colnames(DESIGN$design))],'AOD', sep = '.')
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Fit contrast
contrast = makeContrasts(contrasts=c(
  "Tissue.DiagnosisCBE.AD.AOD-Tissue.DiagnosisCBE.CONTROL.AOD",
  "Tissue.DiagnosisTCX.AD.AOD-Tissue.DiagnosisTCX.CONTROL.AOD"),
  levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::mutate(Comparison = gsub('Tissue.Diagnosis','',Comparison),
                Study = 'MAYO',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction)) %>%
  tidyr::separate(Comparison, c('from.state','to.state'), sep = '-') %>%
  tidyr::separate(from.state, c('Tissue.ref','Dx.ref','AOD.ref'), sep = '\\.') %>%
  tidyr::separate(to.state, c('Tissue.ag','Dx.ag','AOD.ag'), sep = '\\.') %>%
  tidyr::unite(Comparison, Dx.ref, Dx.ag, sep = '-') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique())
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
#tmp = DE %>%
#  dplyr::select(ensembl_gene_id, Comparison, Direction, Tissue.ref, AOD.ref) %>%
#  group_by(Tissue.ref, AOD.ref, Comparison, Direction) %>%
#  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
#  spread(Direction, FDR_0_05_FC_1.2) 
#kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(Comparison+.~Tissue.ref+AOD.ref, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22092490';
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective protien coding transcript level counts with CQN normalisation ( TCX, CBE )';
thisFileName <- 'Mayo_TranscriptNorm.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts/',thisFileName))
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh38 with CQN normalisation (CBE, TCX)';
CODE <- syn_temp$store(synapseclient$Folder(name = "Mayo", parentId = parentId))
#Set Used SynIDs For Provenance
Syns_Used <- c( "syn10499227","syn10507722","syn3817650","syn5223705","syn8698214","syn8698211","syn22087693","syn6126119","syn6126114","syn20801992")

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= c('TCX', 'CBE'), 
  study = 'MAYO', 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh38'
)
# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'MAYO_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'MAYO_SVA_Diagnosis_Covariates.tsv', row.names = F, sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_SVA_Diagnosis_Covariates.tsv', name = 'Covariates', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'covariates'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_Counts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_Counts.tsv', name = 'Counts (filtered raw)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store Estimated counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_Counts.tsv', name = 'Counts (estimated)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'estimatedCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_logCPM.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredLCPM'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_offset.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_offset.tsv', name = 'Gene length and GC content offset', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'offset'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_Design.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_Design.tsv', name = 'Design Matrix', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'designMatrix'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_SurVar.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_SurVar.tsv', name = 'Surrogate Variables', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'surrogateVariables'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'MAYO_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_Diagnosis_Regressed_netResidualExpression.tsv', name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store differential expression results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'MAYO_CBE_TCX_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='MAYO_CBE_TCX_DiffExpression.tsv', name = 'Differential Expression Results (Tissue.Diagnosis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'diffExp'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

stopCluster(cl)
```

```{r DelFiles, echo=T, results='hide'}
system('rm *.tsv')
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp  <- synapseclient$Synapse()
syn_temp$login()
setwd("~/AD_TargetRank/code/")
source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
createAndKnitToFolderEntityClient(file = "Mayo_TranscriptNorm.Rmd",
                                          parentId ="syn22092490",
                                          folderName = 'Mayo')
```

### 2020-05-23 code/prepscripts/Mayo_TranscriptNorm.Rmd
Trans DE MSBB

@@
 activityName = 'Covariate and Diagnosis Regression';
-activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with GRCh38 with CQN normalisation (CBE, TCX)';
+activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with CQN normalisation (CBE, TCX)';
 CODE <- syn_temp$store(synapseclient$Folder(name = "Mayo", parentId = parentId))

### 2020-05-23 code/prepscripts/MSBB_TranscriptNorm.Rmd
Trans DE MSBB
File Paths Fix

@@
 # Source modified utility functions from limma and sva package
-source('../utilityFunctions/parallelDuplicateCorrelation.R')
-source('../utilityFunctions/irwsva.build.R')
-source('../utilityFunctions/f.pvalue.R')
+source('~/AD_TargetRank/utilityFunctions/parallelDuplicateCorrelation.R')
+source('~/AD_TargetRank/utilityFunctions/irwsva.build.R')
+source('~/AD_TargetRank/utilityFunctions/f.pvalue.R')
 knitr::opts_chunk$set(
@@
 
-#_#stopCluster(cl)
+stopCluster(cl)
 ```
@@
 ```{r DelFiles, echo=T, results='hide'}
-#system('rm *.tsv')
+system('rm *.tsv')
 ```

### 2020-05-23 code/prepscripts/Mayo_TranscriptNorm.Rmd
File Paths Fix

@@
 # Source modified utility functions from limma and sva package
-source('../../utilityFunctions/parallelDuplicateCorrelation.R')
-source('../../utilityFunctions/irwsva.build.R')
-source('../../utilityFunctions/f.pvalue.R')
+source('~/AD_TargetRank/utilityFunctions/parallelDuplicateCorrelation.R')
+source('~/AD_TargetRank/utilityFunctions/irwsva.build.R')
+source('~/AD_TargetRank/utilityFunctions/f.pvalue.R')
 knitr::opts_chunk$set(

### 2020-05-23 code/prepscripts/MSBB_TranscriptNorm.Rmd
File Paths Fix

@@
 Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`
-
-Distribution of samples are: `r kable(tmp)`
 
@@
 source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
-createAndKnitToFolderEntityClient(file = "MSBB_TranscriptNorm.Rmd",
+createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/prepscripts/MSBB_TranscriptNorm.Rmd",
                                           parentId ="syn22092490",

### 2020-05-23 code/prepscripts/Mayo_TranscriptNorm.Rmd
File Paths Fix

@@
 
-Distribution of samples are: `r kable(tmp)`
-
 #### Gene outliers
@@
 source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
-createAndKnitToFolderEntityClient(file = "Mayo_TranscriptNorm.Rmd",
+createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/prepscripts/Mayo_TranscriptNorm.Rmd",
                                           parentId ="syn22092490",

### 2020-05-26 code/prepscripts/ROSMAP_TranscriptNorm.Rmd
Rosmap DE

---
title: Covariate and differential expression analysis of ROSMAP reprocessed counts
author: "Thanneer Perumal"
editor_options:
  chunk_output_type: console
---
Date of analysis update: "`r date()`"

### Load Libraries
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
## Load required libraries
library(CovariateAnalysis) # get the package from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(plyr)
library(tidyverse)
library(psych)
library(limma)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
library(knitr)
library(doParallel)
library(foreach)
cl = makeCluster(14)
registerDoParallel(cl)
options(xtable.type="html")
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Data download
Obtain count matrix and metadata from synapse
```{r download.data}
# Download expression data
COUNT_ID <- 'syn10507739';
ALL_USED_IDs <- COUNT_ID
COUNT_OBJ <- syn_temp$get(COUNT_ID)
COUNT <- read.table(COUNT_OBJ$path, header=T, sep='\t', check.names = F, row.names = 1)
COUNT[,grep('150_120419', colnames(COUNT))[2]] = NULL
# Convert rownames of counts from tracking id to ensemble gene id
tmp = data.frame(Gene.ID = rownames(COUNT)) %>%
  dplyr::mutate(ID = Gene.ID) %>%
  tidyr::separate(ID, c('ensembl_gene_id', 'position'), sep = '\\.')
rownames(tmp) = tmp$Gene.ID
rownames(COUNT) = tmp[rownames(COUNT), 'ensembl_gene_id']
# Get clinical metadata
METADATA.CLINICAL_ID <- 'syn3191087'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID
METADATA.CLINICAL_OBJ <- syn_temp$get(METADATA.CLINICAL_ID, version = 3)
METADATA.CLINICAL <- read.table(METADATA.CLINICAL_OBJ$path,sep=',',header=T)
# Get clinical metadata with uncensored ages
METADATA.CLINICAL_ID1 <- 'syn7116000'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.CLINICAL_ID1
METADATA.CLINICAL_OBJ1 <- syn_temp$get(METADATA.CLINICAL_ID1, version = 1)
METADATA.CLINICAL1 <- read.table(METADATA.CLINICAL_OBJ1$path,sep=',',header=T)
# Get technical covariates
METADATA.TECH_ID <- 'syn4300313'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.TECH_ID
METADATA.TECH_OBJ <- syn_temp$get(METADATA.TECH_ID, version = 1)
METADATA.TECH <- read.table(METADATA.TECH_OBJ$path,sep='\t',header=T)
# Get picard metrics from synapse
METADATA.PICARD_ID <- 'syn8698240';
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA.PICARD_ID
METADATA.PICARD <- syn_temp$get(METADATA.PICARD_ID)$path %>%
  data.table::fread() %>%
  dplyr::rename(Sampleid = sample)
# Fix error in technical covariates data
KEY_ID <- 'syn3382527'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = KEY_ID  
KEY <- syn_temp$get(KEY_ID, version = 7)$path %>%
  read.csv %>% 
  dplyr::filter(!is.na(rnaseq_id)) %>%
  dplyr::select(projid, rnaseq_id) %>%
  tidyr::separate(rnaseq_id, c('a','b','batch'), sep = '_') %>% 
  unite(Sampleid, a, b) %>%
  dplyr::select(-batch) %>%
  unique
# Match technical and clinical covariates
METADATA <- METADATA.TECH %>%
  dplyr::left_join(METADATA.PICARD) %>%
  dplyr::select(-projid) %>%
  dplyr::left_join(KEY) %>%
  dplyr::left_join(METADATA.CLINICAL) %>%
  dplyr::select(-age_first_ad_dx, -age_death, -age_at_visit_max) %>%
  dplyr::left_join(METADATA.CLINICAL1)
# Pick higher quality RIN batch for sample 492_120515
METADATA <- METADATA %>%
  dplyr::group_by(Sampleid) %>%
  dplyr::top_n(1, RINcontinuous)
colnames(METADATA) = gsub('AlignmentSummaryMetrics__','',colnames(METADATA))
colnames(METADATA) = gsub('RnaSeqMetrics__','',colnames(METADATA))
# Get gene specific parameters from synapse

ALL_USED_IDs[length(ALL_USED_IDs)+1] = 'syn22087693'  
biomart_results <- read.table(syn_temp$get('syn22087693')$path, header=T, sep='\t', stringsAsFactors = F)

# Duplicate Ensembl Ids are collapsed into a single entry
collapse_duplicate_trans <- function(biomart_results){
  biomart_results %>%
    dplyr::group_by(.data$ensembl_transcript_id) %>%
    dplyr::mutate(hgnc_symbol = paste(.data$hgnc_symbol, collapse = ", ")) %>%
    unique()
}
biomart_results <- as.data.frame( collapse_duplicate_trans(biomart_results), stringsAsFactors = F )
row.names(biomart_results) <- biomart_results$ensembl_transcript_id
biomart_results <- biomart_results[ row.names(COUNT), ]

GENE.PARAM = biomart_results
GENE.GC = dplyr::select(GENE.PARAM, ensembl_transcript_id, percentage_gene_gc_content) %>% unique()
colnames(GENE.GC) = c('ensembl_gene_id', 'percentage_gc_content')

GENE.LEN = dplyr::select(GENE.PARAM, ensembl_transcript_id, length) %>% unique()
colnames(GENE.LEN) = c('ensembl_gene_id', 'gene.length')

SAMPLES.EXCLUDE = c(read.table(syn_temp$get('syn6126119', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name, read.table(syn_temp$get('syn6126114', version = 3)$path, sep = '\t', header = T, stringsAsFactors = F)$Sample.Name)
```

### Data preprocessing
```{r preprocess.data, cache=TRUE}
# Remove samples with no cogdx, RIN, PMI scores and age_death
names(METADATA)[13] <- "PF_ALIGNED_BASES_alt"
METADATA <- METADATA %>%
  ungroup %>%
  dplyr::filter(Sampleid %in% colnames(COUNT)) %>%
  dplyr::filter(!is.na(cogdx), !is.na(braaksc), !is.na(ceradsc)) %>%
  dplyr::filter(!is.na(RINcontinuous)) %>%
  dplyr::filter(!is.na(PCT_INTRONIC_BASES)) %>%
  dplyr::filter(!is.na(pmi)) %>%
  dplyr::filter(!is.na(age_death)) %>%
  as.data.frame()
# Add harmonised case-control status
METADATA$Diagnosis = 'OTHER'
METADATA$Diagnosis[METADATA$cogdx == 1 & METADATA$braaksc <= 3 & METADATA$ceradsc >= 3] = 'CONTROL'
METADATA$Diagnosis[METADATA$cogdx == 4 & METADATA$braaksc >= 4 & METADATA$ceradsc <= 2] = 'AD'
# Add sex variable 
METADATA$Sex = 'FEMALE'
METADATA$Sex[METADATA$msex == 1] = 'MALE'
# Add apoe4 genotype (0, 1, 2)
METADATA$APOE4 = 0
METADATA$APOE4[METADATA$apoe_genotype %in% c(24, 34)] = 1
METADATA$APOE4[METADATA$apoe_genotype %in% c(44)] = 2
# METADATA$APOE4[is.na(METADATA$apoe_genotype)] = NA
# Get square of RIN
METADATA$RINcontinuous2 = METADATA$RINcontinuous^2
# Match covariates to expression data
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))
removedIDs = setdiff(colnames(COUNT), METADATA$Sampleid)
COUNT = COUNT[,indToRetain]
rownames(METADATA) = METADATA$Sampleid
METADATA = METADATA[indToRetain,]
```
Dorsolateral prefrontal cortex of `r dim(COUNT)[2]` subjects from the ROS and MAP cohorts are used for the analysis. Following sample are removed due to missing metadata `r paste(removedIDs, collapse = ',')`

### Covariate clustering
Determine relationship between covariates
```{r covariates.clustering, cache=FALSE}
primaryVariable <- c("cogdx", "Diagnosis", "APOE4")
FactorCovariates <- c("Batch", "Sex", "race", "spanish", "cogdx", "Diagnosis", "APOE4")
ContCovariates <- c("RINcontinuous", "RINcontinuous2", "age_death", "pmi", "educ", 
                    "PCT_PF_READS_ALIGNED", "PCT_CODING_BASES",
                    "PCT_INTERGENIC_BASES", "PCT_INTRONIC_BASES", 
                    "PCT_RIBOSOMAL_BASES")
# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES <- data.frame(lapply(COVARIATES,function(x){x <- sapply(x,function(y){str_replace_all(as.character(y),'\\+','')})}))
rownames(COVARIATES) <- METADATA$Sampleid
# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.character)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], as.numeric)
```

Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, cache=FALSE, fig.width=9, fig.height=6}
COVARIATES.CORRELATION = CovariateAnalysis::getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left', padding  = unit(c(18,2,2,18), 'mm')) 
```

### Explore metadata
```{r data.explore, cache=FALSE, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')
# AgeAtDeath
p[[2]] = ggplot(COVARIATES, aes(x = Diagnosis, y = age_death)) + geom_boxplot()
p[[2]] = p[[2]] + ggtitle('AgeOfDeath') + theme(legend.position = 'top')
# PMI
p[[3]] = ggplot(COVARIATES, aes(x = Diagnosis, y = pmi)) + geom_boxplot()
p[[3]] = p[[3]] + ggtitle('PMI') + theme(legend.position = 'top')
# Education
p[[4]] = ggplot(COVARIATES, aes(x = Diagnosis, y = educ)) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Education') + theme(legend.position = 'top')
# Intronic bases
p[[5]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_INTRONIC_BASES)) + geom_boxplot()
p[[5]] = p[[5]] + ggtitle('Fraction Intronic Bases') + theme(legend.position = 'top')
# Ribosomal bases
p[[6]] = ggplot(COVARIATES, aes(x = Diagnosis, y = PCT_RIBOSOMAL_BASES)) + geom_boxplot()
p[[6]] = p[[6]] + ggtitle('Fraction Ribosomal Bases') + theme(legend.position = 'top')
multiplot(plotlist = p, cols = 2)
```

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object

Ensemble2HGNC <- biomart_results[ biomart_results$ensembl_transcript_id %in% row.names(PROCESSED_COUNTS$filteredExprMatrix$genes), ]

summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

### Library Normalisation
Library normalisation is performed using cqn (conditional quantile normalisation)
```{r cqn}
# Compute offset for gene length and gc content
CQN.GENE_EXPRESSION = cqn(PROCESSED_COUNTS$filteredExprMatrix$counts, 
                          x = GENE.GC[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'percentage_gc_content'],
                          lengths = GENE.LEN[PROCESSED_COUNTS$filteredExprMatrix$genes$genes, 'gene.length'],
                          lengthMethod = "smooth", 
                          verbose = FALSE)
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$y + CQN.GENE_EXPRESSION$offset
```

### Outlier Analysis
#### Sample outliers
Outlier analysis is performed before library normalisation with raw cpm counts
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c('380_120503', '500_120515')
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- dplyr::left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  dplyr::mutate(label = SampleID)
plotdata$label[!(plotdata$SampleID %in% indToRemove)] = ''
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  dplyr::left_join(COVARIATES %>%
                     rownameToFirstColumn('SampleID'))
p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% indToRemove),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top')
p
indToRetain = setdiff(colnames(PROCESSED_COUNTS$filteredExprMatrix$counts), indToRemove)
PROCESSED_COUNTS$filteredExprMatrix$counts = PROCESSED_COUNTS$filteredExprMatrix$counts[,indToRetain]
CQN.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E[,indToRetain]
COVARIATES = COVARIATES[indToRetain,]
tmp = COVARIATES %>%
  dplyr::group_by(Diagnosis, cogdx) %>%
  dplyr::summarise(count = n()) %>%
  tidyr::spread(Diagnosis, count)
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

Based on the expression pattern following samples were tagged as outliers: `r paste(indToRemove, collapse = ', ')`

#### Gene outliers
Assign NA values to genes that are above and below 3 std deviation of its distribution
```{r winsorise.data}
# Set gene counts in specific samples that are deviating 3 sd from other samples to 3SD limit
LOG.CPM = apply(CQN.GENE_EXPRESSION$E, 1, function(x){
  mn = mean(x, na.rm = T)
  std.dev = sd(x, na.rm = T)

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
CQN.GENE_EXPRESSION$E.no.na = CQN.GENE_EXPRESSION$E
CQN.GENE_EXPRESSION$E.no.na[is.na(CQN.GENE_EXPRESSION$E.no.na)] = 0
LIB.SIZE = colSums(PROCESSED_COUNTS$filteredExprMatrix$counts)
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

### Sample clustering
PCA based clustering of samples
```{r decompse.normalise.data, cache=FALSE, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(CQN.GENE_EXPRESSION$E.no.na, scale.=T, center = T)
# Plot first 2 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=cogdx, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```

Tree based clustering of samples
```{r decompse.normalise.data.1, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Batch", "Sex", "Diagnosis")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(CQN.GENE_EXPRESSION$E.no.na)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Distribution of samples (log cpm)
```{r lcpm.dist, cache=FALSE, fig.height=10, fig.width=20}
# Plot abberent distribution of logcpm counts
tmp1 = CQN.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Gene.ID') %>%
  tidyr::gather(SampleID, logCPM, -Gene.ID) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))
p = ggplot(tmp1, aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'NONE') + facet_grid(.~Diagnosis, scale = 'free')
p
```

Coexpression of genes 
```{r coexp1, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(CQN.GENE_EXPRESSION$E.no.na))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates are used to find significant covariates
```{r preadj.covariates, cache=FALSE}
# Find correlation between PC's of gene expression with covariates
preAdjustedSigCovars = runPCAandPlotCorrelations(CQN.GENE_EXPRESSION$E.no.na, 
                                                 COVARIATES,
                                                 'NULL design(voom-normalized)', 
                                                 isKeyPlot=TRUE, 
                                                 MIN_PVE_PCT_PC = 1)
```

Significant covariates to adjust at FDR 0.1 are `r preAdjustedSigCovars$significantCovars`
```{r preadj.covariates.plot, cache=FALSE, fig.width=12, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```

### Normalisation (iterative design)
Since many covariates are correlated, re-normalising and re-adjusting COUNTS with an iterative design matrix
1. Adding Batch and Sex a priori to variable selection
2. Primary variable of interest Diagnosis is excluded from the pool of available covariates for selection
```{r iterative.norm, cache=FALSE, results='asis'}
# Primary variable of interest
postAdjustCovars = c('Batch', 'Sex');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
loopCount = 0 
while(length(residualSigCovars$significantCovars)!=0 && loopCount <= 20){
  writeLines(paste('Using following covariates in the model:',
                   paste(postAdjustCovars, collapse=', '),
                   'as fixed effects'))

  # Post adjusted design matrix
  DM1 = getDesignMatrix(COVARIATES[,postAdjustCovars,drop=F],Intercept = F)
  DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

  # Estimate voom weights for dispersion control
  cnts = NEW.COUNTS
  cnts[is.na(cnts)] = 0
  VOOM.GENE_EXPRESSION = voom(cnts, 
                              design=DM1$design, 
                              plot=F)#, na.rm = T)

  # Fit linear model using new weights and new design
  VOOM.ADJUSTED.FIT = lmFit(CQN.GENE_EXPRESSION$E,
                            design = DM1$design,
                            weights = VOOM.GENE_EXPRESSION$weights)

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                CQN.GENE_EXPRESSION$E)

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates), c(postAdjustCovars, primaryVariable))

  # Find PC of residual gene expression and significant covariates that are highly correlated with PCs
  expr = RESIDUAL.GENE_EXPRESSION
  expr[is.na(expr)] = 0
  residualSigCovars = runPCAandPlotCorrelations(expr, 
                                                COVARIATES[, residCovars, drop=F], 
                                                'adjusted design(voom-normalized)',
                                                isKeyPlot=TRUE)

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

  postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects)))
  loopCount = loopCount + 1
}
modelStr <- paste(paste(gsub('_','\\\\_',postAdjustCovars), collapse=', '),
                  'as fixed effects')
tmp <- paste('Using following covariates in the final model:', modelStr)
```
`r tmp`

### Sanity check
```{r residual.adj, cache=FALSE, fig.width=12, fig.height=8}
# Find PC of residual gene expression and significant covariates that are highly correlated with PCs
residualSigCovars = runPCAandPlotCorrelations(expr, 
                                              COVARIATES,
                                              'adjusted design(voom-normalized)',
                                              isKeyPlot=TRUE)
residualSigCovars[["PC_res"]][[2]]$plotData
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
#cr = cor(t(expr))
#hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
PCA of residual data
```{r decompse.normalise.data2.1, cache=FALSE, fig.height=6, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)
# Plot first 4 PCs
plotdata <- data.frame(SampleID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2])
plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID'))
p <- ggplot(plotdata, aes(x=PC1, y=PC2)) 
p <- p + geom_point(aes(color=Batch, shape=Diagnosis, size=RINcontinuous))
p <- p + theme_bw() + theme(legend.position="right")
p
```
Tree based clustering of residual data
```{r decompse.normalise.data2.2, cache=FALSE, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
svCOVARIATES.tmp = data.matrix(COVARIATES[,c('Batch', 'Sex', primaryVariable)])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
WGCNA::plotDendroAndColors(tree, 
                           colors = cols, 
                           dendroLabels = FALSE, 
                           abHeight = 0.80, 
                           main = "Sample dendrogram",
                           groupLabels = colnames(COVARIATES.tmp))
```

### Adjust data with covariates for Network Analysis
Identified covariates are regressed out from the expression matrix for network analysis
```{r network.adjust}
# Get design matrix
DESIGN.NET = getDesignMatrix(COVARIATES[, postAdjustCovars, drop = F], Intercept = F)
DESIGN.NET = DESIGN.NET$design[,linColumnFinder(DESIGN.NET$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.NET.WEIGHTS = voom(cnts, design=DESIGN.NET, plot=F)
# Re-estimate voom weights
VOOM.NET.WEIGHTS = voom(cnts, 
                        design = DESIGN.NET, plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS$E,
                     weights = VOOM.NET.WEIGHTS$weights,
                     design = VOOM.NET.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.NET.GENE_EXPRESSION = residuals.MArrayLM(VOOM.NET.FIT,
                                                  CQN.GENE_EXPRESSION$E)
#save.image('Mayo_Image.RData')

```

### SVA Adjustments for eQTL analysis
Conditioned on primary variable (Tissue.Diagnosis) and identified covariates estimate surrogate variables using SVA package
```{r sva.adjust}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Re-estimate voom weights
VOOM.WEIGHTS = voom(cnts,
                    design=DESIGN$design, plot=F,
                    block = COVARIATES$Donor_ID, 
                    correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)
# Get (null) design matrix
MODEL0 = DESIGN$design[,-grep(primaryVariable[2],colnames(DESIGN$design))] # Get null model by removing variable of interest
MODEL0 = MODEL0[,linColumnFinder(MODEL0)$indepCols]
MODEL1 = DESIGN$design
MODEL1 = MODEL1[,linColumnFinder(MODEL1)$indepCols]
# Get residuals from differential expression model
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(FIT, CQN.GENE_EXPRESSION$E)
# Compute actual variance of all principal components
expr = RESIDUAL.SVA.GENE_EXPRESSION
expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)
# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])
# Plot variance components
var.comp = data.frame(component = 1:length(actual.var), 
                      residual = actual.var, 
                      permuted.residual = permuted.var) %>%
  tidyr::gather(data, value, -component)
p = ggplot(var.comp %>% filter(component <= round(NUM.SV*1.2)), aes(x = component, y = value, color = data)) + geom_point()
p = p + geom_hline(yintercept=permuted.var[1], linetype = 'dashed') + xlab('Singular Dimension Index')
p = p + geom_vline(xintercept=NUM.SV, linetype = 'dashed') + ylab('Fraction of Variance Explained')
p = p + theme(legend.position = c(0.8, 0.8), legend.title = element_blank())
p
# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E.no.na,
                        MODEL1, MODEL0, n.sv = NUM.SV, 
                        B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)
# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(cnts, design=cbind(MODEL1, SURR.VAR), plot=F)
# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Donor_ID, 
                        correlation = correlation$cor)
# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS$E,
                     weights = VOOM.SVA.WEIGHTS$weights,
                     design = VOOM.SVA.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)
# Residuals after normalisation
RESIDUAL.SVA.GENE_EXPRESSION = residuals.MArrayLM(VOOM.SVA.FIT,
                                                  CQN.GENE_EXPRESSION$E)
# Add variable of interest back to the residuals
varsToAddIn = grep("Tissue.Diagnosis", colnames(MODEL1), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])

```

### Differential expression analysis (with cogdx as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of cogdx scores
1. NCI, No cognitive impairment (No impaired domains)
2. MCI, Mild cognitive impairment (One impaired domain) and NO other cause of CI
3. MCI, Mild cognitive impairment (One impaired domain) AND another cause of CI
4. AD, Alzheimer's disease and NO other cause of CI (NINCDS PROB AD)
5. AD, Alzheimer's disease AND another cause of CI (NINCDS POSS AD)
6. Other dementia. Other primary cause of dementia

Genes that are differentially expressed at an FDR <= 0.05 are
```{r diff.exp, fig.height=6, fig.width=18, cache=FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[1], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Fit contrast
contrast = makeContrasts(contrasts=c("cogdx2-cogdx1",
                                     "cogdx4-cogdx1",
                                     "cogdx4-cogdx2"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:3, function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)

#_# Reformat Gene Param for combination:
GENE.PARAM <- GENE.PARAM[ row.names(DE[[1]]) , ]
colnames(GENE.PARAM)[ colnames(GENE.PARAM) == 'ensembl_transcript_id' ] <- 'ensembl_gene_id'

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  dplyr::left_join(GENE.PARAM %>%
                     dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
                     unique()) %>%
  dplyr::mutate(Study = 'ROSMAP', Region = 'DLPFC',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
writeLines('Differentially expressed genes at an FDR 0.05 and logFC 1.2')
tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
all.diff.exp = list(cogdx = DE)
all.fit = list(cogdx = FIT)
```

### Differential expression analysis (with Diagnosis as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of Diagnosis
1. AD: cogdx = 4, braaksc >= 4 and ceradsc <= 2
2. CONTROL: cogdx = 1, braaksc <= 3 and ceradsc >= 3
3. OTHER: All the other samples

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp1, fig.height=6, fig.width=18, cache=FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[2], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Fit contrast
contrast = makeContrasts(contrasts=c("DiagnosisAD-DiagnosisCONTROL",
                                     "SexFEMALE"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)

names(DE) = c('AD-CONTROL','FEMALE-MALE')
#DE$`AD-CONTROL`$Comparison <- 'AD-CONTROL'
#DE$`FEMALE-MALE`$Comparison <- 'FEMALE-MALE'

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique()) %>%
  dplyr::mutate(Study = 'ROSMAP', Region = 'DLPFC',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
writeLines('Differentially expressed genes at an FDR 0.05 and logFC 1.2')
tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)
p = DE %>% dplyr::filter(Comparison == 'AD-CONTROL') %>% 
  ggplot(aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis = DE))
all.fit = c(all.fit, list(Diagnosis = FIT))
```

### Associate differential expression results with gc content, gene length and average expression
```{r associate.de, fig.height=10, fig.width=15}
pl = list()
DE = DE %>% dplyr::filter(Comparison == 'AD-CONTROL')
pl[[1]] = ggplot(DE, aes(x = log10(length), y = logFC, color = Direction)) + geom_point() 
pl[[1]] = pl[[1]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = log10(length), y = logFC))
pl[[1]] = pl[[1]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[1]] = pl[[1]] + theme(legend.position = 'top')
pl[[2]] = ggplot(DE, aes(x = percentage_gene_gc_content, y = logFC, color = Direction)) + geom_point()
pl[[2]] = pl[[2]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = percentage_gene_gc_content, y = logFC))
pl[[2]] = pl[[2]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[2]] = pl[[2]] + theme(legend.position = 'top')
pl[[3]] = ggplot(DE, aes(x = AveExpr, y = logFC, color = Direction)) + geom_point()
pl[[3]] = pl[[3]] + geom_smooth(method = 'loess', inherit.aes = FALSE, aes(x = AveExpr, y = logFC))
pl[[3]] = pl[[3]] + facet_grid(Comparison~.) + scale_color_manual(values = c('green','grey','red'))
pl[[3]] = pl[[3]] + theme(legend.position = 'top')
multiplot(plotlist = pl, cols = 3)
```

### Differential expression analysis (with APOE status as primary variable)
Differential expression is performed on the primary variable by controlling for covariates identified above

Interpretation of apoe_genotype
1. 0: No apoe4 allele
2. 1: 1 apoe4 allele
3. 2: 2 apoe4 allele

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp2, fig.height=6, fig.width=18, cache=FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c(primaryVariable[3], postAdjustCovars), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
DESIGN$design = na.omit(DESIGN$design)
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts[,rownames(DESIGN$design)], design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E[,rownames(DESIGN$design)],
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Fit contrast
contrast = makeContrasts(contrasts=c("APOE42-APOE40", "APOE42-APOE41", "APOE41-APOE40"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = colnames(contrast)
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique()) %>%
  dplyr::mutate(Study = 'ROSMAP', Region = 'DLPFC',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
writeLines('Differentially expressed genes at an FDR 0.05 and logFC 1.2')
tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Apoe = DE))
all.fit = c(all.fit, list(Apoe = FIT))
```

### Differential expression analysis (with Diagnosis and Sex)
Differential expression is performed on the Diagnosis x Sex variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and folde change of 1.2 are
```{r diff.exp3, fig.height=6, fig.width=18, cache=FALSE}
## Modeling Diagnosis, msex and age_death conjointly
COVARIATES$Diagnosis.Sex = paste(COVARIATES$Diagnosis,COVARIATES$Sex, sep = '.') %>%
  factor
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('Diagnosis.Sex', setdiff(postAdjustCovars, 'Sex')), drop = F], Intercept = F)
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
colnames(DESIGN$design) = gsub('Diagnosis.Sex','',colnames(DESIGN$design))
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Fit contrast
contrast = makeContrasts(contrasts=c("(AD.FEMALE+AD.MALE)/2-(CONTROL.FEMALE+CONTROL.MALE)/2",
                                     "(AD.FEMALE+CONTROL.FEMALE)/2-(AD.MALE+CONTROL.MALE)/2",
                                     "(AD.FEMALE-CONTROL.FEMALE)/2-(AD.MALE-CONTROL.MALE)/2",
                                     "AD.FEMALE-CONTROL.FEMALE",
                                     "AD.MALE-CONTROL.MALE"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differnetial expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = c('AD-CONTROL','FEMALE-MALE','AD-CONTROL.IN.FEMALE-MALE','AD-CONTROL.IN.FEMALE', 'AD-CONTROL.IN.MALE')
DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique()) %>%
  dplyr::mutate(Study = 'ROSMAP', Region = 'DLPFC',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE$Direction[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2)] = 'NONE'
writeLines('Differentially expressed genes at an FDR 0.05 and logFC 1.2')
tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point() + xlim(c(-1,1))
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(.~Comparison, scales = 'fixed')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis.Sex = DE))
all.fit = c(all.fit, list(Diagnosis.Sex = FIT))
```

### Differential expression analysis (with Diagnosis and AOD)
Differential expression is performed on the Diagnosis x AOD variable by controlling for covariates identified above

Genes that are differentially expressed at an FDR <= 0.05 and fold change of 1.2 are
```{r diff.exp4, fig.height=6, fig.width=18, cache=FALSE}
# Get design matrix
DESIGN = getDesignMatrix(COVARIATES[, c('Diagnosis', setdiff(postAdjustCovars, 'age_death')), drop = F], Intercept = F)
ind = grep('Diagnosis', colnames(DESIGN$design))
DESIGN$design[,ind] = DESIGN$design[,ind] * COVARIATES[rownames(DESIGN$design),'age_death']
DESIGN$design = DESIGN$design[,linColumnFinder(DESIGN$design)$indepCols]
colnames(DESIGN$design) = gsub('Diagnosis','AOD.',colnames(DESIGN$design))
# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(cnts, design=DESIGN$design, plot=F)
# Fit linear model using new weights and new design
FIT = lmFit(CQN.GENE_EXPRESSION$E,
            design = DESIGN$design,
            weights = VOOM.WEIGHTS$weights)
# Fit contrast
contrast = makeContrasts(contrasts=c("AOD.AD-AOD.CONTROL",
                                     "(AOD.AD+AOD.CONTROL)/2"),
                         levels = colnames(FIT$coefficients))
FIT.CONTR = contrasts.fit(FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
# Get differential expression
DE = lapply(1:dim(contrast)[2], function(i, FIT){
  topTable(FIT, coef=i, number = dim(FIT)[1], confint = T) %>%
    rownameToFirstColumn('ensembl_gene_id')
}, FIT.CONTR)
names(DE) = c('AOD.AD-AOD.CONTROL','AOD')

DE = DE %>% 
  rbindlist(idcol = 'Comparison') %>%
  left_join(GENE.PARAM %>%
              dplyr::select(ensembl_gene_id, hgnc_symbol, percentage_gene_gc_content, length) %>%
              unique()) %>%
  dplyr::mutate(Study = 'ROSMAP', Region = 'DLPFC',
                Direction = logFC/abs(logFC),
                Direction = factor(Direction, c(-1,1), c('-1' = 'DOWN', '1' = 'UP')),
                Direction = as.character(Direction))
DE[DE$adj.P.Val > 0.05 | abs(DE$logFC) < log2(1.2), ]$Direction = 'NONE'
writeLines('Differentially expressed genes at an FDR 0.05 and logFC 1.2')
tmp = DE %>%
  dplyr::select(ensembl_gene_id, Comparison, Direction) %>%
  group_by(Comparison, Direction) %>%
  dplyr::summarise(FDR_0_05_FC_1.2 = length(unique(ensembl_gene_id))) %>%
  spread(Direction, FDR_0_05_FC_1.2) 
kable(tmp)
p = ggplot(DE, aes(y = -log10(adj.P.Val), x = logFC, color = Direction)) + geom_point()
p = p + scale_color_manual(values = c('green','grey','red'))
p = p + facet_grid(.~Comparison, scales = 'free_x')
p
all.diff.exp = c(all.diff.exp, list(Diagnosis.AOD = DE))
all.fit = c(all.fit, list(Diagnosis.AOD = FIT))
```

### Store files in synapse

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22092490';
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective protien coding transcript level counts with CQN normalisation ( DLPFC )';
thisFileName <- 'ROSMAP_TranscriptNorm.Rmd'
# Github link
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/prepscripts/',thisFileName))
```

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
activityName = 'Covariate and Diagnosis Regression';
activityDescription = 'Covariate analysis and Regrsison of AD Diagnosis of aligned effective counts with CQN normalisation DLPFC';
CODE <- syn_temp$store(synapseclient$Folder(name = "ROSMAP", parentId = parentId))
#Set Used SynIDs For Provenance
Syns_Used <- c("syn10507739","syn3191087","syn7116000","syn4300313","syn8698240","syn3382527","syn22087693","syn6126119","syn6126114")

# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= c('ROSMAP'), 
  study = 'ROSMAP', 
  organism = 'HomoSapiens',
  consortium	= 'AMPAD',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh38'
)
# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store covariates
COVARIATES = rownameToFirstColumn(COVARIATES, 'SampleID')
write.table(COVARIATES, file = 'ROSMAP_SVA_Diagnosis_Covariates.tsv', row.names = F, sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_SVA_Diagnosis_Covariates.tsv', name = 'Covariates', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'covariates'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_Counts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_Counts.tsv', name = 'Counts (filtered raw)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store Estimated counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_Counts.tsv', name = 'Counts (estimated)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'estimatedCounts'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store logCPM
CQN.GENE_EXPRESSION$y %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_logCPM.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_logCPM.tsv', name = 'Counts (filtered logCPM)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'filteredLCPM'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store cqn offsets
CQN.GENE_EXPRESSION$offset %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_offset.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_offset.tsv', name = 'Gene length and GC content offset', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'offset'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_Design.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_Design.tsv', name = 'Design Matrix', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'designMatrix'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store surrogate variables
SURR.VAR %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_SurVar.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_SurVar.tsv', name = 'Surrogate Variables', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'surrogateVariables'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_Diagnosis_Regressed_netResidualExpression.tsv', name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store differential expression results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'ROSMAP_DLPFC_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_DLPFC_DiffExpression.tsv', name = 'Differential Expression Results (Tissue.Diagnosis)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'diffExp'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

stopCluster(cl)
```

```{r DelFiles, echo=T, results='hide'}
system('rm *.tsv')
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp  <- synapseclient$Synapse()
syn_temp$login()
setwd("~/AD_TargetRank/code/")
source("~/AD_TargetRank/utilityFunctions/knitfile2synapseClient.R")
source("~/AD_TargetRank/utilityFunctions/hook_synapseMdSyntax_plot.R")
createAndKnitToFolderEntityClient(file = "~/AD_TargetRank/code/ROSMAP_TranscriptNorm.Rmd",
                                          parentId ="syn22092490",
                                          folderName = 'ROSMAP')
```

### 2020-05-28 code/02-Transcript_Master.Rmd
Transcript Profilers

---
title: "GENEX: Transcript Report"
author: "Jake Gockley"
date: "05/28/2020"
header-includes:
  - \usepackage{multicol}
  - \newcommand{\btwocol}{\begin{multicols}{2}}
  - \newcommand{\etwocol}{\end{multicols}}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: default
  html_document: default
classoption: landscape
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(readr)
library(biomaRt)
library(knitr)
library(kableExtra)
library(tidyverse)
library(grid)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(yaml)
knitr::opts_chunk$set(
  #root.dir = '/home/jgockley/AD_TargetRank/',
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)

#webshot::install_phantomjs()
```

```{r import, eval = T, include=F }
reticulate::use_python("/usr/bin/python", required = TRUE)
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

Gene <- "GENEX"
#Gene <- "GFAP"

#DE Results
MSBB_DE <- as.data.frame( read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/MSBB_DE.tsv', header= T, sep ='\t', stringsAsFactors = F  )) 
Mayo_DE <- as.data.frame( read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/Mayo_DE.tsv', header= T, sep ='\t', stringsAsFactors = F ) )
Rosmap_DE <- as.data.frame( read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/RosMap_DE.tsv', header= T, sep ='\t', stringsAsFactors = F ) )

#Normalized Expression
MSBB_EXP <- read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/MSBB_Exp.tsv', header= T, sep ='\t' )
colnames( MSBB_EXP ) <- gsub( 'X', '', colnames( MSBB_EXP ) )

Mayo_EXP <- read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/Mayo_Exp.tsv', header= T, sep ='\t' )
colnames( Mayo_EXP ) <- gsub( 'X', '', colnames( Mayo_EXP ) )
colnames( Mayo_EXP ) <- gsub( '_TC', '_TCX', colnames( Mayo_EXP ) )

Rosmap_EXP <- read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/RosMap_Exp.tsv', header= T, sep ='\t' )
colnames( Rosmap_EXP ) <- gsub( 'X', '', colnames( Rosmap_EXP ) )

#Covariates
MSBB_Cov <- data.table::fread( syn_temp$get('syn22092576')$path )
row.names(MSBB_Cov) <- MSBB_Cov$SampleID
Mayo_Cov <- data.table::fread( syn_temp$get('syn22093739')$path )
row.names(Mayo_Cov) <- Mayo_Cov$SampleID
Rosmap_Cov <- data.table::fread( syn_temp$get('syn22097275')$path )
row.names(Rosmap_Cov) <- Rosmap_Cov$SampleID

Trans <- read.table( '~/AD_TargetRank/Transcript_Profiles/Temp_Files/Translate.tsv', header= T, sep ='\t', stringsAsFactors = F )

#Adjust the P-Value for total Transcripts tested:
Rosmap_Adj <- as.numeric(table(Trans$ensembl_transcript_id %in% row.names(Rosmap_EXP))[2])
MSBB_Adj <- as.numeric(table(Trans$ensembl_transcript_id %in% row.names(MSBB_EXP))[2])
Mayo_Adj <- as.numeric(table(Trans$ensembl_transcript_id %in% row.names(Mayo_EXP))[2])
colnames(Mayo_DE)[ colnames(Mayo_DE) == 'Tissue.ref'] <- 'Tissue'

Colors <- c( "black", "red4", "navyblue" )
names(Colors) <- c( "NONE", "UP", "DOWN" )

#Plot Log2FC's
Keepers <- Trans[ Trans$hgnc_symbol == Gene,]$ensembl_transcript_id

Log_Plot <- function( data, tiss, comp, model, adj){
  #'@data data frame eg Rosmap_DE
  #'@tiss the tissue to use eg DLPFC
  #'@comp the comparison to use eg AD-CONTROL
  #'@model the model to use eg Diagnosis.Sex
  #'@adj multicomps adjustment val eg Rosmap_Adj

  eval(parse(text=paste0( 'pDat <- data[ data$Comparison == \'', comp,'\' & data$Model == \'', model,'\' & data$ensembl_gene_id %in% Keepers & data$Tissue == \'', tiss ,'\', ]' )))
  pDat$adj.P.Val <- p.adjust( pDat$P.Value, method = 'fdr', n=adj )
  pDat$Direction[pDat$adj.P.Val <= 0.05 & pDat$logFC <= -log2(1.2)] = 'DOWN'
  pDat$Direction[pDat$adj.P.Val <= 0.05 & pDat$logFC >= log2(1.2)] = 'UP'

  # Create a text
  #grob <- grobTree(textGrob(paste0(tiss,': ',Gene), x=0.98,  y=1, hjust=0,
  #gp=gpar(col="black", fontsize=13, fontface="bold")))

  if( (tiss %in% c('DLPFC','TCX','PHG')) == T){
    plot <- ggplot(  pDat   , aes(x=logFC, y=ensembl_gene_id , colour=Direction)) +
      scale_y_discrete() + 
      scale_color_manual(values=c("#000000", "#7F0000" , "#000099"), breaks =c("NONE", "UP" , "DOWN")) +
      geom_errorbar(aes(xmin=CI.L, xmax=CI.R), width=.1) +
      geom_line() +
      geom_point() +
      #annotation_custom(grob) 
      ggtitle(paste0(tiss,': ',Gene)) + theme(legend.position = "none") + theme(axis.title.y = element_blank()) +
      theme(axis.text.y = element_text(face="bold", size=6, angle=45)) +
      geom_vline(xintercept = 0, linetype="dotted", color = "black", size=1.5)
  }else{
     plot <- ggplot(  pDat   , aes(x=logFC, y=ensembl_gene_id , colour=Direction)) +
      scale_y_discrete() + 
      scale_color_manual(values=c("#000000", "#7F0000" , "#000099"), breaks =c("NONE", "UP" , "DOWN")) +
      geom_errorbar(aes(xmin=CI.L, xmax=CI.R), width=.1) +
      geom_line() +
      geom_point() +
      #annotation_custom(grob) 
      ggtitle(paste0(tiss,': ',Gene)) + theme(legend.position = "none") + 
      theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
      geom_vline(xintercept = 0, linetype="dotted", color = "black", size=1.5)
  } 
  return(plot)
}

DLPFC <- Log_Plot( Rosmap_DE, 'DLPFC', 'AD-CONTROL', 'Diagnosis.Sex', Rosmap_Adj )
TCX <- Log_Plot( Mayo_DE, 'TCX', 'AD-CONTROL', 'Diagnosis', Mayo_Adj  )
CBE <- Log_Plot( Mayo_DE, 'CBE', 'AD-CONTROL', 'Diagnosis', Mayo_Adj  )
PHG <- Log_Plot( MSBB_DE, 'PHG', 'AD-CONTROL', 'Diagnosis', MSBB_Adj  )
IFG <- Log_Plot( MSBB_DE, 'IFG', 'AD-CONTROL', 'Diagnosis', MSBB_Adj  )
FP <- Log_Plot( MSBB_DE, 'FP', 'AD-CONTROL', 'Diagnosis', MSBB_Adj  )
STG <- Log_Plot( MSBB_DE, 'STG', 'AD-CONTROL', 'Diagnosis', MSBB_Adj  )
```

### Log2 Fold Change of Transcripts By Brain Region Case V. Control
Red indicates over expressed and Blue indicates under expressed in cases versus Controls \newline
AD = CDR >= 1 & BRAAK >= 4 & CERAD >= 2 \newline
CONTROL = CDR <= 0.5 & BRAAK <= 3 & CERAD <= 1 \newline

```{r PlotLog, cache = FALSE, fig.height = 5, fig.width = 14}
grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,  ncol=7)
```

```{r pagebreak, results='asis', eval=is_latex_output()}
cat('\\pagebreak')
```

### Transcript Expression Level Boxplot Case V. Control
Red indicates over expressed and Blue indicates under expressed in cases versus Controls \newline
AD = CDR >= 1 & BRAAK >= 4 & CERAD >= 2 - Red \newline
CONTROL = CDR <= 0.5 & BRAAK <= 3 & CERAD <= 1 - Blue \newline
Other - Green

```{r formatBoxPlot, eval = T, include=F }
Rosmap_Cov$Tissue <- 'DLPFC'

BoxPloter <- function( EXP, Cov, Tiss ){
  #'@EXP the expression dataset to plot eg Rosmap_EXP
  #'@Cov the covariate data frame eg Rosmap_Cov
  #'@Tiss The target tissue to plot eg. DLPFC

  temp <- as.data.frame( EXP[ row.names( EXP )[ row.names( EXP ) %in% Keepers], ] )
  temp$Transcript <- row.names(temp)

  Cov <- Cov[ as.character(Cov$Tissue) == Tiss, ]

  foo <- reshape2::melt( t(temp), id="Transcript" )
  Cov <- as.data.frame(Cov)
  row.names(Cov) <- Cov$SampleID
  foo$Diagnosis <- Cov[ as.character(foo$Var1), ]$Diagnosis
  foo$value <- as.numeric(as.character(foo$value))

  foo <- foo[ (foo$Diagnosis %in% c('AD', 'CONTROL', 'OTHER'))==T, ]

  plot <- ggplot(data = foo, aes(x=Var2, y=value)) + 
    geom_violin(  ) + geom_boxplot( width=0.1, outlier.shape = NA) +  theme(legend.position = "none") +
    geom_jitter(aes(col=Diagnosis), alpha = .4, size=.2) + # geom_jitter() +
    theme(axis.text.x = element_text(face="bold", size=6, angle=45)) + ggtitle(paste0(Tiss,': ',Gene)) +
    theme(axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none" )

  return( plot )
}

DLPFC <- BoxPloter( Rosmap_EXP, Rosmap_Cov, 'DLPFC' )

Mayo_Cov$Diagnosis <- do.call(rbind, strsplit( Mayo_Cov$Tissue.Diagnosis, '[.]' ))[,2]
Mayo_Cov$Tissue <- do.call(rbind, strsplit( Mayo_Cov$Tissue.SourceDiagnosis, '[.]' ))[,1]

TCX <- BoxPloter( Mayo_EXP, Mayo_Cov, 'TCX' )
CBE <- BoxPloter( Mayo_EXP, Mayo_Cov, 'CBE' )

MSBB_Cov$Diagnosis <- do.call(rbind, strsplit( MSBB_Cov$Tissue.Diagnosis, '[.]' ))[,2]
MSBB_Cov$Tissue <- do.call(rbind, strsplit( MSBB_Cov$Tissue.Diagnosis, '[.]' ))[,1]

IFG <- BoxPloter( MSBB_EXP, MSBB_Cov, 'IFG' )
PHG <- BoxPloter( MSBB_EXP, MSBB_Cov, 'PHG' )
FP <- BoxPloter( MSBB_EXP, MSBB_Cov, 'FP' )
STG <- BoxPloter( MSBB_EXP, MSBB_Cov, 'STG' )
```

```{r PlotLog2, cache = FALSE, fig.height = 8, fig.width = 14}
grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,ncol=4, nrow=2)
```

### 2020-05-28 code/01-Transcript_Profiler_Initializer.R

@@ -1,0 +1,147 @@
+#This script takes a user specified Yaml file and creates a markdown file from it
+#To run: Rscript code/01-Transcript_Profiler_Initializer.R <Configuration YAML FILE>
+#eg. Rscript code/01-Initializer.r configs/Initial_49_Targets.yaml
+
+library(yaml)
+args <- commandArgs(trailingOnly=TRUE)
+
+#READ IN Config
+#configuration<-'configs/Initial_49_Targets.yaml'
+configuration <- args[1]
+config <- read_yaml(configuration)
+
+setwd( config$filedir )
+
+#Load Synapse and our DE Data:
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
+#DE Results
+MSBB_DE <- data.table::fread( syn_temp$get('syn22092611')$path )
+colnames(Mayo_DE)[ colnames(Mayo_DE) == 'Tissue.ref' ] <- 'Tissue'
+Mayo_DE <- data.table::fread( syn_temp$get('syn22093747')$path )
+Rosmap_DE <- data.table::fread( syn_temp$get('syn22097286')$path )
+colnames(Rosmap_DE)[ colnames(Rosmap_DE) == 'Region' ] <- 'Tissue'
+
+#Normalized Expression
+MSBB_EXP <- as.data.frame(data.table::fread( syn_temp$get('syn22092607')$path ))
+row.names(MSBB_EXP) <- MSBB_EXP$ensembl_gene_id
+MSBB_EXP <- MSBB_EXP[,colnames(MSBB_EXP) != 'ensembl_gene_id']
+
+Mayo_EXP <- as.data.frame( data.table::fread( syn_temp$get('syn22093746')$path ) )
+row.names(Mayo_EXP) <- Mayo_EXP$ensembl_gene_id
+Mayo_EXP <- Mayo_EXP[,colnames(Mayo_EXP) != 'ensembl_gene_id']
+
+Rosmap_EXP <- as.data.frame( data.table::fread( syn_temp$get('syn22097285')$path ) )
+row.names(Rosmap_EXP) <- Rosmap_EXP$ensembl_gene_id
+Rosmap_EXP <- Rosmap_EXP[,colnames(Rosmap_EXP) != 'ensembl_gene_id']
+
+#Covariates
[…]

### 2020-05-28 code/02-Transcript_Master.Rmd
Transcript Profilers

@@
 Gene <- "GENEX"
-#Gene <- "GFAP"
+#Gene <- "CTHRC1"
 
@@
   eval(parse(text=paste0( 'pDat <- data[ data$Comparison == \'', comp,'\' & data$Model == \'', model,'\' & data$ensembl_gene_id %in% Keepers & data$Tissue == \'', tiss ,'\', ]' )))
+  if( dim(pDat)[1] > 0 ){
     pDat$adj.P.Val <- p.adjust( pDat$P.Value, method = 'fdr', n=adj )
@@
         geom_vline(xintercept = 0, linetype="dotted", color = "black", size=1.5)
+
+    } 
+  }else{
+    plot <- NULL
   }
@@
 STG <- Log_Plot( MSBB_DE, 'STG', 'AD-CONTROL', 'Diagnosis', MSBB_Adj  )
+
+LIST <- NULL
+for( test in c('DLPFC','TCX','CBE','PHG','IFG','FP','STG') ){
+  if( eval(parse(text= paste0( 'is.null(', test, ')' ))) == TRUE ){
+  }else{
+    LIST <- c(LIST, test)
+  }
+}
 ```
@@
 ```{r PlotLog, cache = FALSE, fig.height = 5, fig.width = 14}
-grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,  ncol=7)
+N<-length(LIST) 
+if(length(LIST) == 0 ){
+
+}else{
+  PLOTS <- paste0(LIST, collapse=',')
+  eval(parse(text= paste0( 'grid.arrange(', PLOTS , ', ncol=', N, ' )' )))
+}
 ```
@@
   temp <- as.data.frame( EXP[ row.names( EXP )[ row.names( EXP ) %in% Keepers], ] )
+  if( dim(temp)[1] == 0 ){
+    plot <- NULL
+  }else{
     temp$Transcript <- row.names(temp)
@@
       theme(axis.title.y = element_blank(), axis.title.x = element_blank(), legend.position = "none" )
-
+  }
   return( plot )
@@
 STG <- BoxPloter( MSBB_EXP, MSBB_Cov, 'STG' )
+
+LIST <- NULL
+for( test in c('DLPFC','TCX','CBE','PHG','IFG','FP','STG') ){
+  if( eval(parse(text= paste0( 'is.null(', test, ')' ))) == TRUE ){
+  }else{
+    LIST <- c(LIST, test)
+  }
+}
+
 ```
@@
 ```{r PlotLog2, cache = FALSE, fig.height = 8, fig.width = 14}
-grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,ncol=4, nrow=2)
+N<-length(LIST) 
+if(length(LIST) == 0 ){
+
+}else{
+  if(length(LIST) > 4 ){
+    PLOTS <- paste0(LIST, collapse=',')
+    eval(parse(text= paste0( 'grid.arrange(', PLOTS , ', ncol=4, nrow=2 )' )))
+  }else{
+    PLOTS <- paste0(LIST, collapse=',')
+    eval(parse(text= paste0( 'grid.arrange(', PLOTS , ', ncol=4, nrow=1 )' )))
+  }
+}
+
+#grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,ncol=4, nrow=2)
 ```

### 2020-05-28 code/01-Transcript_Profiler_Initializer.R

@@ -130,7 +130,7 @@ thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref=
 thisFile <- githubr::getPermlink( repository = thisRepo, repositoryPath=paste0( 'code/',thisFileName ))
 thisFile2 <- githubr::getPermlink( repository = thisRepo, repositoryPath=paste0( 'code/',thisFileName2 ))
 
-Syns_used <- c("syn22092611", "syn22093747", "syn22097286", "syn22092607", "syn22093746", "syn22097285", "syn22092576", "syn22093739", "syn22097275")
+Syns_Used <- c("syn22092611", "syn22093747", "syn22097286", "syn22092607", "syn22093746", "syn22097285", "syn22092576", "syn22093739", "syn22097275")
 CODE <- syn_temp$store(synapseclient$Folder(name = config$runname, parentId = parentId))
 
 for( name in genes ){
@@ -144,4 +144,17 @@ for( name in genes ){
   system( paste0( 'rm ~/AD_TargetRank/code/03-', name, '.Rmd' ) )
   system( paste0( 'rm -r ~/AD_TargetRank/Transcript_Profiles/Initial_49_Targets/03-', name, '_files/' ) )
   
+  reticulate::use_python("/usr/bin/python", required = TRUE)
+  synapseclient <- reticulate::import("synapseclient")
+  syn_temp <- synapseclient$Synapse()
+  syn_temp$login()
+  
+  ENRICH_OBJ <- syn_temp$store( synapseclient$File( path= paste0( '~/AD_TargetRank/Transcript_Profiles/', config$runname,'/03-', name,'.pdf'), name = paste0( 'Transcript analysis for ', name ), 
+                                                     parentId=CODE$properties$id ), 
+                                                     used = Syns_Used,
+                                                     activityName = activityName, 
+                                                     executed = c(thisFile,thisFile2),
+                                                     activityDescription = activityDescription)
 }
+
+

### 2020-09-08 code/02-Transcript_Master.Rmd
MetaAnalysisCode

@@
 library(yaml)
+library(MASS)
 knitr::opts_chunk$set(
@@
 #Gene <- "CTHRC1"
+#Gene <- "GFAP"
+#Gene <- 'C4B'
 
@@
 Red indicates over expressed and Blue indicates under expressed in cases versus Controls \newline
-AD = CDR >= 1 & BRAAK >= 4 & CERAD >= 2 \newline
-CONTROL = CDR <= 0.5 & BRAAK <= 3 & CERAD <= 1 \newline
+AD = CDR >= 1 & Braak >= 4 & CERAD >= 2 \newline
+CONTROL = CDR <= 0.5 & Braak <= 3 & CERAD <= 1 \newline
 
@@
 
+### Log2 Fold Change of Transcripts By Cognitive Assesment
+Red indicates over expressed and Blue indicates under expressed in cases versus Controls \newline
+ROSMAP: Cogdx = 4 Versus CogDx = 1
+\begin{itemize}
+\item[] Cogdx = 4: AD: Alzheimer’s dementia and NO other cause of CI (NINCDS PROB AD)
+\item[] Cogdx = 1: NCI: No cognitive impairment (No impaired domains)
+\end{itemize}
+
+MSBB: CDR = 5 Versus CDR = 0
+\begin{itemize}
+\item[] CDR = 5: Severe cognitive decline
+\item[] CDR = 0: No cognitive decline
+\end{itemize}
+
+```{r Cognition, eval = T, include=F}
+DLPFC <- Log_Plot( Rosmap_DE, 'DLPFC', 'cogdx4-cogdx1', 'cogdx', Rosmap_Adj )
+
+PHG <- Log_Plot( MSBB_DE, 'PHG', 'CDR', 'SourceDiagnosis', MSBB_Adj  )
+IFG <- Log_Plot( MSBB_DE, 'IFG', 'CDR', 'SourceDiagnosis', MSBB_Adj  )
+FP <- Log_Plot( MSBB_DE, 'FP', 'CDR', 'SourceDiagnosis', MSBB_Adj  )
+STG <- Log_Plot( MSBB_DE, 'STG', 'CDR', 'SourceDiagnosis', MSBB_Adj  )
+
+LIST <- NULL
+for( test in c('DLPFC','PHG','IFG','FP','STG') ){
+  if( eval(parse(text= paste0( 'is.null(', test, ')' ))) == TRUE ){
+  }else{
+    LIST <- c(LIST, test)
+  }
+}
+```
+
+```{r PlotLogCog, cache = FALSE, fig.height = 5, fig.width = 14}
+N<-length(LIST) 
+if(length(LIST) == 0 ){
+
+}else{
+  PLOTS <- paste0(LIST, collapse=',')
+  eval(parse(text= paste0( 'grid.arrange(', PLOTS , ', ncol=', N, ' )' )))
+}
+```
+
+```{r pagebreak2, results='asis', eval=is_latex_output()}
+cat('\\pagebreak')
+```
+
+### Neuropathology Associations
+All associations are logistic ordinal regression models preformed on Braak Score with P-Values correct to the total number of transcripts belonging to the user input gene list expressed in the respective tissue.
+\newline
+
+```{r WrangleNeuroPath, eval = T, include=F }
+MSBB_N   <- read.table( file = '~/AD_TargetRank/Transcript_Profiles/Temp_Files/MSBB_NeuroPath.tsv', header = T, sep = '\t', stringsAsFactors = F )
+row.names(MSBB_N) <- MSBB_N$sampleIdentifier 
+ROSMAP_N <- read.table( file = '~/AD_TargetRank/Transcript_Profiles/Temp_Files/Rosmap_NeuroPath.tsv', header = T, sep = '\t', stringsAsFactors = F )
+row.names(ROSMAP_N) <- ROSMAP_N$Sampleid 
+Mayo_N   <- read.table( file = '~/AD_TargetRank/Transcript_Profiles/Temp_Files/Mayo_NeuroPath.tsv', header = T, sep = '\t', stringsAsFactors = F )
+row.names(Mayo_N) <- Mayo_N$SampleID
+
+MSBB_Braak <- MSBB_N[ , c( "sampleIdentifier", "Tissue1", "bbscore") ]
+colnames(MSBB_Braak) <- c( "Sample", 'Tissue', "Braak" )
+ROSMAP_N$Tissue <- 'DLPFC'
+ROSMAP_Braak <- ROSMAP_N[ , c("Sampleid", "Tissue", "braaksc") ] 
+colnames(ROSMAP_Braak) <- c( "Sample", 'Tissue', "Braak" )
+Mayo_Braak <- Mayo_N[ ,c("SampleID","Tissue","Braak")]
+colnames(Mayo_Braak) <- c( "Sample", 'Tissue', "Braak" )
+
+Braak <- as.data.frame(rbind(Mayo_Braak,ROSMAP_Braak,MSBB_Braak))
+Braak$Sample <- as.character( Braak$Sample )
+Braak$Tissue <- as.character( Braak$Tissue )
+
+Results <- data.frame( matrix( NA, 1 ,10 ) )
+colnames( Results ) <- c( 'Transcript', 'Gene', 'Symbol', 'Tissue', 'Coeff', 
+                          '95_Conf_L', '95_Conf_U', 'Z_Score', 'P_Val', 'FDR_PVal'
+                        )
+
+# Plot Function
+Braak_Plot <- function( data, tiss, gene ){
+  #'@data data frame eg Results
+  #'@tiss the tissue to use eg DLPFC
+  #'@gene the gene to use eg gene
+
+  eval(parse(text=paste0( 'pDat <- data[ data$Symbol %in% gene & data$Tissue == \'', tiss ,'\', ]' )))
+
+  if( dim(pDat)[1] > 0 ){
+
+    if( (tiss %in% c('DLPFC','TCX','PHG')) == T){
+      plot <- ggplot(  pDat   , aes(x=Coeff, y=Transcript , colour=Direction)) +
+        scale_y_discrete() + 
+        scale_color_manual(values=c("#000000", "#7F0000" , "#000099"), breaks =c("NONE", "UP" , "DOWN")) +
+        geom_errorbar(aes(xmin=`95_Conf_L`, xmax=`95_Conf_U`), width=.1) +
+        geom_line() +
+        geom_point() +
+        #annotation_custom(grob) 
+        ggtitle(paste0(tiss,': ',gene)) + theme(legend.position = "none") + theme(axis.title.y = element_blank()) +
+        theme(axis.text.y = element_text(face="bold", size=6, angle=45)) +
+        geom_vline(xintercept = 0, linetype="dotted", color = "black", size=1.5)
+    }else{
+       plot <- ggplot(  pDat   , aes( x=Coeff, y=Transcript, colour=Direction )) +
+        scale_y_discrete() + 
+        scale_color_manual(values=c("#000000", "#7F0000" , "#000099"), breaks =c("NONE", "UP" , "DOWN")) +
+        geom_errorbar(aes(xmin=`95_Conf_L`, xmax=`95_Conf_U`), width=.1) +
+        geom_line() +
+        geom_point() +
+        #annotation_custom(grob) 
+        ggtitle(paste0(tiss,': ',gene)) + theme(legend.position = "none") + 
+        theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
+        geom_vline(xintercept = 0, linetype="dotted", color = "black", size=1.5)
+    } 
+  }else{
+    plot <- NULL
+  }
+  return(plot)
+}
+
+#RunAnalysis  
+for( TI in c( 'DLPFC', 'TCX', 'CBE', 'PHG', 'IFG', 'STG', 'FP' ) ){
+  eval(parse(text= paste0( 'Braak_temp <- Braak[ Braak$Tissue == \'', TI, '\', ]' )))
+  Braak_temp$Braak <- as.factor( Braak_temp$Braak )
+
+  if( TI == 'DLPFC' ){
+    #table(row.names(Rosmap_EXP) %in% Trans$ensembl_transcript_id)
+    EXP_temp <- Rosmap_EXP[,Braak_temp$Sample]
+  }else{
+    if( (TI %in% c('TCX','CBE')) == T  ){
+      EXP_temp <- Mayo_EXP[,Braak_temp$Sample]
+    }else{
+      if( (TI %in% c( 'PHG', 'IFG', 'STG', 'FP')) == T  ){
+        EXP_temp <- MSBB_EXP[,Braak_temp$Sample]
+      }else{
+        warning("Improper Tissue Type")
+      }
+    }
+  }
+
+  row.names(Trans) <- Trans$ensembl_transcript_id
+
+  if( any(Trans[ Trans$hgnc_symbol == Gene,]$ensembl_transcript_id %in% row.names(EXP_temp)) ){
+    Pull <- Trans[ Trans$hgnc_symbol == Gene,]$ensembl_transcript_id[ Trans[ Trans$hgnc_symbol == Gene,]$ensembl_transcript_id %in% row.names(EXP_temp) ]
+    EXP_temp <- EXP_temp[ Pull , ]
+
+    for( i in 1:dim(EXP_temp)[1]){
+
+      Transcript <- row.names(EXP_temp)[i]
+      Ge <- Trans[Transcript,]$hgnc_symbol
+      ENSG <- Trans[Transcript,]$ensembl_gene_id
+
+      foo <- cbind( Braak_temp[,1:3], 
+                    t(EXP_temp[ row.names(EXP_temp)[i], row.names(Braak_temp) ])
+                  )
+      foo$Braak <- as.factor(foo$Braak)
+      colnames(foo)[4] <- "Ge"
+      m <- polr( Braak ~ Ge, data = foo, Hess=TRUE)
+
+      summary(m)
+      ctable <- coef(summary(m))
+      p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
+      ctable <- cbind(ctable, "p value" = p)
+      ci <- confint(m)
+
+      Adj_P <- p.adjust( ctable['Ge','p value'], method = 'fdr', n = dim(EXP_temp)[1] )
+
+      Entry <- c( Transcript, ENSG, Ge, TI, ctable['Ge','Value'], 
+                  as.numeric(ci[1]), as.numeric(ci[2]), ctable['Ge', "t value"] , ctable['Ge','p value'],
+                  Adj_P 
+                )
+      Results <- rbind( Results[1:dim(Results)[1], ], Entry)
+    }
+  }
+  else{
+  }
+  Results <- Results[ complete.cases(Results), ]
+  if( dim(Results[ complete.cases(Results), ])[1] > 0){
+    Results$Direction <- 'NONE'
+    Results$Direction[Results$FDR_PVal <= 0.05 & Results$Z_Score < 0 ] = 'DOWN'
+    Results$Direction[Results$FDR_PVal <= 0.05 & Results$Z_Score > 0] = 'UP'
+
+    Results$Coeff <- as.numeric(as.character( Results$Coeff ))
+    Results$`95_Conf_L` <- as.numeric(as.character( Results$`95_Conf_L` ))
+    Results$`95_Conf_U` <- as.numeric(as.character( Results$`95_Conf_U` ))
+    Results$Z_Score <- as.numeric(as.character( Results$Z_Score ))
+    Results$P_Val  <- as.numeric(as.character( Results$P_Val ))
+    Results$FDR_PVal <- as.numeric(as.character( Results$FDR_PVal ))
+
+    eval(parse(text= paste0( '\'', TI, '\' <- Braak_Plot( Results, \'', TI, '\', Gene )' )))
+  }else{
+  }
+}
+
+LIST <- NULL
+for( test in c('DLPFC','TCX','CBE','PHG','IFG','FP','STG') ){
+  if( eval(parse(text= paste0( 'is.null(', test, ')' ))) == TRUE ){
+  }else{
+    LIST <- c(LIST, test)
+  }
+}
+```
+
+```{r PlotLogBrak, cache = FALSE, fig.height = 5, fig.width = 14}
+N<-length(LIST) 
+if(length(LIST) == 0 ){
+
+}else{
+  PLOTS <- paste0(LIST, collapse=',')
+  eval(parse(text= paste0( 'grid.arrange(', PLOTS , ', ncol=', N, ' )' )))
+}
+```
+
+```{r pagebreak3, results='asis', eval=is_latex_output()}
+cat('\\pagebreak')
+```
+
 ### Transcript Expression Level Boxplot Case V. Control
-Red indicates over expressed and Blue indicates under expressed in cases versus Controls \newline
-AD = CDR >= 1 & BRAAK >= 4 & CERAD >= 2 - Red \newline
-CONTROL = CDR <= 0.5 & BRAAK <= 3 & CERAD <= 1 - Blue \newline
+AD = CDR >= 1 & Braak >= 4 & CERAD >= 2 - Red \newline
+CONTROL = CDR <= 0.5 & Braak <= 3 & CERAD <= 1 - Blue \newline
 Other - Green
@@
 }
+```
 
-#grid.arrange(DLPFC, TCX, CBE, PHG, IFG, FP, STG,ncol=4, nrow=2)
+```{r pagebreak4, results='asis', eval=is_latex_output()}
+cat('\\pagebreak')
 ```
+
+### Gene Model
+Only expressed trascripts i.e. those in the analysis above are used in the gene model. \newline
+
+```{r ExonPlot, eval = T, include=F }
+
+library(Gviz)
+library(rtracklayer)
+library(trackViewer)
+library(org.Hs.eg.db)
+library( TxDb.Hsapiens.UCSC.hg38.knownGene )
+library("biomaRt")
+
+gene_id <- Gene
+filters <- 'hgnc_symbol'
+attrs <- c(filters, "ensembl_gene_id", "chromosome_name", "start_position", "end_position", 'strand')
+
+host <- 'ensembl.org'
+organism <- 'hsa'
+ensembl <- biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host = host)
+
+ds <- biomaRt::listDatasets(ensembl)[, "dataset"]
+ds <- grep(paste0("^", organism), ds, value = TRUE)
+if (length(ds) == 0) {
+  stop(paste("Mart not found for:", organism))
+} else if (length(ds) > 1) {
+  message("Found several marts")
+  sapply(ds, function(d) message(paste(which(ds == d), d, sep = ": ")))
+  n <- readline(paste0("Choose mart (1-", length(ds), ") : "))
+  ds <- ds[as.integer(n)]
+}
+ensembl <- biomaRt::useDataset(ds, mart = ensembl)
+
+trans <- getBM(filters = 'hgnc_symbol',
+               attributes = attrs,
+               values = Gene,
+               mart = ensembl )
+
+if( trans$strand < 0 ){
+  Strand <- '-'
+}else{
+  Strand <- '+'
+}
+
+# txdb <- makeTxDbFromEnsembl("Homo sapiens", server="useastdb.ensembl.org")
+TxDB_ENS <- makeTxDbFromEnsembl("Homo sapiens", server="useastdb.ensembl.org")
+
+#entrezID <- get(Gene, org.Hs.egSYMBOL2EG)
+EID <- Trans[ Trans$hgnc_symbol == Gene, ]$ensembl_gene_id[1]
+#geneTrack(EID,TxDB_ENS)[[1]]
+theTrack <- geneTrack(EID,TxDB_ENS)[[1]]
+
+gr <- GRanges( paste0('chr',trans$chromosome_name ), IRanges( trans$start_position, trans$end_position), strand=Strand)
+trs <- geneModelFromTxdb(TxDb.Hsapiens.UCSC.hg38.knownGene,
+  org.Hs.eg.db,
+  gr=gr)
+
+viewerStyle <- trackViewerStyle()
+setTrackViewerStyleParam(viewerStyle, "margin", c(.1, .12, .02, .02))
+
+TIDs <- Trans[ Trans$hgnc_symbol == Gene, ]$ensembl_transcript_id
+TIDs <- TIDs[ !duplicated(TIDs) ]
+Trs <- trs[ names(trs)[ do.call( rbind, strsplit( names(trs), '[.]' ))[,1]  %in% TIDs ] ]
+
+track <- trackList( Trs)
+for( i in 1:length(track) ){
+  setTrackStyleParam( track[[i]], "ylabgp", list(cex=.8) )
+}
+
+if( any( grepl('MHC', gr@seqnames@values )) ){
+  PLOT <- 'NO'
+}else{
+   PLOT <- 'YES'
+}
+```
+
+```{r PlotExon, cache = FALSE, fig.height = 8, fig.width = 14}
+library(trackViewer)
+if( PLOT == 'YES' ){
+  viewTracks(track, 
+    gr=gr, viewerStyle=viewerStyle, 
+    autoOptimizeStyle=TRUE)
+}else{
+  print("Gene model has some transcripts assigned to alternate chromosomes a more detailed look is Required")
+}
+
+```

### 2020-09-08 code/RNA-Seq_Meta_Analysis.Rmd
MetaAnalysisCode

---
title: "Meta analysis of reprocessed RNASeq data from AMP-AD for hg38 Reprocessed Data (all 7 brain regions)"
author: "Adapted from Thanneer Perumal's code"
date: "`r date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
## Load required libraries
library(CovariateAnalysis)
library(data.table)
library(plyr)
library(tidyverse)
#library(synapser)
library(knitr)
library(githubr)
#synLogin()
library(doParallel)
library(foreach)
cl = makeCluster(detectCores()-2)
registerDoParallel(cl)
options(xtable.type="html")
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
#parentId = 'syn18936948';
parentId = 'syn22351719';
activityName = 'Meta analysis of differential expression hg38 Reprocessed';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (all 7 brain regions)';
thisFileName <- 'RNA-Seq_Meta_Analysis.Rmd'
# Github link
thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

## Data download
### Get study specific covariates and adjusted expression from synapse

```{r get.exp}
# Function to read tsv/csv files from synapse
downloadFile <- function(id){ 
  fread(syn_temp$get(id)$path, data.table = F)
}
# Read covariates
covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
all.used.ids = as.character(covar.ids)
covar =  covar.ids %>%
  lapply(downloadFile)
# Read adjusted expression
#_#expr.ids = c(ROSMAP = 'syn8456719', MSSM = 'syn8485027', MAYO = 'syn8466826')
expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
all.used.ids = c(all.used.ids, as.character(expr.ids))
expr =  expr.ids %>%
  lapply(downloadFile) %>%
  plyr::join_all(type = 'inner')
```

```{r modify.covar}
covar$ROSMAP = covar$ROSMAP %>%
  dplyr::select(SampleID, Sex, Diagnosis, APOE4) %>%
  dplyr::mutate(Tissue = 'DLPFC')
covar$MSSM = covar$MSSM %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  dplyr::rename(Sex = Sex) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar$MAYO = covar$MAYO %>%
  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
covar = covar %>%
  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study') %>%
  dplyr::select(-Tissue1)
```

## Perform fixed and random effect meta-analysis for each gene across 7 different brain regions
#### Meta-analysis of AD-CONTROL
```{r ad.cntrl}
# Mean and sd of gene expression in each tissue x diagnosis group
tissue.dx.summary = plyr::ddply(covar, .(Tissue, Diagnosis), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison
meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary, .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.ad_cntrl, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))+geom_point()
p[[2]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.fixed), x = TE.fixed))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
p[[3]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.random), x = TE.random))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
# Get gene sets
tmp = meta.anlz.ad_cntrl %>%
  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('ad.control.', names(gs))
all.gs = gs
```

#### Meta-analysis of AD-CONTROL (sex specific)
```{r ad.cntrl.sex}
# Mean and sd of gene expression in each tissue x diagnosis x sex group
tissue.dx.sex.summary = plyr::ddply(covar, .(Tissue, Diagnosis, Sex), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison (sex specific)
meta.anlz.ad_cntrl.sex = plyr::ddply(tissue.dx.sex.summary, .(ensembl_gene_id, Sex), .fun = function(x){
  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.ad_cntrl.sex, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
p[[1]] = p[[1]] + geom_point() + facet_grid(.~Sex)
p[[2]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.fixed), x = TE.fixed))
p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
p[[3]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.random), x = TE.random))
p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
ggpubr::ggarrange(plotlist = p, ncol = 1, nrow = 3)
# Get gene sets
tmp = meta.anlz.ad_cntrl.sex %>%
  dplyr::select(ensembl_gene_id, Sex, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id, -Sex) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Sex, Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('ad.control.', names(gs))
all.gs = c(all.gs, gs)
```

#### Meta-analysis of APOE42-APOE40
```{r apoe2.0}
# Mean and sd of gene expression in each Tissue x APOE4 category
tissue.apoe4.summary = plyr::ddply(covar, .(Tissue, APOE4), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$ensembl_gene_id,
             n = dim(x)[1],
             mn = rowMeans(y[,x$SampleID], na.rm = T),
             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison (apoe4 2 vs 0)
meta.anlz.apoe4_2_0 = plyr::ddply(tissue.apoe4.summary, .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, APOE4 == '2')
  rownames(exp.effect) = exp.effect$Tissue
  cntrl.effect = dplyr::filter(x, APOE4 == '0')
  rownames(cntrl.effect) = cntrl.effect$Tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.apoe4_2_0, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
p[[1]] = p[[1]] + geom_point()
p[[2]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.fixed), x = TE.fixed))
p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
p[[3]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.random), x = TE.random))
p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
# Get gene sets
tmp = meta.anlz.apoe4_2_0 %>%
  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('apoe4.2.0.', names(gs))
all.gs = c(all.gs, gs)
```

## Store results in synapse
```{r syn.store, cache = FALSE, include=FALSE}
# Create folder to store results in synapse
#CODE = Folder(name = 'MetaAnalysis', parentId = parentId)
#CODE = synStore(CODE)
CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))

Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )
# Write results to files
fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'AD-Control meta-analysis across 7 brain regions hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(meta.anlz.ad_cntrl.sex, file = 'meta.anlz.ad_cntrl.sex.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.sex.tsv', name = 'AD-Control meta-analysis across 7 brain regions (sex specific) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(meta.anlz.apoe4_2_0, file = 'meta.anlz.apoe4_2_0.sex.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.apoe4_2_0.sex.tsv', name = 'AD-Control meta-analysis across 7 brain regions (APOE4) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
save(list = c('all.gs'), file = 'all.gensets.RData')
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='all.gensets.RData', name = 'All gene sets (in RData format) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

```

```{r knit2synapse, eval=FALSE}
synapseclient <- reticulate::import("synapseclient")
syn_temp <- synapseclient$Synapse()
syn_temp$login()

setwd("~/TWAS/code/")
source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")

createAndKnitToFolderEntityClient(file = "metaAnalysisForDiffExp_hg38.Rmd",
                                 parentId = "syn18936948",
                                 folderName = "MetaAnalysis")
```

## Source code

### 2020-09-08 code/01-Transcript_Profiler_Initializer.R

@@ -3,6 +3,9 @@
 #eg. Rscript code/01-Initializer.r configs/Initial_49_Targets.yaml
 
 library(yaml)
+library(tidyr)
+library(dplyr)
+library(plyr)
 args <- commandArgs(trailingOnly=TRUE)
 
 #READ IN Config
@@ -120,6 +123,256 @@ if( dir.exists( paste0( '~/AD_TargetRank/Transcript_Profiles/', config$runname )
   system(paste0( 'mkdir ~/AD_TargetRank/Transcript_Profiles/', config$runname ))
 }
 
+#Wrangle MetaData for Neuropath data
+
+METADATA_TECH_ID <- 'syn6100548'
+METADATA_TECH <- syn_temp$get(METADATA_TECH_ID, version = 7)$path %>%
+  read.table(sep=',',header=T, row.names=1, stringsAsFactors = F) %>%
+  group_by(sampleIdentifier) %>%
+  top_n(1, -rRNA.rate) %>%
+  dplyr::select(sampleIdentifier, BrodmannArea, barcode, individualIdentifier, batch, RIN) %>%
+  unique %>%
+  ddply(.(barcode), .fun = function(x){
+    x = dplyr::mutate(x, batch = paste(batch, collapse = ''), RIN = sum(RIN, na.rm = T))  
+  }) 
+
+# Get 90+ age
+METADATA_AGE_ID = 'syn10156693'
+METADATA_AGE <- syn_temp$get(METADATA_AGE_ID, version = 3)$path %>%
+  read.table(sep='\t', header=T, stringsAsFactors = F)
+# Get clinical metadata
+METADATA_CLINICAL_ID <- 'syn6101474'
+METADATA_CLINICAL <- read.csv( syn_temp$get(METADATA_CLINICAL_ID)$path,  header = T, stringsAsFactors = F)
+
+row.names(METADATA_CLINICAL) <- METADATA_CLINICAL$individualIdentifier
+
+###Left Here:
+colnames(METADATA_CLINICAL)[ colnames(METADATA_CLINICAL) == 'individualID'] <- 'individualIdentifier'
+colnames(METADATA_CLINICAL)[ colnames(METADATA_CLINICAL) == 'ageDeath'] <- 'AOD'
+row.names(METADATA_TECH) <- METADATA_TECH$sampleIdentifier
+METADATA <- dplyr::full_join(METADATA_TECH, METADATA_CLINICAL)
+
[…]

### 2020-09-08 code/RankAllExpression.R

@@ -1,0 +1,357 @@
+## This script Creates the RNA-Expression Raw Ranks for every gene Across all Data Sets
+## Non-significant DE between Case and Control are pushed to Zero
+
+library(yaml)
+
+library(readr)
+library(biomaRt)
+library(knitr)
+library(kableExtra)
+library(tidyverse)
+library(grid)
+library(dplyr)
+library(ggplot2)
+library(ggrepel)
+library(gridExtra)
+library(yaml)
+
+setwd('~/AD_TargetRank')
+
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
+print(getwd())
+#Assign your YAML formated config file to a varable
+#configuration <- "configs/EmoryTargets.yaml"
+#config <- read_yaml(file=paste0('./',configuration))
+
+setwd(config$filedir)
+plotdir=paste0( 'runs/totalRank/figures' )
+tabledir=paste0( 'runs/totalRank/tables' )
+
+Tissues <- config$tissue
+
+#Create Synapse Run Folder and retrieve parentID
+CODE <- syn_temp$store(synapseclient$Folder(name = 'totalRank', parentId = 'syn21532475'))
+
+foo <- syn_temp$getChildren('syn21532475')
+Temp <- reticulate::iterate(foo, simplify = F)
+for( i in 1:length(Temp) ){
+  if( Temp[[i]]$name == 'totalRank' ){
+    RunParent <- Temp[[i]]$id
+  }else{
+  }
+}
+
+
+#Load in the differential expression data
+DE <- read.table( syn_temp$get('syn21534583')$path, header=T, sep ="\t")
+#Genes <- Tab$ENSG
+Genes <- as.character(DE$ensembl_gene_id)
+Genes <- Genes[!duplicated(Genes)]
+
+#Construct the rank model
+ModelConstruct <- function( de, Comp, Mod, Sex ){
+  #C#reates a model and plots the target genes
[…]

### 2020-09-16 code/RNA-Seq_Meta_Analysis.Rmd
RNA MetaFinalize

@@
 library(tidyverse)
-#library(synapser)
 library(knitr)
 library(githubr)
-#synLogin()
 library(doParallel)
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
 #parentId = 'syn18936948';
 parentId = 'syn22351719';
-activityName = 'Meta analysis of differential expression hg38 Reprocessed';
-activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (all 7 brain regions)';
+activityName = 'Meta analysis of differential expression hg38 Reprocessed RNASeq';
+activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (6 brain regions - No CBE)';
 thisFileName <- 'RNA-Seq_Meta_Analysis.Rmd'
@@
   lapply(downloadFile)
+
 # Read adjusted expression
-#_#expr.ids = c(ROSMAP = 'syn8456719', MSSM = 'syn8485027', MAYO = 'syn8466826')
 expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
@@
 }, expr)
-# Perform meta-analysis for AD-CONTROL comparison
-meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary, .(ensembl_gene_id), .fun = function(x){
+# Perform meta-analysis for AD-CONTROL comparison - Exclude CBE
+meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$Tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
   exp.effect = dplyr::filter(x, Diagnosis == 'AD')
@@
 
-#### Meta-analysis of AD-CONTROL (sex specific)
-```{r ad.cntrl.sex}
-# Mean and sd of gene expression in each tissue x diagnosis x sex group
-tissue.dx.sex.summary = plyr::ddply(covar, .(Tissue, Diagnosis, Sex), .fun = function(x, y){
-  data.frame(ensembl_gene_id = y$ensembl_gene_id,
-             n = dim(x)[1],
-             mn = rowMeans(y[,x$SampleID], na.rm = T),
-             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
-}, expr)
-# Perform meta-analysis for AD-CONTROL comparison (sex specific)
-meta.anlz.ad_cntrl.sex = plyr::ddply(tissue.dx.sex.summary, .(ensembl_gene_id, Sex), .fun = function(x){
-  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
-  rownames(exp.effect) = exp.effect$Tissue
-  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
-  rownames(cntrl.effect) = cntrl.effect$Tissue
-  cntrl.effect = cntrl.effect[rownames(exp.effect), ]
+#### Adjust the AD-Control Model for weights
+```{r weights}
+meta.anlz.ad_cntrl$TE.random.abs <- abs( meta.anlz.ad_cntrl$TE.random )
 
-  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
-                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
-                 studlab = exp.effect$Tissue,
-                 sm = 'SMD', method.smd = 'Hedges',
-                 method.tau = 'REML')
+#Assign Sig Or No
+meta.anlz.ad_cntrl$Sig <- 'NO'
+meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$pval.random < 0.05, ]$Sig <- 'YES'
 
-  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
-                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
-                   'Q', 'tau', 'H', 'I2')]))
-}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
-  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
-                fdr.random = p.adjust(pval.random, method = 'fdr'))
-p = list()
-p[[1]] = ggplot(meta.anlz.ad_cntrl.sex, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
-p[[1]] = p[[1]] + geom_point() + facet_grid(.~Sex)
-p[[2]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.fixed), x = TE.fixed))
-p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
-p[[3]] = ggplot(meta.anlz.ad_cntrl.sex, aes(y = -log10(fdr.random), x = TE.random))
-p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red') + facet_grid(.~Sex)
-ggpubr::ggarrange(plotlist = p, ncol = 1, nrow = 3)
-# Get gene sets
-tmp = meta.anlz.ad_cntrl.sex %>%
-  dplyr::select(ensembl_gene_id, Sex, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
-  tidyr::gather(Feature, Value, -ensembl_gene_id, -Sex) %>%
-  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
-  tidyr::spread(Feature1, Value) %>%
-  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
-tmp$Direction[tmp$TE <= 0] = 'DOWN'
-tmp$Direction[tmp$TE >= 0] = 'UP'
-gs = tmp %>%
-  plyr::dlply(.(Sex, Model, Direction), .fun = function(x){
-    unique(x$ensembl_gene_id)
-  })
-names(gs) = paste0('ad.control.', names(gs))
-all.gs = c(all.gs, gs)
-```
+meta.anlz.ad_cntrl <- meta.anlz.ad_cntrl[ order(meta.anlz.ad_cntrl$TE.random.abs), ]
+meta.anlz.ad_cntrl$y <- c( 1:dim(meta.anlz.ad_cntrl)[1] )/dim(meta.anlz.ad_cntrl)[1]
 
-#### Meta-analysis of APOE42-APOE40
-```{r apoe2.0}
-# Mean and sd of gene expression in each Tissue x APOE4 category
-tissue.apoe4.summary = plyr::ddply(covar, .(Tissue, APOE4), .fun = function(x, y){
-  data.frame(ensembl_gene_id = y$ensembl_gene_id,
-             n = dim(x)[1],
-             mn = rowMeans(y[,x$SampleID], na.rm = T),
-             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
-}, expr)
-# Perform meta-analysis for AD-CONTROL comparison (apoe4 2 vs 0)
-meta.anlz.apoe4_2_0 = plyr::ddply(tissue.apoe4.summary, .(ensembl_gene_id), .fun = function(x){
-  exp.effect = dplyr::filter(x, APOE4 == '2')
-  rownames(exp.effect) = exp.effect$Tissue
-  cntrl.effect = dplyr::filter(x, APOE4 == '0')
-  rownames(cntrl.effect) = cntrl.effect$Tissue
-  cntrl.effect = cntrl.effect[rownames(exp.effect), ]
+meta.anlz.ad_cntrl$Type <- 'Actual Rank'
 
-  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
-                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
-                 studlab = exp.effect$Tissue,
-                 sm = 'SMD', method.smd = 'Hedges',
-                 method.tau = 'REML')
+mylogit <- glm(meta.anlz.ad_cntrl$y  ~ abs( meta.anlz.ad_cntrl$TE.random.abs), data = meta.anlz.ad_cntrl, family = "binomial")
 
-  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
-                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
-                   'Q', 'tau', 'H', 'I2')]))
-}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
-  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
-                fdr.random = p.adjust(pval.random, method = 'fdr'))
-p = list()
-p[[1]] = ggplot(meta.anlz.apoe4_2_0, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))
-p[[1]] = p[[1]] + geom_point()
-p[[2]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.fixed), x = TE.fixed))
-p[[2]] = p[[2]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
-p[[3]] = ggplot(meta.anlz.apoe4_2_0, aes(y = -log10(fdr.random), x = TE.random))
-p[[3]] = p[[3]] + geom_point() + geom_hline(yintercept = -log10(0.05), color = 'red')
-ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
-# Get gene sets
-tmp = meta.anlz.apoe4_2_0 %>%
-  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
-  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
-  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
-  tidyr::spread(Feature1, Value) %>%
-  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
-tmp$Direction[tmp$TE <= 0] = 'DOWN'
-tmp$Direction[tmp$TE >= 0] = 'UP'
-gs = tmp %>%
-  plyr::dlply(.(Model, Direction), .fun = function(x){
-    unique(x$ensembl_gene_id)
-  })
-names(gs) = paste0('apoe4.2.0.', names(gs))
-all.gs = c(all.gs, gs)
+#Make Weight Template
+Work <- meta.anlz.ad_cntrl[, c('ensembl_gene_id','TE.random', 'seTE.random','fdr.random','TE.random.abs')]
+Work$log_abs_PVal <- abs( meta.anlz.ad_cntrl$TE.random.abs )
+colnames(Work)[ colnames(Work) == 'ensembl_gene_id' ] <- 'ENSG'
+
+#Calculate predicted Weights
+Work2 <- Work  
+Work2$Type <- "Predicted Weight"
+#Work2 <- Work2[ , c('ENSG', 'log_abs_PVal',"Type") ]
+Work2$y <- 1/(1+exp( -( mylogit$coefficients[1]+mylogit$coefficients[2]* abs( meta.anlz.ad_cntrl$TE.random.abs ) ) ))
+Work2$PreAdjWeight <- Work2$y
+
+#Add Actual Ranks to Template
+Work$y <- c( 1:dim(Work)[1] )/dim(Work)[1]
+Work$Type <- 'Actual Rank'
+mylogit <- glm(Work$y  ~ abs( Work$TE.random.abs), data = Work, family = "binomial")
+Work$PreAdjWeight <- Work$y
+
+Weights <- as.data.frame( rbind(Work,Work2) )
+Weights[ Weights$fdr.random > 0.05,]$y <- 0
+Weights <- Weights[ order(-Weights$y), ]
+
+#Annotate with Gene Names
+library('biomaRt')
+mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
+genes <- as.character(Weights$ENSG[ !duplicated(Weights$ENSG) ])
+G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol"),values=genes,mart= mart)
+
+G_list$ensembl_gene_id <- as.character(G_list$ensembl_gene_id)
+G_list$hgnc_symbol <- as.character(G_list$hgnc_symbol)
+
+G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]
+Miss <- cbind( ensembl_gene_id = as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG),
+               hgnc_symbol = rep("NA", length(as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG)) )
+               )
+G_list <- rbind( Miss, G_list)
+G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]
+
+row.names(G_list) <- G_list$ensembl_gene_id
+Weights$Gene <- G_list[ as.character(Weights$ENSG),]$hgnc_symbol
+
+Weights <- Weights[, c( "ENSG", "Gene", "TE.random", "seTE.random", "fdr.random", 'TE.random.abs', "log_abs_PVal", "Type", "PreAdjWeight", "y" )]
 ```
@@
 ```{r syn.store, cache = FALSE, include=FALSE}
-# Create folder to store results in synapse
-#CODE = Folder(name = 'MetaAnalysis', parentId = parentId)
-#CODE = synStore(CODE)
 CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))
+Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )
 
-Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )
 # Write results to files
 fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'AD-Control meta-analysis across 7 brain regions hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
-
-# Write results to files
-fwrite(meta.anlz.ad_cntrl.sex, file = 'meta.anlz.ad_cntrl.sex.tsv', sep = '\t', row.names = F, quote = F)
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.sex.tsv', name = 'AD-Control meta-analysis across 7 brain regions (sex specific) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
-
-# Write results to files
-fwrite(meta.anlz.apoe4_2_0, file = 'meta.anlz.apoe4_2_0.sex.tsv', sep = '\t', row.names = F, quote = F)
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.apoe4_2_0.sex.tsv', name = 'AD-Control meta-analysis across 7 brain regions (APOE4) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
-
-# Write results to files
-save(list = c('all.gs'), file = 'all.gensets.RData')
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='all.gensets.RData', name = 'All gene sets (in RData format) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
-
-```
-
-```{r knit2synapse, eval=FALSE}
-synapseclient <- reticulate::import("synapseclient")
-syn_temp <- synapseclient$Synapse()
-syn_temp$login()
-
-setwd("~/TWAS/code/")
-source("~/TWAS/utilityFunctions/knitfile2synapseClient.R")
-source("~/TWAS/utilityFunctions/hook_synapseMdSyntax_plot.R")
-
-createAndKnitToFolderEntityClient(file = "metaAnalysisForDiffExp_hg38.Rmd",
-                                 parentId = "syn18936948",
-                                 folderName = "MetaAnalysis")
+ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
 ```

### 2020-09-16 code/ProteomicsProcessing.R

@@ -1,0 +1,1100 @@
+#Final Code to develop DE Proteomic Analysis for Agora
+#Note: Output will vary slightly between runs as a result of the bootstraping regression
+##Special Thanks to Eric Dammer and the Bioinformatics Team at Emory for an extrodinary amount of help
+
+library("readxl")
+library(sva)
+library(impute)
+library(variancePartition)
+library(doParallel)
+library(affy)
+library(impute)
+library(preprocessCore)
+library("doParallel")
+library(boot)
+library(WGCNA)
+#install.packages("BiocManager")
+#BiocManager::install("WGCNA")
+#sudo apt-get install libjpeg-dev
+library(Cairo)
+#Woof this Sucks
+#sudo apt-get install r-base-dev, sudo apt-get install libxt-dev
+#sudo apt-get install xfonts-base, sudo apt-get install xauth, sudo apt-get install libjpeg-dev
+#sudo apt-get install xvfb, sudo apt-get install libcairo2-dev, sudo apt-get install libgtk2.0-dev
+library(NMF)
+#install.packages('NMF')
+library(biomaRt)
+library(variancePartition)
+#BiocManager::install("variancePartition")
+library(limma)
+
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
+Process <- function( FILE, Opt ){
+  ##Processes the Raw data and collapse to GeneName|UniProtID 
+  #'@FILE matrix of raw protien.groups file
+  #'@Opt a one or a zero 
+  
+  options(stringsAsFactors = FALSE)
+  proteinGroups <- FILE #FILE
+  
+  decoyCount = length(which(proteinGroups$Reverse=="+"))
+  originalRows = nrow(proteinGroups)
+  decoyIndices = which(proteinGroups$Reverse=="+")
+  FDR = paste0(round(length(decoyIndices)/nrow(proteinGroups)*100, 2), "% FDR")
+  LFQindices = which(grepl("LFQ.intensity.", colnames(proteinGroups)))
[…]

### 2020-09-21 code/RNA-Seq_Meta_Analysis.Rmd
Final  Proteomics

@@
 # Write results to files
-fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+fwrite(meta.anlz.ad_cntrl, file = 'RNASeq_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='RNASeq_meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+
+# Write results to files
+fwrite(Weights, file = 'RNASeq_meta_weights.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='RNASeq_meta_weights.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis Weights', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
 ```

### 2020-09-21 code/ProteomicsProcessing.R

@@ -1092,8 +1092,582 @@ colnames(LF_DE)[c(1,2)] <- c('GName','ProtID')
 LF_DE <- as.data.frame(LF_DE,  stringsAsFactors =F )
 LF_DE$GName <- as.character(LF_DE$GName)
 
+####### Assign Averaged Weight Values, and Denote the Bidirectional Genes
+Process_Meta <- LF_DE
+table(LF_DE$GName)[ table(LF_DE$GName) > 8]
+LF_DE[ LF_DE$GName == 'TPM1', ] 
+
+Process_Meta$ENSG <- NA 
+
+table( LF_DE[ table(LF_DE$GName)[ table(LF_DE$GName) > 1], ]$Rand_Sig )
+# NO YES 
+# 888 147 
+
+Dups <- names( table(LF_DE$GName)[ table(LF_DE$GName) > 1] )
+Dups <- Dups[ !duplicated(Dups) ]
+
+Col2Av <- c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed', 'TE.random',
+            'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random', 'Q', 'tau', 'H', 'I2', 
+            'fdr.fixed', 'fdr.random')
+
+IndiciesToToss <- NULL
+Process_Meta$Averaged <- 'NO'
+Process_Meta$BiDirectional <- 'NO'
+AbigousRows <- NULL
+AbigousIDs <- NULL
+
+#Fix the TUBAL3/TBAL3 issue (Same ProtID but one is given deprecated Gene Name)
+Process_Meta[ Process_Meta$GName=='TBAL3' & Process_Meta$peptide_id=='TBAL3|A6NHL2-2' ,  ]$GName <- 'TUBAL3'
+Process_Meta[ Process_Meta$GName=='TUBAL3' & Process_Meta$peptide_id=='TBAL3|A6NHL2-2' ,  ]$peptide_id <- 'TUBAL3|A6NHL2-2'
+
+#Fix Deprecated HIST1H4A  Gene Name
+Process_Meta[ Process_Meta$GName=='HIST1H4A' & Process_Meta$peptide_id=='HIST1H4A|P62805' ,  ]$GName <- 'H4C1'
+Process_Meta[ Process_Meta$GName=='H4C1' & Process_Meta$peptide_id=='HIST1H4A|P62805' ,  ]$peptide_id <- 'H4C1|P62805'
[…]

### 2020-09-21 code/ProteomicsProcessing.R
Final  Proteomics

@@ -1648,8 +1648,8 @@ activityDescription = 'Fixed and random effect meta-analysis of AMP-AD Proteomic
 thisFileName <- 'ProteomicsProcessing.R'
 
 # Github link
-thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
-thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
+thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
+thisFile <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
 
 CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))
 Syns_Used <- c( 'syn18914942', 'syn18914952', 'syn18914920', 'syn9637748', 'syn6100414',
@@ -1657,14 +1657,14 @@ Syns_Used <- c( 'syn18914942', 'syn18914952', 'syn18914920', 'syn9637748', 'syn6
                 'syn18914936', 'syn18914939' )
 
 # Write results to files
-fwrite(LF_DE, file = 'Raw_Proteomics_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+data.table::fwrite(LF_DE, file = 'Raw_Proteomics_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
 ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='Raw_Proteomics_meta.anlz.ad_cntrl.tsv', name = 'Proteomics AD-Control meta-analysis across 4 brain regions', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
 
-fwrite(Process_Meta, file = 'Processed_Proteomics_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+data.table::fwrite(Process_Meta, file = 'Processed_Proteomics_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
[…]

### 2020-09-23 code/Omics_Harness.R
Omics Score 2.0

@@ -1,0 +1,264 @@
+# Code to Combine RNA and Proteomics Weights and Apply Scoring Harness
+library(biomaRt)
+
+reticulate::use_python("/usr/bin/python", required = TRUE)
+synapseclient <- reticulate::import("synapseclient")
+syn_temp <- synapseclient$Synapse()
+syn_temp$login()
+
+Prot <- read.table( syn_temp$get('syn22686575')$path, header=T, sep='\t', stringsAsFactors=F)
+RNA  <- read.table( syn_temp$get('syn22414716')$path, header=T, sep='\t', stringsAsFactors=F)
+
+ensgs <- c( Prot$ENSG, RNA$ENSG) 
+ensgs <- ensgs[ !duplicated(ensgs) ]
+
+#Build Combined File Template
+Comb <- as.data.frame( matrix(NA, length(ensgs), 12) )
+rownames(Comb) <- ensgs
+colnames( Comb ) <- c( 'ENSG', 'GName',
+                        'RNA_TE', 'RNA_fdr_CorPVal', 'RNA_Sig', 'RNA_Direction', 'RNA_Weight',
+                        'Pro_TE', 'Pro_fdr_CorPVal', 'Pro_Sig', 'Pro_Direction', 'Pro_Weight'
+                        )
+
+#Seed the Dataframe
+Comb$ENSG <- row.names( Comb )
+
+prot <- Prot[ Prot$Type == 'Predicted Weight' , ]
+row.names(prot) <- prot$ENSG
+
+rna <- RNA[ RNA$Type == 'Predicted Weight' , ]
+row.names(rna) <- rna$ENSG
+
+#ProtVals
+Comb[ row.names(prot), ]$Pro_Direction <- prot$Direction
+Comb[ row.names(prot), ]$Pro_fdr_CorPVal <- prot$fdr.random
+Comb[ row.names(prot), ]$Pro_Sig <- prot$Sig
+Comb[ row.names(prot), ]$Pro_TE <- prot$TE.random
+Comb[ row.names(prot), ]$GName <- prot$GName
+Comb[ row.names(prot), ]$Pro_Weight <- prot$y
+
+#RNA Vals
+Comb[ row.names(rna), ]$RNA_fdr_CorPVal <- rna$fdr.random
+Comb[ row.names(rna), ]$RNA_TE <- rna$TE.random
+Comb[ row.names(rna), ]$GName <- rna$Gene
[…]

### 2020-09-23 code/Omics_Harness.R
Omics Score 2.0

@@ -217,6 +217,7 @@ activityDescription = 'Omics Scoring harnesss based on RNA-Seq and Proteomics me
 thisFileName <- 'Omics_Harness.R'
 
 # Github link
+library(githubr)
 thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
 thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
 

### 2020-09-23 code/Omics_Harness.R
Omics Score 2.0

@@ -262,4 +262,28 @@ ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='~/AD_TargetRank/OmicsSc
                                activityName = activityName, 
                                executed = thisFile, 
                                activityDescription = activityDescription)
+#Push Table
+cols = list( synapseclient$Column( name='ENSG', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='GName', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='RNA_TE', columnType='DOUBLE'),
+             synapseclient$Column( name='RNA_fdr_CorPVal', columnType='DOUBLE'),
+             synapseclient$Column( name='RNA_Sig', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='RNA_Direction', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='RNA_Weight', columnType='DOUBLE'),
+             synapseclient$Column( name='Pro_TE', columnType='DOUBLE'),
+             synapseclient$Column( name='Pro_fdr_CorPVal', columnType='DOUBLE'),
+             synapseclient$Column( name='Pro_Sig', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='Pro_Direction', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='Pro_Weight', columnType='DOUBLE'),
+             synapseclient$Column( name='Harness', columnType='DOUBLE'),
+             synapseclient$Column( name='Rank', columnType='DOUBLE'),
+             synapseclient$Column( name='Predicted_Weight', columnType='DOUBLE'),
+             synapseclient$Column( name='TYPE', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='Final_Weight', columnType='DOUBLE'),
+             synapseclient$Column( name='OmicsScore', columnType='DOUBLE')
+           )
 
+schema = synapseclient$Schema(name='Omics Scores Version 2.0', columns=cols, parent='syn22351719')
+table = synapseclient$Table(schema, "OmicsScores.csv")
+table = syn_temp$store(table)
+  

### 2020-09-23 code/Omics_Harness.R
Omics Score 2.0

@@ -263,27 +263,51 @@ ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='~/AD_TargetRank/OmicsSc
                                executed = thisFile, 
                                activityDescription = activityDescription)
 #Push Table
+Comb$RNA_TE <- as.double(Comb$RNA_TE) 
+Comb$RNA_fdr_CorPVal <- as.double(Comb$RNA_fdr_CorPVal) 
+Comb$RNA_Weight <- as.double(Comb$RNA_Weight) 
+
+Comb$Pro_TE <- as.double(Comb$Pro_TE) 
+Comb$Pro_fdr_CorPVal <- as.double(Comb$Pro_fdr_CorPVal) 
+Comb$Pro_Weight <- as.double(Comb$Pro_Weight) 
+Comb$Harness <- as.double(Comb$Harness) 
+Comb$Rank <- as.double(Comb$Rank) 
+
+Comb$Predicted_Weight <- as.double(Comb$Predicted_Weight) 
+Comb$Final_Weight <- as.double(Comb$Final_Weight) 
+Comb$OmicsScore <- as.double(Comb$OmicsScore) 
+
+Comb$Pro_fdr_CorPVal <- signif(Comb$Pro_fdr_CorPVal , digits = 6)
+Comb$RNA_fdr_CorPVal <- signif(Comb$RNA_fdr_CorPVal , digits = 6)
+Comb$Harness <- signif(Comb$Harness , digits = 6)
+Comb$Predicted_Weight <- signif(Comb$Predicted_Weight , digits = 6)
+Comb$RNA_TE <- signif(Comb$RNA_TE , digits = 6)
+Comb$Pro_TE   <- signif(Comb$Pro_TE   , digits = 6)
+
 cols = list( synapseclient$Column( name='ENSG', columnType='STRING', maximumSize=20),
-             synapseclient$Column( name='GName', columnType='STRING', maximumSize=20),
-             synapseclient$Column( name='RNA_TE', columnType='DOUBLE'),
-             synapseclient$Column( name='RNA_fdr_CorPVal', columnType='DOUBLE'),
+             synapseclient$Column( name='GName', columnType='STRING', maximumSize=40),
+             synapseclient$Column( name='RNA_TE', columnType='STRING', maximumSize=20),
+             synapseclient$Column( name='RNA_fdr_CorPVal', columnType='STRING', maximumSize=20),
              synapseclient$Column( name='RNA_Sig', columnType='STRING', maximumSize=20),
[…]

### 2021-02-05 code/RNA-Seq_Meta_Analysis.Rmd
Update RNA SeqMeta

@@
 # Read covariates
-covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
+#covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
+covar.ids = c(ROSMAP = 'syn23643706', MSSM = 'syn8484996', MAYO = 'syn8466814')
 all.used.ids = as.character(covar.ids)
@@
 # Read adjusted expression
-expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
+#expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
+expr.ids = c(ROSMAP = 'syn23643727', MSSM = 'syn21285564', MAYO = 'syn21291908')
 all.used.ids = c(all.used.ids, as.character(expr.ids))
@@
 covar$ROSMAP = covar$ROSMAP %>%
-  dplyr::select(SampleID, Sex, Diagnosis, APOE4) %>%
-  dplyr::mutate(Tissue = 'DLPFC')
+  dplyr::rename(Sex = sex) %>%
+  dplyr::rename(APOE4 = apoe4) %>%
+  tidyr::separate(Tissue.Diagnosis, c('Tissue', 'Diagnosis'), sep = '\\.') %>%
+  tidyr::separate(Tissue.Diagnosis.SEX, c('Tissue1', 'Diagnosis1', 'Sex1'), sep = '\\.') %>%
+  dplyr::select(SampleID, Sex, Tissue1, Diagnosis, Tissue, APOE4) #%>%
+  #dplyr::mutate(Tissue = 'DLPFC')
 covar$MSSM = covar$MSSM %>%

### 2021-11-17 code/RNA-Seq_Meta_Analysis.Rmd
New version RNA-Seq Weights

--- 2020-09-16 code/RNA-Seq_Meta_Analysis.Rmd
@@
 ```{r synapse.parameters, include=FALSE, cache=TRUE}
-reticulate::use_python("/usr/bin/python", required = TRUE)
-synapseclient <- reticulate::import("synapseclient")
-syn_temp <- synapseclient$Synapse()
-syn_temp$login()
-
-#parentId = 'syn18936948';
+synapser::synLogin()
 parentId = 'syn22351719';
@@
 # Function to read tsv/csv files from synapse
-downloadFile <- function(id){ 
-  fread(syn_temp$get(id)$path, data.table = F)
+downloadFile <- function(id, tissue){ 
+  dat <- fread(synapser::synGet(id)$path, data.table = F)
+  dat$tissue <- tissue
+  return(data)
 }
 # Read covariates
-covar.ids = c(ROSMAP = 'syn8456631', MSSM = 'syn8484996', MAYO = 'syn8466814')
+covar.ids = c(ROSMAP_DLPFC = 'syn26340637 ', 
+              ROSMAP_PCC = 'syn26350606',
+              ROSMAP_ACC = 'syn26374583 ',
+              MSSM_FP = 'syn26453689', 
+              MSSM_IFG = 'syn26454154', 
+              MSSM_PHG = 'syn26453943', 
+              MSSM_STG = 'syn26454639', 
+              MAYO_TCX = 'syn26374583')
+#_tissue
+
 all.used.ids = as.character(covar.ids)
@@
   lapply(downloadFile)
+covar$ROSMAP_DLPFC$tissue <- 'DLPFC'
+covar$ROSMAP_PCC$tissue <- 'PCC'
+covar$ROSMAP_ACC$tissue <- 'ACC'
+covar$MSSM_FP$tissue <- 'FP'
+covar$MSSM_IFG$tissue <- 'IFG'
+covar$MSSM_PHG$tissue <- 'PHG'
+covar$MSSM_STG$tissue <- 'STG'
+covar$MAYO_TCX$tissue <- 'TCX'
 
 # Read adjusted expression
-expr.ids = c(ROSMAP = 'syn21292041', MSSM = 'syn21285564', MAYO = 'syn21291908')
+expr.ids = c(ROSMAP_DLPFC = 'syn26340641', 
+              ROSMAP_PCC = 'syn26374587',
+              ROSMAP_ACC = 'syn26350610',
+              MSSM_FP = 'syn26453693', 
+              MSSM_IFG = 'syn26454158', 
+              MSSM_PHG = 'syn26453947', 
+              MSSM_STG = 'syn26454643', 
+              MAYO_TCX = 'syn26387667')
+
 all.used.ids = c(all.used.ids, as.character(expr.ids))
+downloadFile <- function(id, tissue){ 
+  fread(synapser::synGet(id)$path, data.table = F)
+}
 expr =  expr.ids %>%
   lapply(downloadFile) %>%
-  plyr::join_all(type = 'inner')
+  plyr::join_all(type = 'full')
 ```
@@
 ```{r modify.covar}
-covar$ROSMAP = covar$ROSMAP %>%
-  dplyr::select(SampleID, Sex, Diagnosis, APOE4) %>%
-  dplyr::mutate(Tissue = 'DLPFC')
-covar$MSSM = covar$MSSM %>%
-  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
-  dplyr::rename(Sex = Sex) %>%
-  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
-  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
-covar$MAYO = covar$MAYO %>%
-  dplyr::select(SampleID, Sex, Tissue.Diagnosis, Tissue.APOE4) %>%
-  tidyr::separate(Tissue.Diagnosis, c('Tissue1', 'Diagnosis'), sep = '\\.') %>%
-  tidyr::separate(Tissue.APOE4, c('Tissue', 'APOE4'), sep = '\\.')
 covar = covar %>%
-  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study') %>%
-  dplyr::select(-Tissue1)
+  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study')
 ```
@@
 # Mean and sd of gene expression in each tissue x diagnosis group
-tissue.dx.summary = plyr::ddply(covar, .(Tissue, Diagnosis), .fun = function(x, y){
-  data.frame(ensembl_gene_id = y$ensembl_gene_id,
+tissue.dx.summary = plyr::ddply(covar, .(tissue, diagnosis), .fun = function(x, y){
+  data.frame(ensembl_gene_id = y$feature,
              n = dim(x)[1],
-             mn = rowMeans(y[,x$SampleID], na.rm = T),
-             sd = apply(y[,x$SampleID], 1, sd, na.rm = T))
+             mn = rowMeans(y[,x$specimenID], na.rm = T),
+             sd = apply(y[,x$specimenID], 1, sd, na.rm = T))
 }, expr)
 # Perform meta-analysis for AD-CONTROL comparison - Exclude CBE
-meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$Tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
-  exp.effect = dplyr::filter(x, Diagnosis == 'AD')
-  rownames(exp.effect) = exp.effect$Tissue
-  cntrl.effect = dplyr::filter(x, Diagnosis == 'CONTROL')
-  rownames(cntrl.effect) = cntrl.effect$Tissue
+meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
+  exp.effect = dplyr::filter(x, diagnosis == 'AD')
+  rownames(exp.effect) = exp.effect$tissue
+  cntrl.effect = dplyr::filter(x, diagnosis == 'CT')
+  rownames(cntrl.effect) = cntrl.effect$tissue
   cntrl.effect = cntrl.effect[rownames(exp.effect), ]
@@
 ```{r weights}
+meta.anlz.ad_cntrl <- meta.anlz.ad_cntrl[complete.cases(meta.anlz.ad_cntrl),]
 meta.anlz.ad_cntrl$TE.random.abs <- abs( meta.anlz.ad_cntrl$TE.random )
@@
 meta.anlz.ad_cntrl$Sig <- 'NO'
-meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$pval.random < 0.05, ]$Sig <- 'YES'
+meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$fdr.random < 0.05, ]$Sig <- 'YES'
 
@@
 
-#Annotate with Gene Names
-library('biomaRt')
-mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
-genes <- as.character(Weights$ENSG[ !duplicated(Weights$ENSG) ])
-G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","hgnc_symbol"),values=genes,mart= mart)
+G_list <- fread(synapser::synGet('syn26453691')$path)
+all.used.ids <- c(all.used.ids,'syn26453691')
 
+G_list <- as.data.frame(G_list)
 G_list$ensembl_gene_id <- as.character(G_list$ensembl_gene_id)
@@
 G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]
-Miss <- cbind( ensembl_gene_id = as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG),
-               hgnc_symbol = rep("NA", length(as.character(Weights[ (Weights$ENSG %in% G_list$ensembl_gene_id) == F,]$ENSG)) )
-               )
-G_list <- rbind( Miss, G_list)
+
 G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]
@@
 ```{r syn.store, cache = FALSE, include=FALSE}
-CODE <- syn_temp$store(synapseclient$Folder(name = "MetaAnalysis", parentId = parentId))
-Syns_Used <- c( 'syn8456631', 'syn8484996', 'syn8466814', 'syn21292041', 'syn21285564', 'syn21291908' )
+CODE <- synapser::synStore(synapser::Folder(name = "MetaAnalysis", parentId = parentId))
+Syns_Used <- all.used.ids
 
 # Write results to files
-fwrite(meta.anlz.ad_cntrl, file = 'meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
-ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+fwrite(meta.anlz.ad_cntrl, file = 'RNASeq_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
+
+# Write results to files
+fwrite(Weights, file = 'RNASeq_meta_weights.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
+ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta_weights.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis Weights', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
 ```

### 2021-11-17 code/RNA-Seq_Meta_Analysis.Rmd
New version RNA-Seq Weights

@@
 library(foreach)
-cl = makeCluster(detectCores()-2)
-registerDoParallel(cl)
+#cl = makeCluster(detectCores()-2)
+#registerDoParallel(cl)
 options(xtable.type="html")
@@
 
+covar = covar %>%
+  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study')
+
 # Read adjusted expression
@@
   plyr::join_all(type = 'full')
-```
-
-```{r modify.covar}
-covar = covar %>%
-  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study')
 ```

### 2021-11-17 code/RNA-Seq_Meta_Analysis.Rmd
New version RNA-Seq Weights

@@
 # Function to read tsv/csv files from synapse
-downloadFile <- function(id, tissue){ 
-  dat <- fread(synapser::synGet(id)$path, data.table = F)
-  dat$tissue <- tissue
-  return(data)
+downloadFile <- function(id){ 
+  fread(synapser::synGet(id)$path, data.table = F)
 }
@@
 all.used.ids = c(all.used.ids, as.character(expr.ids))
-downloadFile <- function(id, tissue){ 
-  fread(synapser::synGet(id)$path, data.table = F)
-}
+
 expr =  expr.ids %>%

### 2021-11-17 code/RNA-Seq_Meta_Analysis.Rmd
New version RNA-Seq Weights

@@
 library(foreach)
-#cl = makeCluster(detectCores()-2)
-#registerDoParallel(cl)
+
 options(xtable.type="html")

### 2021-11-17 code/Omics_Harness.R
New version RNA-Seq Weights

@@ -345,7 +345,30 @@ cols = list( synapser::Column( name='ENSG', columnType='STRING', maximumSize=20)
 
 write.csv( Comb, 'OmicsScores.csv', 
            row.names = F, quote = F)
-schema = synapser::Schema(name='Omics Scores Version 3.0', columns=cols, parent='syn21532474')
-table = synapser::Table(schema, "OmicsScores.csv")
-synapser::synStore(table)
+
+# Pull Project
+project_treatad <- synapser::synGetEntity('syn21532474')
+
+# Pull Omics Score Table
+omics_table <- synapser::synTableQuery(sprintf("select * from %s ", 'syn22758536'))
+
+# Create a snapshot of the Table Entity
+body_json <- rjson::toJSON(c(snapshotComment = "Added TMT and SageSeqR Processed", snapshotLabel = "v4"))
+snapshot <-  synapser::synRestPOST(paste0("/entity/", omics_table$tableId, "/table/snapshot"), body = body_json)
+
+# Pull Version 3 of the Omics scores:
+omics_table <- synapser::synTableQuery(sprintf("select * from %s ", 'syn22758536'))
+scores <- Comb #read.csv(synapser::synGet('syn22758171', version=2)$path)
+
+# Change the entire Table Entity
+deleted <- synapser::synDelete(omics_table)
+scores <- scores[, c('ENSG', 'GName', 'RNA_TE', 'RNA_fdr_CorPVal', 'RNA_Sig', 'RNA_Direction',
+                     'RNA_Weight',	'Pro_TE',	'Pro_fdr_CorPVal',	'Pro_Sig',	'Pro_Direction',
+                     'Pro_Weight',	'Harness',	'Rank',	'Predicted_Weight',	'TYPE',	'Final_Weight',
+                     'OmicsScore')]
+scores$RNA_TE <- signif(scores$RNA_TE,5)
+scores$Pro_TE <- signif(scores$Pro_TE,5)
+
+synapser::synStore(synapser::Table(omics_table$tableId, scores))
+
   

### 2022-02-25 code/RNA-Seq_Meta_Analysis.Rmd
update the weights

@@
 # Read covariates
-covar.ids = c(ROSMAP_DLPFC = 'syn26340637 ', 
-              ROSMAP_PCC = 'syn26350606',
-              ROSMAP_ACC = 'syn26374583 ',
-              MSSM_FP = 'syn26453689', 
-              MSSM_IFG = 'syn26454154', 
-              MSSM_PHG = 'syn26453943', 
-              MSSM_STG = 'syn26454639', 
-              MAYO_TCX = 'syn26374583')
+#covar.ids = c(ROSMAP_DLPFC = 'syn26340637 ', 
+#              ROSMAP_PCC = 'syn26350606',
+#              ROSMAP_ACC = 'syn26374583 ',
+#              MSSM_FP = 'syn26453689', 
+#              MSSM_IFG = 'syn26454154', 
+#              MSSM_PHG = 'syn26453943', 
+#              MSSM_STG = 'syn26454639', 
+#              MAYO_TCX = 'syn26374583')
+
+covar.ids = c(ROSMAP = 'syn26967450 ', 
+              MSSM = 'syn27068753', 
+              MAYO = 'syn27024950')
 #_tissue
@@
   lapply(downloadFile)
-covar$ROSMAP_DLPFC$tissue <- 'DLPFC'
-covar$ROSMAP_PCC$tissue <- 'PCC'
-covar$ROSMAP_ACC$tissue <- 'ACC'
-covar$MSSM_FP$tissue <- 'FP'
-covar$MSSM_IFG$tissue <- 'IFG'
-covar$MSSM_PHG$tissue <- 'PHG'
-covar$MSSM_STG$tissue <- 'STG'
-covar$MAYO_TCX$tissue <- 'TCX'
+#covar$ROSMAP_DLPFC$tissue <- 'DLPFC'
+#covar$ROSMAP_PCC$tissue <- 'PCC'
+#covar$ROSMAP_ACC$tissue <- 'ACC'
+#covar$MSSM_FP$tissue <- 'FP'
+#covar$MSSM_IFG$tissue <- 'IFG'
+#covar$MSSM_PHG$tissue <- 'PHG'
+#covar$MSSM_STG$tissue <- 'STG'
+#covar$MAYO_TCX$tissue <- 'TCX'
 
@@
 # Read adjusted expression
-expr.ids = c(ROSMAP_DLPFC = 'syn26340641', 
-              ROSMAP_PCC = 'syn26374587',
-              ROSMAP_ACC = 'syn26350610',
-              MSSM_FP = 'syn26453693', 
-              MSSM_IFG = 'syn26454158', 
-              MSSM_PHG = 'syn26453947', 
-              MSSM_STG = 'syn26454643', 
-              MAYO_TCX = 'syn26387667')
+#expr.ids = c(ROSMAP_DLPFC = 'syn26340641', 
+#              ROSMAP_PCC = 'syn26374587',
+#              ROSMAP_ACC = 'syn26350610',
+#              MSSM_FP = 'syn26453693', 
+#              MSSM_IFG = 'syn26454158', 
+#              MSSM_PHG = 'syn26453947', 
+#              MSSM_STG = 'syn26454643', 
+#              MAYO_TCX = 'syn26387667')
+expr.ids = c(ROSMAP_DLPFC = 'syn26967454', 
+              MSSM_STG = 'syn27068759', 
+              MAYO_TCX = 'syn27024966')
 
@@
   plyr::join_all(type = 'full')
+#expr <- as.data.frame(expr) %>%
+#  column_to_rownames('feature')
+
 ```
@@
 Weights <- Weights[, c( "ENSG", "Gene", "TE.random", "seTE.random", "fdr.random", 'TE.random.abs', "log_abs_PVal", "Type", "PreAdjWeight", "y" )]
+
+Weights$plotname <- NA
+label_name <- c( "MTCH1", "MTOR", "APPL1", "BAX", "STMN2", "SLC9A9", "TOMM40L", "GABRA2",  "MTCH2"  )
+Weights[ Weights$Gene %in% label_name, ]$plotname <- Weights[ Weights$Gene %in% label_name, ]$Gene
+
+A <- Weights[ Weights$Type == 'Actual Rank', ]
+row.names(A) <- A$ENSG
+colnames(A)[colnames(A)=='PreAdjWeight'] <- 'PreAdjWeight_Actual'
+colnames(A)[colnames(A)=='y'] <- 'y_Actual'
+
+B <- Weights[ Weights$Type == 'Predicted Weight', ]
+row.names(B) <- B$ENSG
+colnames(B)[colnames(B)=='PreAdjWeight'] <- 'PreAdjWeight_Predicted'
+colnames(B)[colnames(B)=='y'] <- 'y_Predicted'
+
+plotframe <- as.data.frame(cbind(A[,c('ENSG','Gene', 'TE.random', 'seTE.random', 'fdr.random', 'TE.random.abs', 'log_abs_PVal', 'PreAdjWeight_Actual','y_Actual', 'plotname')], B[row.names(A), c('PreAdjWeight_Predicted','y_Predicted')]))
+
+plotframe <- plotframe[ plotframe$fdr.random < 0.05, ]
+plotframe$logTE <- log(abs(plotframe$TE.random.abs))
+
+library(ggplot2)
+library(ggrepel)
+ggplot2::ggplot(data=plotframe) + 
+  ggplot2::geom_point( ggplot2::aes(x=logTE, y=y_Predicted)) +
+  ggrepel::geom_label_repel(data = plotframe, 
+                            ggplot2::aes(x=logTE, y=y_Predicted, label = plotname, size = NULL, color = NULL),
+                            nudge_x = .45,
+                            nudge_y = -.05,
+                            segment.size  = 0.2,
+                            segment.color = "grey50",
+                            direction     = "x"
+) +
+  ggplot2::xlab("Log2( absolute(Effect Size))") + 
+  ggplot2::ylab("Assigned weight") +
+  ggplot2::labs(title="Weights of Differentially Expressed Transcripts") + 
+  ggplot2::theme(plot.title =  ggplot2::element_text(hjust = 0.5))
+
 ```

### 2022-02-25 code/Omics_Harness.R

@@ -177,13 +177,26 @@ Comb[ Comb$Pro_Sig %in% 'YES' & Comb$RNA_Sig %in% 'YES', ]$TYPE <- 'Both_Sig'
 Comb[ (Comb$Pro_Sig %in% 'NO' | is.na(Comb$Pro_Sig)) & (Comb$RNA_Sig %in% 'NO' | is.na(Comb$RNA_Sig)), ]$TYPE <- 'None_Sig'
 Comb$TYPE<-as.factor(Comb$TYPE)
 
+Comb$plotname <- NA
+#label_name <- c('CPEB1', 'SIRT2', 'DLG5', 'MTCH1', 'MTOR', 'ADAM17', 'APPL1', 'CNTNAP2', 'NEFL')
+label_name <- c("MTCH1", "MTOR", "APPL1", 'BAX','STMN2', 'SLC9A9','TOMM40L','GABRA2', 'MTCH2')
+Comb[ Comb$GName %in% label_name, ]$plotname <- Comb[ Comb$GName %in% label_name, ]$GName
+
 ggplot( data=Comb ) + 
   geom_line( aes( y = test, x=Harness  )) +
   geom_jitter( aes( y = test, x=Harness, col= TYPE),alpha = 0.2, width=0.05) 
 
 ggplot( data=Comb[ Comb$RNA_Sig %in% 'YES' | Comb$Pro_Sig %in% 'YES', ] ) + 
   geom_line( aes( y = test, x=Harness  )) +
-  geom_jitter( aes( y = test, x=Harness, col= TYPE),alpha = 0.4, shape = 16, size = .5, width=0.05) 
+  geom_jitter( aes( y = test, x=Harness, col= TYPE),alpha = 0.4, shape = 16, size = .5, width=0.05) +
+  ggrepel::geom_label_repel(data = Comb[ Comb$RNA_Sig %in% 'YES' | Comb$Pro_Sig %in% 'YES', ] , 
+                            ggplot2::aes(x=Harness, y=test, label = plotname, size = NULL, color = NULL),
+                            nudge_x = .45,
+                            nudge_y = -.15,
+                            segment.size  = 0.2,
+                            segment.color = "grey50",
+                            direction     = "x"
+  )
 
 
 Comb$TYPE<-as.character(Comb$TYPE)
@@ -219,7 +232,17 @@ pdf(paste0(plots,'Target_Type_Filt.pdf'))
[…]

### 2022-02-25 code/Proteomics_Metaanalysis_V2.R

@@ -1,9 +1,9 @@
 ### Proteomics Meta Analysis Version 2
-
+library(dplyr)
 source('utilityFunctions/loadsyndata.R')
 system( 'curl http://www.compgen.pitt.edu/GemTools/GemTools.R > GemTools.R')
 source('GemTools.R')
-
+synapser::synLogin()
 load(synapser::synGet('syn26427476')$path)
 load(synapser::synGet('syn26433992')$path)
 
@@ -316,6 +316,44 @@ tWork$Sig <- ifelse(tWork$fdr.random < 0.05, "YES", "NO" )
 tWork <- tWork[ order(-abs(tWork$y)), ]
 tWork$y_Final <- tWork$y
 tWork[ tWork$Sig == 'NO',]$y_Final <- 0
+
+tWork$plotname <- NA
+label_name <- c( "MTCH1", "MTOR", "APPL1", "BAX", "STMN2", "SLC9A9", "TOMM40L", "GABRA2",  "MTCH2"  )
+tWork[ tWork$SYMBOL %in% label_name, ]$plotname <- tWork[ tWork$SYMBOL %in% label_name, ]$SYMBOL
+
+A <- tWork[ tWork$Type == 'Actual Rank', ]
+row.names(A) <- A$ENSG
+colnames(A)[colnames(A)=='y_Final'] <- 'y_Final_Actual'
+colnames(A)[colnames(A)=='y'] <- 'y_Actual'
+
+B <- tWork[ tWork$Type == 'Predicted Weight', ]
+row.names(B) <- B$ENSG
+colnames(B)[colnames(B)=='y_Final'] <- 'y_Final_Predicted'
+colnames(B)[colnames(B)=='y'] <- 'y_Predicted'
+
+plotframe <- as.data.frame(cbind(A[,c('peptide_id', 'EnsemblGene','SYMBOL', 'TE.random', 'seTE.random', 'fdr.random', 'abs_TE', 'y_Actual','y_Final_Actual', 'plotname')], B[row.names(A), c('y_Predicted','y_Final_Predicted')]))
+
+plotframe <- plotframe[ plotframe$fdr.random < 0.05, ]
+plotframe$logTE <- log(abs(plotframe$abs_TE))
+
+library(ggplot2)
+library(ggrepel)
+ggplot2::ggplot(data=plotframe) + 
+  ggplot2::geom_point( ggplot2::aes(x=logTE, y=y_Predicted)) +
+  ggrepel::geom_label_repel(data = plotframe, 
[…]
